### 2016-05-04 main.tex
Base files update

\documentclass[12pt,draftcls,onecolumn]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[dvips]{graphicx}
% Title Page
\title{Understanding Loss recovery in MultiPath TCP }
\author{KAU}

\begin{document}
\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}
TCP has two mechanisms for detecting and recovering from packet losses namely Fast Recovery (FR) and Retransmission Timeouts(RTOs). For long flows where there are sufficiently large number of packets 
to be transferred, FR is quicker than RTOs in detecting and recovering packet losses. On the other hand shorter flows which are a majority in the web traffic have to depend on RTOs for packet loss detection 
and recovery incurring more latency. A single packet loss in a short flow may take many RTTs to detect and recover. This scenario is also applicable to the packets at the end of the flow (or tail) in a long 
flow.

TCP Loss Probe (TLP)~\cite{ietftlp} a mechanism that allows flows to detect and recover from tail losses much faster than an RTO, thereby speeding up short transfers. With TLP a packet loss in the middle 
of a packet train as well as at the tail end will now trigger the same fast recovery mechanisms. It assumes other algorithms such as early retransmit~\cite{rfc5827} and FACK threshold based recovery are 
present.

In Multi-path-TCP(MPTCP)~\cite{rfc6824}, a connection can have multiple TCP sub flows using different interfaces on different routes. Packet losses in each sub flow are assumed to be detected and recovered 
in a similar fashion as that of TCP. However, It is not completely clear how the loss recovery happens in the implementation and which sub-flow retransmits the lost packets. If the recovery is handled at the 
meta level, the lost packet may be rescheduled and retransmitted at the available sub flow with lowest RTT. If the recovery is handled at the flow level, the packet may be retransmitted in the same sub-flow.

Through this set of preliminary experiments, we would like to understand the behavior of MPTCP with and without TLP and ER by monitoring burst completion time, congestion window evolution and timeout 
behavior.
\section{Related Work}\label{relwork}

\section{Scope}\label{scope}

We try to understand the loss recovery and retransmission policies in state of the art MPTCP linux implementation.

\section{Experimental Setup}\label{exsetup}

Planned experiments use CORE emulation platform with a simple topology as depicted in Figure~\ref{fig1}.
Client is connected to two wireless interfaces 3G and WLAN. Server is connected to a wired router.
Characteristics of the connection setup and assumptions about the parameters are provided in table~\ref{tab1}.
Traffic generation is planned with iperf, netperf and d-itg as different runs.

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.9\textwidth]{images/fortest.ps}
\caption{Topology used for Emulation}\label{fig1}
\end{center}
\label{prl}
\end{figure}
\begin{center}
\begin{table}
\begin{center}
\begin{tabular}{|c|cccccccccc|}
      \hline
      \multicolumn{1}{c}{} & & \\[\dimexpr-\normalbaselineskip-\arrayrulewidth]
      \textbf{Burst Size} & \multicolumn{10}{c|}{7 Packets} \\
      \hline
      \textbf{Separation Time} & \multicolumn{10}{c|}{2s} \\
      \hline

      \textbf{Run Length} & \multicolumn{10}{c|}{10min}\\
      \hline 	
      \textbf{RTT} & \multicolumn{5}{c|}{200ms(3G)} & \multicolumn{5}{c|}{20ms(WLAN)} \\
      \hline 	
      \textbf{Bandwidth} & \multicolumn{5}{c|}{7Mbps(3G)} & \multicolumn{5}{c|}{54Mbps(WLAN)} \\
      \hline
      \textbf{Loss} & \multicolumn{5}{c|}{0(3G)} & \multicolumn{5}{c|}{2\%(WLAN)} \\
      \hline
      \textbf{Loss Model} & \multicolumn{10}{c|}{Random}\\
      \hline
\end{tabular}
\caption{Connection parameters}\label{tab1}
\end{center}
\end{table}
\end{center}

\section{Observations and Discussion}\label{disc}
Preliminary experiments on CORE emulator with the setup mentioned in section\ref{exsetup} showed several cases where the behavior on losses differ.
These cases can be categorized as 
\begin{itemize}
 \item SYN Loss
 \item First packet loss (in second subflow)
 \item RTO 
 \item Unwanted RST 
 \item Tail loss
 \item Specific packet loss - in meta level and mptcp level

\end{itemize}

xplot analysis indicates that SYN loss triggers an RTO in the worst case and incurring a latency of 3 RTTs at the sender.

\section{Tail loss scenarios}
We tried to understand the behavior of MPTCP retransmissions at different flow levels in the following tail loss scenarios~\ref{ret3}.
MPTCP is sensitive to path asymmetry in general due to the default scheduler being shortest RTT based scheduler. Wireshark and tcptrace
analysis provides information on the TCP retransmissions at the meta level. mptcptrace provides information on flow sequence split at 
subflow level. Subflow level retransmissions are difficult to test from the output of mptcptrace analysis. Further deep packet inspection
is required to understand the losses at the subflow level.

\begin{table}[]
\centering
\caption{Tail loss scenarios with tcp.early.retrans = 3 default}
\label{ret3}
\begin{tabular}{|l|l|l|l|l|}
\hline
 testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \hline
\end{tabular}
\end{table}

The linux sysctl setting tcp\_early\_retrans is 3 by default. This option enables both Early retransmit and TLP. 
Further testcases should be considered with options 0 and 4.

\bibliographystyle{IEEEtran}
\bibliography{mptcp-retrans}
\end{document}

### 2016-10-05 main.tex
initial merge of retrans and tlp

@@
 % Title Page
-\title{Understanding Loss recovery in MultiPath TCP }
+\title{Loss recovery and tail loss performance in Multipath TCP}
 \author{KAU}
@@
       \multicolumn{1}{c}{} & & \\[\dimexpr-\normalbaselineskip-\arrayrulewidth]
-      \textbf{Burst Size} & \multicolumn{10}{c|}{7 Packets} \\
+      \textbf{Burst Size} & \multicolumn{10}{c|}{160 Packets} \\
       \hline
@@
 
+\subsection{Testing with Deterministic loss patterns}
+In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
+
 \section{Observations and Discussion}\label{disc}
 Preliminary experiments on CORE emulator with the setup mentioned in section\ref{exsetup} showed several cases where the behavior on losses differ.
-These cases can be categorized as 
-\begin{itemize}
- \item SYN Loss
- \item First packet loss (in second subflow)
- \item RTO 
- \item Unwanted RST 
- \item Tail loss
- \item Specific packet loss - in meta level and mptcp level
-
-\end{itemize}
-
-xplot analysis indicates that SYN loss triggers an RTO in the worst case and incurring a latency of 3 RTTs at the sender.
 
@@
 Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
-Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \hline
+Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                &  \\ \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[]
+\centering
+\caption{Tail loss scenarios with tcp.early.retrans = 0 ER disabled}
+\label{ret3}
+\begin{tabular}{|l|l|l|l|l|}
+\hline
+ testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
+Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
+Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
+Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
+Tail drop 4 & 20ms - 120ms &  			&		&  \\ \hline
+Tail drop 5 & 120ms - 20ms &  			& 		& \\ \hline 
+\end{tabular}
+\end{table}
+
+\begin{table}[]
+\centering
+\caption{Tail loss scenarios with tcp.early.retrans = 4 TLP disabled}
+\label{ret3}
+\begin{tabular}{|l|l|l|l|l|}
+\hline
+ testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
+Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
+Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
+Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
 \end{tabular}
@@
 
+\section{Improvements to TLP}
+The goal of Tail Loss Probe(TLP) is to reduce tail latency of short flows.
+It achieves this by converting retransmission timeouts (RTOs) occuring due to tail losses (losses at end of transactions) into fast recovery.
+TLP transmits one packet in two round-trips when a connection is in Open state and isn't receiving any ACKs.
+The transmitted packet, aka loss probe, can be either new or a retransmission.
+When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout.
+
+Current implementation of TLP is at the TCP flow level.
+In the event of RTO timeout the MPTCP retransmission happens with a reinjection in to scheduler along with sending lost packet on the same path.
+But in the Event of Probe timeout, the lost packet is being sent on the same path without injecting in to scheduler.
+Modifications should reinject the lost packet in to the mptcp scheduler in the event of PTO.
+
+\subsection{Example scenario where MPTCP TLP could incur more latency}
+
+Use a traffic pattern where there is new burst to send before the RTO after a tail loss triggering TLP on a flow.
+
+\section{Proposed changes and expected implications}
+
+\subsection{Implementation Plan}
+
+Things to do:
+
+When PTO fires: in code trigger call on PTO in tcp level
+It calls ${tcp\_push\_one()}$ in tcp-output.c and then ${mptcp\_write\_xmit()}$.
+As it is reinjection (due to only one case currently in MPTCP), that packet will have path mask
+to put it on the same subflow. Here we need to add a case for reinjection due to PTO that should
+send the packet to $reinject\_queue$ of ${mptcp\_cb}$ as well.
+
+Main flow packet queue and subflow packet queue: ${meta\_sk}$ refers to the
+main mptcp flow.
+
+Loss probe timer used in Linux: ${ICSK\_TIME\_LOSS\_PROBE}$.
+
+\section{Evaluation of proposal}
+
+Initial testing with customized program to send bursts of packets with fixed burst seperation. Adjust the burst seperation to get the desired scenario.
+
 \bibliographystyle{IEEEtran}

### 2016-10-18 plots/er4.py
Update

@@ -1,0 +1,61 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+#data = np.loadtxt(sys.argv[1])
+
+n_groups = 5 
+
+s2020 = (183,481)
+
+s2030 = (185,542)
+
+s3020 = (214,532)
+
+s20120 = (185,552)
+
+s12020 = (236,887)
+
+
+fig, ax = plt.subplots()
+
+index = np.arange(n_groups)
+bar_width = 0.18
+
+opacity = 0.4
+
+rects1 = plt.bar(index, s2020 , bar_width,
+                 alpha=opacity,
+                 color='b',
+                 label='20-20')
+
+rects2 = plt.bar(index + bar_width, s2030, bar_width,
+                 alpha=opacity,
+                 color='r',
+                 label='20-30')
+
+rects3 = plt.bar(index + 2*bar_width, s3020, bar_width,
+                 alpha=opacity,
+                 color='g',
+                 label='30-20')
+
+rects4 = plt.bar(index + 3*bar_width, s20120, bar_width,
+                 alpha=opacity,
+                 color='m',
+                 label='20-120')
+
+rects5 = plt.bar(index + 4*bar_width, s12020, bar_width,
+		alpha=opacity,
+		color='c',
+		label='120-20')
+
+
+plt.xlabel('Scenarios')
+plt.ylabel('AvgDelay (ms)')
+#plt.title('AvgDelay with Std. Dev ')
+plt.xticks(0.185 + index + bar_width, ('20-20', '20-30', '30-20', '20-120', '120-20'))
+#plt.legend()
+plt.tight_layout()
+plt.legend(loc='upper center', bbox_to_anchor=(0.5, 1.05), ncol=4)
+plt.savefig('avgDelay.pdf')
+
+#plt.show()

### 2016-10-18 plots/er4.py
Update

@@ -5,15 +5,18 @@ import matplotlib.pyplot as plt
 
 n_groups = 5 
 
-s2020 = (183,481)
+b1 = (183,185,214,185,236)
+b2 = (481,542,532,552,887)
 
-s2030 = (185,542)
+#s2020 = (183,481)
 
-s3020 = (214,532)
+#s2030 = (185,542)
 
-s20120 = (185,552)
+#s3020 = (214,532)
 
-s12020 = (236,887)
+#s20120 = (185,552)
+
+#s12020 = (236,887)
 
 
 fig, ax = plt.subplots()
@@ -23,30 +26,15 @@ bar_width = 0.18
 
 opacity = 0.4
 
-rects1 = plt.bar(index, s2020 , bar_width,
+rects1 = plt.bar(index, b1 , bar_width,
                  alpha=opacity,
                  color='b',
-                 label='20-20')
+                 label='Burst1')
 
-rects2 = plt.bar(index + bar_width, s2030, bar_width,
+rects2 = plt.bar(index + bar_width, b2, bar_width,
                  alpha=opacity,
                  color='r',
-                 label='20-30')
-
-rects3 = plt.bar(index + 2*bar_width, s3020, bar_width,
-                 alpha=opacity,
-                 color='g',
-                 label='30-20')
-
-rects4 = plt.bar(index + 3*bar_width, s20120, bar_width,
-                 alpha=opacity,
-                 color='m',
-                 label='20-120')
-
-rects5 = plt.bar(index + 4*bar_width, s12020, bar_width,
-		alpha=opacity,
-		color='c',
-		label='120-20')
+                 label='Burst2')
 
 
 plt.xlabel('Scenarios')
@@ -55,7 +43,7 @@ plt.ylabel('AvgDelay (ms)')
 plt.xticks(0.185 + index + bar_width, ('20-20', '20-30', '30-20', '20-120', '120-20'))
 #plt.legend()
 plt.tight_layout()
-plt.legend(loc='upper center', bbox_to_anchor=(0.5, 1.05), ncol=4)
+plt.legend(loc='upper center', bbox_to_anchor=(0.5, 1.05), ncol=5)
 plt.savefig('avgDelay.pdf')
 
 #plt.show()

### 2016-11-21 main.tex
Update
update

@@
 \usepackage[utf8]{inputenc}
-\usepackage[dvips]{graphicx}
+\usepackage{graphicx}
+
 % Title Page
@@
 
-We try to understand the loss recovery and retransmission policies in state of the art MPTCP linux implementation.
+We try to understand the loss recovery and retransmission policies in state of the art MPTCP linux implementation. As per the specification of Multipath TCP introduced in 
+MPTCP features acknowledgements at connection-level as well as
+   subflow-level acknowledgements, in order to provide a robust service
+   to the application.
 
@@
 \begin{center}
-\includegraphics[angle=0, width=0.9\textwidth]{images/fortest.ps}
+\includegraphics[angle=0, width=0.9\textwidth]{images/fortest.pdf}
 \caption{Topology used for Emulation}\label{fig1}
 \end{center}
-\label{prl}
 \end{figure}
 \begin{center}
+
 \begin{table}
@@
 In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
+Losses generated by using associating netem with corresponding interfaces and dropping the packets(Cite KAU paper).\ref{timing}\ref{1p}\ref{1pp}\ref{2p}
 
@@
 Preliminary experiments on CORE emulator with the setup mentioned in section\ref{exsetup} showed several cases where the behavior on losses differ.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.9\textwidth, natwidth=610, natheight=400]{images/timing.pdf}
+\end{center}
+\caption{Timing diagram of different tail loss scenarios}\label{timing}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.9\textwidth,natwidth=578.16,natheight=433.62]{plots/1P.pdf}
+\caption{Single packet tail loss}\label{1p}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.9\textwidth, natwidth=578.16,natheight=433.62]{plots/1PP.pdf}
+\caption{Single packet tail loss together with probe loss}\label{1pp}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.9\textwidth,natwidth=578.16,natheight=433.62]{plots/2P.pdf}
+\caption{Two packet tail loss}\label{2p}
+\end{center}
+\end{figure}
 
@@
 
-\begin{table}[]
+\begin{table}[!ht]
 \centering
@@
 
-\begin{table}[]
+\begin{table}[!ht]
 \centering
 \caption{Tail loss scenarios with tcp.early.retrans = 0 ER disabled}
-\label{ret3}
+\label{ret0}
 \begin{tabular}{|l|l|l|l|l|}
@@
 
-\begin{table}[]
+\begin{table}[!ht]
 \centering
 \caption{Tail loss scenarios with tcp.early.retrans = 4 TLP disabled}
-\label{ret3}
+\label{ret4}
 \begin{tabular}{|l|l|l|l|l|}

### 2016-11-21 main.tex
update

@@
 behavior.
+
 \section{Related Work}\label{relwork}
-
+TBD.
 \section{Scope}\label{scope}
 
-We try to understand the loss recovery and retransmission policies in state of the art MPTCP linux implementation. As per the specification of Multipath TCP introduced in 
-MPTCP features acknowledgements at connection-level as well as
-   subflow-level acknowledgements, in order to provide a robust service
-   to the application.
+We try to understand the loss recovery and retransmission policies in state of the art MPTCP linux implementation version 0.91. As per the specification of Multipath TCP introduced in 
+MPTCP features acknowledgements at connection-level as well as subflow-level, in order to provide a robust transport service to the application.
 
@@
 
-Planned experiments use CORE emulation platform with a simple topology as depicted in Figure~\ref{fig1}.
-Client is connected to two wireless interfaces 3G and WLAN. Server is connected to a wired router.
+Experiments use CORE emulation platform with a simple topology as depicted in Figure~\ref{fig1}.
+Client is connected to two wireless interfaces 3G/4G and WLAN. Server is connected to a wired router.
 Characteristics of the connection setup and assumptions about the parameters are provided in table~\ref{tab1}.
-Traffic generation is planned with iperf, netperf and d-itg as different runs.
+There is no attempt in this study to focus on the effect of link characteristics in retransmission performance.
 
@@
 
-      \textbf{Run Length} & \multicolumn{10}{c|}{10min}\\
+      \textbf{RTT} & \multicolumn{5}{c|}{20ms-120ms(3G/4G)} & \multicolumn{5}{c|}{20ms-120ms(WLAN)} \\
       \hline 	
-      \textbf{RTT} & \multicolumn{5}{c|}{200ms(3G)} & \multicolumn{5}{c|}{20ms(WLAN)} \\
+      \textbf{Bandwidth} & \multicolumn{5}{c|}{54Mbps(3G/4G)} & \multicolumn{5}{c|}{54Mbps(WLAN)} \\
       \hline
-      \textbf{Bandwidth} & \multicolumn{5}{c|}{7Mbps(3G)} & \multicolumn{5}{c|}{54Mbps(WLAN)} \\
-      \hline
-      \textbf{Loss} & \multicolumn{5}{c|}{0(3G)} & \multicolumn{5}{c|}{2\%(WLAN)} \\
-      \hline
-      \textbf{Loss Model} & \multicolumn{10}{c|}{Random}\\
+      \textbf{Loss Model} & \multicolumn{10}{c|}{Deterministic}\\
       \hline
@@
 In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
-Losses generated by using associating netem with corresponding interfaces and dropping the packets(Cite KAU paper).\ref{timing}\ref{1p}\ref{1pp}\ref{2p}
+Losses generated by using associating netem with corresponding interfaces and dropping the packets. This process is simplified by using the KAUNetem tool~\cite{Garcia2016}. 
+MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between client and server on the MPTCP protocol 
+support and interface availability. So one has to wait in time and packets for the second subflow establishment to understand the MPTCP retransmissions. We send two bursts of 80 
+packets each and drop tail packets on one of the interface. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that
+subflow. Three test cases are evaluated with one packet tail loss, two packet tail loss and  one packet along with probe loss. The expected retransmission behavior is shown in~\ref{timing}.
 
-\section{Observations and Discussion}\label{disc}
-Preliminary experiments on CORE emulator with the setup mentioned in section\ref{exsetup} showed several cases where the behavior on losses differ.
+In Linux, TCP retransmission features are controlled by a sysctl setting tcp.early.retrans. It has 5 possible values with ranging from 0 to 4 with 3 being Linux default, enables both
+ER and TLP. For this analysis, we used 0,3 and 4. In each setting, we calculate burst completion time for two burst of 80 packets each. The setup has server and client with MPTCP
+enabled linux running on them. Client has two wireless interfaces with one way delay on each interface ranging from 20ms to 120ms. We consider 5 scenarios with delay pairings 20ms-20ms
+20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance.
 
@@
 
+\section{Observations and Discussion}\label{disc}
+
+Experiments performed on CORE emulator with the setup mentioned in section\ref{exsetup}.
+MPTCP is sensitive to path asymmetry in general due to the default scheduler being shortest RTT based scheduler. Wireshark and tcptrace
+analysis provides information on the TCP retransmissions at the meta level. mptcptrace provides information on flow sequence split at 
+subflow level. Subflow level retransmissions are difficult to test from the output of mptcptrace analysis. Further deep packet inspection
+required to understand the losses at the subflow level.
+
+Figure~\ref{1p} provides burst completion time comparison of different ER sysctl settings considered with one packet loss. The results show the improvements
+in burst completion time with ER and TLP enabled compared to both disabled only when there is delay difference. The retransmission of lost packet follows the behavior 
+depicted in Figure~\ref{timing}.
+
 \begin{figure}[!ht]
@@
 
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.9\textwidth, natwidth=578.16,natheight=433.62]{plots/1PP.pdf}
-\caption{Single packet tail loss together with probe loss}\label{1pp}
-\end{center}
-\end{figure}
+In packet trace inspection, it is easy to check the use of TLP with two packet tail loss. The last packet will be retransmitted first in this case.
+With one packet tail loss one has to compare the RTT and last packet retransmission time to check the use of TLP.
+To further validate the results with one packet tail loss, we consider two packet tail loss and the results are shown in~\ref{2p}.
+In this case, ER=3 setting performs better than the other two settings in all scenarios except when the packets lost on high delay interface.
 
@@
 
-\section{Tail loss scenarios}
-We tried to understand the behavior of MPTCP retransmissions at different flow levels in the following tail loss scenarios~\ref{ret3}.
-MPTCP is sensitive to path asymmetry in general due to the default scheduler being shortest RTT based scheduler. Wireshark and tcptrace
-analysis provides information on the TCP retransmissions at the meta level. mptcptrace provides information on flow sequence split at 
-subflow level. Subflow level retransmissions are difficult to test from the output of mptcptrace analysis. Further deep packet inspection
-is required to understand the losses at the subflow level.
+To further study the performance of TLP in improving latency, we tried to drop the probe packet along with the last packet. In this case ER=0
+setting results much lower burst completion times.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.9\textwidth, natwidth=578.16,natheight=433.62]{plots/1PP.pdf}
+\caption{Single packet tail loss together with probe loss}\label{1pp}
+\end{center}
+\end{figure}
 
@@
 
-The linux sysctl setting tcp\_early\_retrans is 3 by default. This option enables both Early retransmit and TLP. 
-Further testcases should be considered with options 0 and 4.
-
 \section{Improvements to TLP}
@@
 When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout.
+Our results show that current TLP implementation does not improve the performance in MPTCP and incur more latency.
 

### 2016-11-21 plots/plot_tlp.py

@@ -25,21 +25,21 @@ margin = 0.05
 bar_width = (1-2*margin)/3
 
 
-rects1 = plt.bar(index+margin, p1pe0 , bar_width,
+rects1 = plt.bar(index+margin, p2e0 , bar_width,
                  color='b',
                  label='er=0')
 
-rects2 = plt.bar(index+margin + bar_width, p1pe3, bar_width,
+rects2 = plt.bar(index+margin + bar_width, p2e3, bar_width,
                  color='r',
                  label='er=3')
 
-rects3 = plt.bar(index+margin + 2*bar_width, p1pe4, bar_width,
+rects3 = plt.bar(index+margin + 2*bar_width, p2e4, bar_width,
                  color='g',
                  label='er=4')
 
 
 plt.xlabel('Scenarios')
-plt.ylabel('AvgDelay (ms)')
+plt.ylabel('Burst Completion Time (ms)')
 #plt.title('AvgDelay with Std. Dev ')
 plt.xticks(index+0.5,('20-20', '20-30', '30-20', '20-120', '120-20'))
 #plt.legend()

### 2016-12-04 main.tex
Fig Changes

@@
 packets each and drop tail packets on one of the interface. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that
-subflow. Three test cases are evaluated with one packet tail loss, two packet tail loss and  one packet along with probe loss. The expected retransmission behavior is shown in~\ref{timing}.
+subflow. Three test cases are evaluated with one packet tail loss, two packet tail loss and  one packet along with probe loss.
 
@@
 enabled linux running on them. Client has two wireless interfaces with one way delay on each interface ranging from 20ms to 120ms. We consider 5 scenarios with delay pairings 20ms-20ms
-20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance.
+20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance. The expected retransmission behavior for different settings of ER is shown 
+in~\ref{timingER0},~/\ref{timingER3} and ~\ref{timingER4}.
 
@@
 \begin{center}
-\includegraphics[angle=0, width=0.9\textwidth, natwidth=610, natheight=400]{images/timing.pdf}
+\includegraphics[angle=0, width=0.9\textwidth, natwidth=610, natheight=400]{images/timingER0.pdf}
 \end{center}
-\caption{Timing diagram of different tail loss scenarios}\label{timing}
+\caption{Timing diagram of different tail loss scenarios in sysctl ER=0}\label{timingER0}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.9\textwidth, natwidth=610, natheight=400]{images/timingER3.pdf}
+\end{center}
+\caption{Timing diagram of different tail loss scenarios in sysctl ER=3}\label{timingER3}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.9\textwidth, natwidth=610, natheight=400]{images/timingER4.pdf}
+\end{center}
+\caption{Timing diagram of different tail loss scenarios in sysctl ER=4}\label{timingER4}
 \end{figure}
@@
 in burst completion time with ER and TLP enabled compared to both disabled only when there is delay difference. The retransmission of lost packet follows the behavior 
-depicted in Figure~\ref{timing}.
+depicted in Figures~\ref{timingER0,timintER3,timingER4}.
 

### 2016-12-10 main.tex
TCP comparison

@@
 
+Figures ~\ref{t1p,t2p,t1pp} represent the comparison of MPTCP and TCP burst completion times for the two bursts from server to client.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.9\textwidth,natwidth=578.16,natheight=433.62]{plots/T1P.pdf}
+\caption{Single packet tail loss}\label{t1p}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.9\textwidth,natwidth=578.16,natheight=433.62]{plots/T2P.pdf}
+\caption{Two packet tail loss}\label{t2p}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.9\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
+\caption{Single packet tail loss together with probe loss}\label{t1pp}
+\end{center}
+\end{figure}
+
 \begin{table}[!ht]

### 2016-12-10 plots/fetchDelay.sh

@@ -1,9 +1,9 @@
 #!/bin/bash
 
-FILES1=/home/vkirany/mptcptests/
-for p in 1P 2P 1P+P
+FILES1=/home/vkirany/mptcptests/tcp/
+for e in 'er0' 'er2' 'er3' 'er4'
 do
-for e in 'er0' 'er3' 'er4'
+for p in 1P 2P 1P+P
 do
 for x in 'tail20-20' 'tail20-30' 'tail30-20' 'tail20-120' 'tail120-20'
 do

### 2016-12-10 plots/plot_tlp_c.py

@@ -43,7 +43,7 @@ fig, ax = plt.subplots()
 index = np.arange(n_groups)
 
 margin = 0.05
-bar_width = (1-2*margin)/4
+bar_width = (1-2*margin)/8
 
 
 rects1 = plt.bar(index+margin, p1pe0 , bar_width,
@@ -62,6 +62,19 @@ rects4 = plt.bar(index+margin + 3*bar_width, p1pe4, bar_width,
                  color='g',
                  label='er=4')
 
+rects5 = plt.bar(index+margin + 4*bar_width, tp1pe0, bar_width,
+                 color='c',
+                 label='ter=0')
+rects6 = plt.bar(index+margin + 5*bar_width, tp1pe2, bar_width,
+                 color='y',
+                 label='ter=2')
+rects7 = plt.bar(index+margin + 6*bar_width, tp1pe3, bar_width,
+                 color='k',
+                 label='ter=3')
+rects8 = plt.bar(index+margin + 7*bar_width, tp1pe4, bar_width,
+                 color='#B47CC7',
+                 label='ter=4')
+
 
 plt.xlabel('Scenarios')
 plt.ylabel('Burst Completion Time (ms)')
@@ -69,8 +82,8 @@ plt.ylabel('Burst Completion Time (ms)')
 plt.xticks(index+0.5,('20-20', '20-30', '30-20', '20-120', '120-20'))
 #plt.legend()
 plt.tight_layout()
-plt.legend(loc='upper center', bbox_to_anchor=(0.5, 1.05), ncol=5)
-#plt.ylim(0,650)
+plt.ylim(0,1850)
+plt.legend(loc='upper center', bbox_to_anchor=(0.5, 1.05), ncol=4)
 plt.savefig('Delay.pdf')
 
 #plt.show()

### 2016-12-18 main.tex
Update

--- 2016-12-04 main.tex
@@
-\documentclass[12pt,draftcls,onecolumn]{IEEEtran}
+\documentclass[10pt,draftcls,twocolumn]{IEEEconf}
 \usepackage[utf8]{inputenc}
@@
 \begin{center}
-\includegraphics[angle=0, width=0.9\textwidth]{images/fortest.pdf}
+\includegraphics[angle=0, width=0.45\textwidth]{images/fortest.pdf}
 \caption{Topology used for Emulation}\label{fig1}
@@
 enabled linux running on them. Client has two wireless interfaces with one way delay on each interface ranging from 20ms to 120ms. We consider 5 scenarios with delay pairings 20ms-20ms
-20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance. The expected retransmission behavior for different settings of ER is shown 
+20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance.
+
+Figures ~\ref{t1p,t2p,t1pp} represent the comparison of TCP burst completion times for the two bursts from server to client.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.46\textwidth,natwidth=578.16,natheight=433.62]{plots/T1P.pdf}
+\caption{Single packet tail loss using TCP}\label{t1p}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.46\textwidth,natwidth=578.16,natheight=433.62]{plots/T2P.pdf}
+\caption{Two packet tail loss using TCP}\label{t2p}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
+\caption{Single packet tail loss together with probe loss using TCP}\label{t1pp}
+\end{center}
+\end{figure}
+
+The expected retransmission behavior for different settings of ER is shown 
 in~\ref{timingER0},~/\ref{timingER3} and ~\ref{timingER4}.
@@
 \begin{center}
-\includegraphics[angle=0, width=0.9\textwidth, natwidth=610, natheight=400]{images/timingER0.pdf}
+\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timingER0.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[angle=0, width=0.9\textwidth, natwidth=610, natheight=400]{images/timingER3.pdf}
+\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timingER3.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[angle=0, width=0.9\textwidth, natwidth=610, natheight=400]{images/timingER4.pdf}
+\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timingER4.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[angle=0, width=0.9\textwidth,natwidth=578.16,natheight=433.62]{plots/1P.pdf}
-\caption{Single packet tail loss}\label{1p}
+\includegraphics[angle=0, width=0.46\textwidth,natwidth=578.16,natheight=433.62]{plots/1P.pdf}
+\caption{Single packet tail loss using MPTCP}\label{1p}
 \end{center}
@@
 \begin{center}
-\includegraphics[angle=0, width=0.9\textwidth,natwidth=578.16,natheight=433.62]{plots/2P.pdf}
-\caption{Two packet tail loss}\label{2p}
+\includegraphics[angle=0, width=0.46\textwidth,natwidth=578.16,natheight=433.62]{plots/2P.pdf}
+\caption{Two packet tail loss using MPTCP}\label{2p}
 \end{center}
@@
 \begin{center}
-\includegraphics[angle=0, width=0.9\textwidth, natwidth=578.16,natheight=433.62]{plots/1PP.pdf}
-\caption{Single packet tail loss together with probe loss}\label{1pp}
+\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/1PP.pdf}
+\caption{Single packet tail loss together with probe loss using MPTCP}\label{1pp}
 \end{center}

### 2016-12-18 plots/plot_tlp.py

@@ -43,24 +43,24 @@ fig, ax = plt.subplots()
 index = np.arange(n_groups)
 
 margin = 0.05
-bar_width = (1-2*margin)/4
+bar_width = (1-2*margin)/3
 
 
-rects1 = plt.bar(index+margin, p1pe0 , bar_width,
+rects1 = plt.bar(index+margin, p2e0 , bar_width,
                  color='b',
                  label='er=0')
 
-rects2 = plt.bar(index+margin + bar_width, p1pe2 , bar_width,
-                 color='m',
+rects2 = plt.bar(index+margin + bar_width, p2e2 , bar_width,
+                 color='g',
                  label='er=2')
 
-rects3 = plt.bar(index+margin + 2*bar_width, p1pe3, bar_width,
+rects3 = plt.bar(index+margin + 2*bar_width, p2e3, bar_width,
                  color='r',
                  label='er=3')
 
-rects4 = plt.bar(index+margin + 3*bar_width, p1pe4, bar_width,
-                 color='g',
-                 label='er=4')
+#rects4 = plt.bar(index+margin + 3*bar_width, p1pe4, bar_width,
+#                 color='g',
+#                 label='er=4')
 
 
 plt.xlabel('Scenarios')

### 2016-12-18 plots/plot_tlp_c.py

@@ -43,37 +43,37 @@ fig, ax = plt.subplots()
 index = np.arange(n_groups)
 
 margin = 0.05
-bar_width = (1-2*margin)/8
+bar_width = (1-2*margin)/3
 
 
-rects1 = plt.bar(index+margin, p1pe0 , bar_width,
+rects1 = plt.bar(index+margin, tp2e0 , bar_width,
                  color='b',
                  label='er=0')
 
-rects2 = plt.bar(index+margin + bar_width, p1pe2 , bar_width,
-                 color='m',
+rects2 = plt.bar(index+margin + bar_width, tp2e2 , bar_width,
+                 color='g',
                  label='er=2')
 
-rects3 = plt.bar(index+margin + 2*bar_width, p1pe3, bar_width,
+rects3 = plt.bar(index+margin + 2*bar_width, tp2e3, bar_width,
                  color='r',
                  label='er=3')
 
-rects4 = plt.bar(index+margin + 3*bar_width, p1pe4, bar_width,
-                 color='g',
-                 label='er=4')
+#rects4 = plt.bar(index+margin + 3*bar_width, p1pe4, bar_width,
+#                 color='g',
+#                 label='er=4')
 
-rects5 = plt.bar(index+margin + 4*bar_width, tp1pe0, bar_width,
-                 color='c',
-                 label='ter=0')
-rects6 = plt.bar(index+margin + 5*bar_width, tp1pe2, bar_width,
-                 color='y',
-                 label='ter=2')
-rects7 = plt.bar(index+margin + 6*bar_width, tp1pe3, bar_width,
-                 color='k',
-                 label='ter=3')
-rects8 = plt.bar(index+margin + 7*bar_width, tp1pe4, bar_width,
-                 color='#B47CC7',
-                 label='ter=4')
+#rects5 = plt.bar(index+margin + 4*bar_width, tp1pe0, bar_width,
+#                 color='c',
+#                 label='ter=0')
+#rects6 = plt.bar(index+margin + 5*bar_width, tp1pe2, bar_width,
+#                 color='y',
+#                 label='ter=2')
+#rects7 = plt.bar(index+margin + 6*bar_width, tp1pe3, bar_width,
+#                 color='k',
+#                 label='ter=3')
+#rects8 = plt.bar(index+margin + 7*bar_width, tp1pe4, bar_width,
[…]

### 2016-12-23 main.tex
edit

@@
 On the other hand shorter flows which are a majority in the web traffic have to depend on RTOs for packet loss detection
-and recovery incurring more latency.
+and recovery thus incurring more latency.
 A single packet loss in a short flow may take many RTTs to detect and recover.
@@
       \multicolumn{1}{c}{} & & \\[\dimexpr-\normalbaselineskip-\arrayrulewidth]
-      \textbf{Burst Size} & \multicolumn{10}{c|}{160 Packets} \\
+      \textbf{Burst Size} & \multicolumn{10}{c|}{80 Packets} \\
       \hline
@@
 
-\subsection{Testing with Deterministic loss patterns}
+\subsection{Testing retransmission with deterministic loss patterns}
 In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
@@
 In Linux, TCP retransmission features are controlled by a sysctl setting tcp.early.retrans. It has 5 possible values with ranging from 0 to 4 with 3 being Linux default, enables both
-ER and TLP. For this analysis, we used 0,3 and 4. In each setting, we calculate burst completion time for two burst of 80 packets each. The setup has server and client with MPTCP
-enabled linux running on them. Client has two wireless interfaces with one way delay on each interface ranging from 20ms to 120ms. We consider 5 scenarios with delay pairings 20ms-20ms
-20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance.
+ER and TLP. For this analysis, apart from the default setting, we also consider 0 that diasables both ER and TLP and 2 that enables delayed ER. Disabling ER and TLP gives us the performance
+of RTO efficiency. Enabling delayed ER gives us combined performance of RTO and delayed ER and provides the possibility to single out the scenarios where TLP is more effective.
+Probe loss should be seen as retransmission loss when tcp.early.retrans is 0 or 2.
 
-Figures ~\ref{t1p,t2p,t1pp} represent the comparison of TCP burst completion times for the two bursts from server to client.
+In each setting, we calculate burst completion time for two bursts of 80 packets each. The setup has server and client with Linux supporting MPTCP running on them. Client has two wireless 
+interfaces with one way delay on each interface ranging from 20ms to 120ms. We consider 5 scenarios with delay pairings 20ms-20ms, 20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand 
+the effect of delay difference in the performance. These scenarios are choosen to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in link delays.
+
+Retransmission behavior of TCP is seen a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
+starts with the results using TCP and compares it with that of the MPTCP. Figures ~\ref{t1p},~\ref{t2p},~\ref{t1pp} represent the comparison of TCP 
+burst completion times.
 
@@
 
-The expected retransmission behavior for different settings of ER is shown 
-in~\ref{timingER0},~/\ref{timingER3} and ~\ref{timingER4}.
+The expected retransmission behavior of MPTCP for different settings of ER is shown 
+in~\ref{timingER0},~\ref{timingER2} and ~\ref{timingER3}.
 
@@
 \begin{center}
-\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timingER3.pdf}
+\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timingER2.pdf}
 \end{center}
-\caption{Timing diagram of different tail loss scenarios in sysctl ER=3}\label{timingER3}
+\caption{Timing diagram of different tail loss scenarios in sysctl ER=2}\label{timingER2}
 \end{figure}
@@
 \begin{center}
-\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timingER4.pdf}
+\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timingER3.pdf}
 \end{center}
-\caption{Timing diagram of different tail loss scenarios in sysctl ER=4}\label{timingER4}
+\caption{Timing diagram of different tail loss scenarios in sysctl ER=3}\label{timingER3}
 \end{figure}

### 2016-12-23 main.tex
Edit

@@
 
+\section{Observations and Discussion}\label{disc}
+
 Retransmission behavior of TCP is seen a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
 starts with the results using TCP and compares it with that of the MPTCP. Figures ~\ref{t1p},~\ref{t2p},~\ref{t1pp} represent the comparison of TCP 
-burst completion times.
+burst completion times. Path asymmetry is irrelevant as the TCP uses single path.
+
+In the case of single packet tail loss, there is no difference in burst completion times of ER setting 0,2 and 3. All the cases depend on the RTO
+to recover from loss. In two packet tail loss case, ER setting 0 and 2 have similar performance while ER setting 3 improves burst completion time
+by using TLP. If the probe or retransmission is lost, then ER 3 still performs better than 0 and 2 as it waits for one RTO and one probe timeout
+instead of two RTOs in other two cases.
 
@@
 
-\section{Observations and Discussion}\label{disc}
-
 Experiments performed on CORE emulator with the setup mentioned in section\ref{exsetup}.
 MPTCP is sensitive to path asymmetry in general due to the default scheduler being shortest RTT based scheduler. Wireshark and tcptrace
-analysis provides information on the TCP retransmissions at the meta level. mptcptrace provides information on flow sequence split at 
-subflow level. Subflow level retransmissions are difficult to test from the output of mptcptrace analysis. Further deep packet inspection
-required to understand the losses at the subflow level.
+analysis provides information on the TCP retransmissions at the subflow level.
 
 Figure~\ref{1p} provides burst completion time comparison of different ER sysctl settings considered with one packet loss. The results show the improvements
-in burst completion time with ER and TLP enabled compared to both disabled only when there is delay difference. The retransmission of lost packet follows the behavior 
-depicted in Figures~\ref{timingER0,timintER3,timingER4}.
+in burst completion time with ER and TLP enabled compared to both disabled only when there is delay difference.
 
@@
 
-In packet trace inspection, it is easy to check the use of TLP with two packet tail loss. The last packet will be retransmitted first in this case.
+In packet trace inspection, it is easy to observe the use of TLP with two packet tail loss. The last packet will be retransmitted first in this case.
 With one packet tail loss one has to compare the RTT and last packet retransmission time to check the use of TLP.
@@
 To further study the performance of TLP in improving latency, we tried to drop the probe packet along with the last packet. In this case ER=0
-setting results much lower burst completion times.
+setting results much lower burst completion times. Use of TLP did not improve the burst completion time as we saw in TCP for ER=3. The reason
+lies in the way the retransmission and loss probe transmission carried out in MPTCP. Retransmission occurs after an RTO time on both paths.
+Even if retransmission fails on one path, the client receives the packet on the other path as shown in Fig~\ref{timingER0}. For TLP, in
+current implementation, the loss probe sent on the actual path of first transmission but not on both paths. This incorrect impartation of
+TLP from TCP to MPTCP led to the increase burst completion time by an RTO for MPTCP.
 
@@
 
-\begin{table}[!ht]
-\centering
-\caption{Tail loss scenarios with tcp.early.retrans = 3 default}
-\label{ret3}
-\begin{tabular}{|l|l|l|l|l|}
-\hline
- testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
-Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
-Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
-Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                &  \\ \hline
-\end{tabular}
-\end{table}
+%\begin{table}[!ht]
+%\centering
+%\caption{Tail loss scenarios with tcp.early.retrans = 3 default}
+%\label{ret3}
+%\begin{tabular}{|l|l|l|l|l|}
+%\hline
+% testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
+%Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
+%Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
+%Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                &  \\ \hline
+%\end{tabular}
+%\end{table}
 
-\begin{table}[!ht]
-\centering
-\caption{Tail loss scenarios with tcp.early.retrans = 0 ER disabled}
-\label{ret0}
-\begin{tabular}{|l|l|l|l|l|}
-\hline
- testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
-Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
-Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
-Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
-Tail drop 4 & 20ms - 120ms &  			&		&  \\ \hline
-Tail drop 5 & 120ms - 20ms &  			& 		& \\ \hline 
-\end{tabular}
-\end{table}
+%\begin{table}[!ht]
+%\centering
+%\caption{Tail loss scenarios with tcp.early.retrans = 0 ER disabled}
+%\label{ret0}
+%\begin{tabular}{|l|l|l|l|l|}
+%\hline
+% testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
+%Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
+%Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
+%Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
+%Tail drop 4 & 20ms - 120ms &  			&		&  \\ \hline
+%Tail drop 5 & 120ms - 20ms &  			& 		& \\ \hline 
+%\end{tabular}
+%\end{table}
 
-\begin{table}[!ht]
-\centering
-\caption{Tail loss scenarios with tcp.early.retrans = 4 TLP disabled}
-\label{ret4}
-\begin{tabular}{|l|l|l|l|l|}
-\hline
- testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
-Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
-Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
-Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
-\end{tabular}
-\end{table}
+%\begin{table}[!ht]
+%\centering
+%\caption{Tail loss scenarios with tcp.early.retrans = 4 TLP disabled}
+%\label{ret4}
+%\begin{tabular}{|l|l|l|l|l|}
+%\hline
+% testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
+%Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
+%Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
+%Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
+%\end{tabular}
+%\end{table}
 
-\section{Improvements to TLP}
+\section{Improvments to TLP for MPTCP}
 The goal of Tail Loss Probe(TLP) is to reduce tail latency of short flows.
@@
 In the event of RTO timeout the MPTCP retransmission happens with a reinjection in to scheduler along with sending lost packet on the same path.
-But in the Event of Probe timeout, the lost packet is being sent on the same path without injecting in to scheduler.
-Modifications should reinject the lost packet in to the mptcp scheduler in the event of PTO.
+But in the Event of Probe timeout, the loss probe packet is being sent on the same path without injecting in to scheduler.
+Modifications should reinject the probe packet in to the mptcp scheduler in the event of PTO.
 

### 2017-01-07 main.tex
Update

@@
 Initial testing with customized program to send bursts of packets with fixed burst seperation. Adjust the burst seperation to get the desired scenario.
+Results comparing the MPTCP-TLP for the considered cases are provided in figures~\ref{1pn},~\ref{2pn} and ~\ref{1ppn}. The improvement in latency
+performance is seen Single packet tail loss with probe loss scenario. Single packet loss and Two packet loss scenarios provide increase in latency.
+Analysis required in these two cases.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/1PNew.pdf}
+\caption{Single packet tail loss comparision with MPTCP-TLP}\label{1pn}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/2PNew.pdf}
+\caption{Two packet tail loss comparison with MPTCP-TLP}\label{2pn}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/1PPNew.pdf}
+\caption{Single packet tail loss together with probe loss with MPTCP-TLP}\label{1ppn}
+\end{center}
+\end{figure}
 

### 2017-01-07 plots/plot_tlp_new.py

@@ -1,0 +1,55 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+#data = np.loadtxt(sys.argv[1])
+
+n_groups = 5 
+
+p1e3=(433,482,463,443,592)
+p2e3=(301,352,355,362,676)
+p1pe3=(646,742,668,698,957)
+
+
+np1e3=(423,390,453,389,581)
+np2e3=(303,407,407,403,674)
+np1pe3=(423,483,452,443,581)
+
+tp1e3=(491,490,541,491,1003)
+tp2e3=(462,460,533,462,1187)
+tp1pe3=(748,748,807,747,1375)
+
+
+
+fig, ax = plt.subplots()
+
+index = np.arange(n_groups)
+
+margin = 0.05
+bar_width = (1-2*margin)/3
+
+
+rects1 = plt.bar(index+margin, p1pe3 , bar_width,
+                 color='b',
+                 label='mptcp')
+
+rects2 = plt.bar(index+margin + bar_width, np1pe3 , bar_width,
+                 color='g',
+                 label='mptcp-New')
+
+rects3 = plt.bar(index+margin + 2*bar_width, tp1pe3, bar_width,
+                 color='r',
+                 label='tcp')
+
+
+
+plt.xlabel('Scenarios')
+plt.ylabel('Burst Completion Time (ms)')
+#plt.title('AvgDelay with Std. Dev ')
+plt.xticks(index+0.5,('20-20', '20-30', '30-20', '20-120', '120-20'))
+#plt.legend()
+plt.tight_layout()
+plt.legend(loc='upper center', bbox_to_anchor=(0.5, 1.05), ncol=5)
+#plt.ylim(0,650)
+plt.savefig('Delay.pdf')
+
+#plt.show()

### 2017-01-15 main.tex
Update

@@
 
-Through this set of preliminary experiments, we would like to understand the behavior of MPTCP with and without TLP and ER by monitoring burst completion time, congestion window evolution and timeout 
-behavior.
-
 \section{Related Work}\label{relwork}
@@
 
-We try to understand the loss recovery and retransmission policies in state of the art MPTCP linux implementation version 0.91. As per the specification of Multipath TCP introduced in 
-MPTCP features acknowledgements at connection-level as well as subflow-level, in order to provide a robust transport service to the application.
+This paper presents the state of the art loss recovery mechanism of MPTCP and presents case for improvement in the tail loss scenario. For understanding the loss recovery and retransmission policies 
+in state of the art MPTCP linux implementation version 0.91, we consider specific cases of tail losses.
 
@@
 
+\subsection{TCP}
 Retransmission behavior of TCP is seen a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
@@
 
+\subsection{MPTCP}
+
 The expected retransmission behavior of MPTCP for different settings of ER is shown 
-in~\ref{timingER0},~\ref{timingER2} and ~\ref{timingER3}.
+in~\ref{timing1P},~\ref{timing1PP} and ~\ref{timing2P}.
 
@@
 \begin{center}
-\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timingER0.pdf}
+\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timing1P.pdf}
 \end{center}
-\caption{Timing diagram of different tail loss scenarios in sysctl ER=0}\label{timingER0}
+\caption{Timing diagram of MPTCP behavior with one packet tail loss a) STD b) LIN}\label{timing1P}
 \end{figure}
@@
 \begin{center}
-\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timingER2.pdf}
+\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timing1PP.pdf}
 \end{center}
-\caption{Timing diagram of different tail loss scenarios in sysctl ER=2}\label{timingER2}
+\caption{Timing diagram of MPTCP behavior with one packet and next probe loss a) STD b) LIN}\label{timing1PP}
 \end{figure}
@@
 \begin{center}
-\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timingER3.pdf}
+\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timing2P.pdf}
 \end{center}
-\caption{Timing diagram of different tail loss scenarios in sysctl ER=3}\label{timingER3}
+\caption{Timing diagram of MPTCP behavior with two packet tail loss a) STD b) LIN}\label{timing2P}
 \end{figure}
@@
 lies in the way the retransmission and loss probe transmission carried out in MPTCP. Retransmission occurs after an RTO time on both paths.
-Even if retransmission fails on one path, the client receives the packet on the other path as shown in Fig~\ref{timingER0}. For TLP, in
+Even if retransmission fails on one path, the client receives the packet on the other path as shown in Fig~\ref{timing1P}. For TLP, in
 current implementation, the loss probe sent on the actual path of first transmission but not on both paths. This incorrect impartation of
@@
 
-\section{Proposed changes and expected implications}
-
-\subsection{Implementation Plan}
-
-Things to do:
-
-When PTO fires: in code trigger call on PTO in tcp level
-It calls ${tcp\_push\_one()}$ in tcp-output.c and then ${mptcp\_write\_xmit()}$.
-As it is reinjection (due to only one case currently in MPTCP), that packet will have path mask
-to put it on the same subflow. Here we need to add a case for reinjection due to PTO that should
-send the packet to $reinject\_queue$ of ${mptcp\_cb}$ as well.
-
-Main flow packet queue and subflow packet queue: ${meta\_sk}$ refers to the
-main mptcp flow.
-
-Loss probe timer used in Linux: ${ICSK\_TIME\_LOSS\_PROBE}$.
-
 \section{Evaluation of proposal}

### 2017-02-27 main.tex
Update

@@
 % Title Page
-\title{Loss recovery and tail loss performance in Multipath TCP}
+\title{Loss recovery with tail loss probe in Multipath TCP}
 \author{KAU}
@@
 \begin{abstract}
+Packet losses are a well known cause of performance hindrance in the Internet. Transport protocols recover from the packet losses in order to provide reliable end to end communication. The efficiency
+of loss recovery greatly influences the completion time of the flows.
+In this paper we focus on the state of the art los recovery mechanisms for TCP and Multipath TCP.
+We use controlled tail loss scenarios
+to calculate the burst completion time. We propose an enhancement to the tail loss recovery in Multipath TCP. Our experiment results show performance improvement in controlled scenarios. 
 \end{abstract}
@@
 \section{Introduction}
-TCP has two mechanisms for detecting and recovering from packet losses namely Fast Recovery (FR) and Retransmission Timeouts(RTOs). For long flows where there are sufficiently large number of packets 
-to be transferred, FR is quicker than RTOs in detecting and recovering packet losses.
-On the other hand shorter flows which are a majority in the web traffic have to depend on RTOs for packet loss detection
-and recovery thus incurring more latency.
+
+Multipath TCP is an experimental standard proposed as an extension to TCP~\cite{rfc6824}. It allows devices with multiple interfaces to transmit simultaneously on both interfaces.  
+In Multipath TCP(MPTCP), a connection can have multiple TCP sub flows using different interfaces on different routes. Each TCP subflow imparts the functional behavior of a standard TCP flow. 
+MPTCP uses a connection level congestion window as well as sub flow level congestion windows for each sub flow.
+Packet losses in each sub flow are detected and recovered in a similar fashion as that of TCP.
+However, It is not restricted in the standard on how the loss recovery happens in the implementation and which sub-flow retransmits the lost packets.
+If the recovery is handled at the meta level, the lost packet may be rescheduled and retransmitted at the available sub flow with lowest RTT.
+If the recovery is handled at the flow level, the packet may be retransmitted in the same sub-flow.
+An important part of this recovery process is to follow the TCP semantics.
+
+TCP has two mechanisms for detecting and recovering from packet losses namely Fast Recovery (FR) and Retransmission Timeouts(RTOs).
+Fast recovery is triggered on receipt of predecided number of duplicate acknowledgements, considering it as a loss indication.
+RTOs wait for the packet loss to trigger a timeout for retransmitting that packet.
+For long flows where there are sufficiently large number of packets to be transferred, FR is quicker than RTOs in detecting and recovering packet losses.
+On the other hand shorter flows which are a majority in the web traffic have to depend on RTOs for packet loss detection and recovery thus incurring more latency.
 A single packet loss in a short flow may take many RTTs to detect and recover.
-This scenario is also applicable to the packets at the end of the flow (or tail) in a long
-flow.
+This scenario is also applicable to the packets at the end of the flow (or tail) in a long flow.
 
-TCP Loss Probe (TLP)~\cite{ietftlp} a mechanism that allows flows to detect and recover from tail losses much faster than an RTO, thereby speeding up short transfers.
+TCP Loss Probe (TLP)~\cite{Flach:2013} a mechanism that allows flows to detect and recover from tail losses much faster than an RTO, thereby speeding up short transfers.
 With TLP a packet loss in the middle
@@
 
-In Multi-path-TCP(MPTCP)~\cite{rfc6824}, a connection can have multiple TCP sub flows using different interfaces on different routes.
-Packet losses in each sub flow are assumed to be detected and recovered
-in a similar fashion as that of TCP.
-However, It is not completely clear how the loss recovery happens in the implementation and which sub-flow retransmits the lost packets.
-If the recovery is handled at the
-meta level, the lost packet may be rescheduled and retransmitted at the available sub flow with lowest RTT.
-If the recovery is handled at the flow level, the packet may be retransmitted in the same sub-flow.
+Retransmission schemes of TCP are of significant importance in achieving lower flow completion times. TLP is the recent scheme proposed and implemented in Linux kernel. Further 
+evaluationsin~\cite{Rajiullah:2015}, showed that TLP provides significant reductions in latency for short flows with one to n-degree tail losses.
+In this paper we provide the related research on retransmission in section~\ref{relwork} leading to the experimental setup.
 
-\section{Related Work}\label{relwork}
-TBD.
+\section{Background and Related Work}\label{relwork}
+
+Multipath TCP uses two levels of sequence numbers to support efficient data transfer, namely data sequence and TCP sequence.
+Data sequence is for the end to end data transfer and TCP sequence number is for individual flow data sequence similar to TCP.
+A loss occurring in an individual TCP flow corresponds to a loss in end to end data.
+Multipath TCP, as an extension of TCP, uses the basics of TCP retransmission.
+However, the interaction with two levels makes the problem of retransmission more challenging than in TCP.
+The dual sequence numbering enables MPTCP to respond to loss of packets or loss of subflows by retransmitting the lost packets on an alternate path.
+It allows a packet to be retransmitted on a different subflow than that of the originally sent flow.
+The Linux implementation of MPTCP uses a set of retransmission heuristics to decide on the retransmissions.
+The data outstanding on timed out subflows should be rescheduled for transmission on different subflows using first timeout as the indicator.
+Fast retransmit on a subflow will not trigger retransmission on another subflow.
+It is often the case in MPTCP connection that there is path asymmetry among the subflow characteristics.
+Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmtting on the fast paths.
+
 \section{Scope}\label{scope}
 
-This paper presents the state of the art loss recovery mechanism of MPTCP and presents case for improvement in the tail loss scenario. For understanding the loss recovery and retransmission policies 
-in state of the art MPTCP linux implementation version 0.91, we consider specific cases of tail losses.
+This paper analyzes the performance of state of the art loss recovery mechanism of MPTCP and presents case for improvement in the tail loss scenario. 
+For understanding the loss recovery and retransmission policies in MPTCP linux implementation version 0.91, we consider specific cases of tail losses.  
+No tuning of TCP settings performed such as adjusting the delayed ACK behavior or time and flip on Nagle algorithm. We consider the standard settings
+of Linux kernel. In the following experiments, we do not consider the congestion control interaction with the retransmission. We consider mainly failure
+induced packet loss and its recovery.
 
@@
 \begin{table}
+\resizebox{0.45\textwidth}{!}{\begin{minipage}{0.45\textwidth}
+\small
 \begin{center}
@@
       \hline
+
       \multicolumn{1}{c}{} & & \\[\dimexpr-\normalbaselineskip-\arrayrulewidth]
@@
 
-      \textbf{RTT} & \multicolumn{5}{c|}{20ms-120ms(3G/4G)} & \multicolumn{5}{c|}{20ms-120ms(WLAN)} \\
+      \textbf{RTT} & \multicolumn{10}{c|}{20ms-120ms(3G/4G/WLAN)}  \\
       \hline 	
-      \textbf{Bandwidth} & \multicolumn{5}{c|}{54Mbps(3G/4G)} & \multicolumn{5}{c|}{54Mbps(WLAN)} \\
+      \textbf{Bandwidth} & \multicolumn{10}{c|}{54Mbps(3G/4G/WLAN)}  \\
       \hline
@@
       \hline
+
 \end{tabular}
@@
 \end{center}
+\end{minipage}}
 \end{table}
@@
 In Linux, TCP retransmission features are controlled by a sysctl setting tcp.early.retrans. It has 5 possible values with ranging from 0 to 4 with 3 being Linux default, enables both
-ER and TLP. For this analysis, apart from the default setting, we also consider 0 that diasables both ER and TLP and 2 that enables delayed ER. Disabling ER and TLP gives us the performance
-of RTO efficiency. Enabling delayed ER gives us combined performance of RTO and delayed ER and provides the possibility to single out the scenarios where TLP is more effective.
-Probe loss should be seen as retransmission loss when tcp.early.retrans is 0 or 2.
+ER and TLP. Standard TCP used by other operating systems might not use ER and TLP. We use RTO to refer sysctl setting of 0 and TLP to refer sysctl setting of 3 throughout the paper.
+In standard TCP (RTO), disabling ER and TLP gives the performance of RTO. In Linux default TCP, enabling delayed ER gives the combined performance of RTO, delayed ER and TLP.
+In standard TCP, probe loss scenario is equivalent to lossing the retransmitted packet.
 
@@
 \section{Observations and Discussion}\label{disc}
+This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, retransmission or probe loss and two packet 
+tail loss.
 
 \subsection{TCP}
-Retransmission behavior of TCP is seen a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
+Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
 starts with the results using TCP and compares it with that of the MPTCP. Figures ~\ref{t1p},~\ref{t2p},~\ref{t1pp} represent the comparison of TCP 
@@
 
-In the case of single packet tail loss, there is no difference in burst completion times of ER setting 0,2 and 3. All the cases depend on the RTO
-to recover from loss. In two packet tail loss case, ER setting 0 and 2 have similar performance while ER setting 3 improves burst completion time
-by using TLP. If the probe or retransmission is lost, then ER 3 still performs better than 0 and 2 as it waits for one RTO and one probe timeout
-instead of two RTOs in other two cases.
+In the case of single packet tail loss, there is no difference in burst completion times of RTO and TLP settings. TLP implementation in linux accounts
+for the delayed ACK and waits 200ms before sending probe. Hence retransmission timeout is the mode of loss recovery in both RTO and TLP. This can be avoided by tuning
+the ACKs as discussed in~\cite{Rajiullah:2015}.
+In the case of two packet tail loss, TLP performs better than RTO as it employs a combination of TLP and ER to recover from losses. 
+If the probe or retransmission is lost, then TLP still performs better than STD as it waits for one RTO and one probe timeout instead of two RTOs.
 
@@
 
-The expected retransmission behavior of MPTCP for different settings of ER is shown 
-in~\ref{timing1P},~\ref{timing1PP} and ~\ref{timing2P}.
+The expected retransmission behavior of MPTCP for standard TCP and Linux default settings is shown 
+in~\ref{timing1P},~\ref{timing1PP} and ~\ref{timing2P}. The actual pattern might differ in cases,
+where the RTT is larger than 200ms.
 
@@
 \end{center}
-\caption{Timing diagram of MPTCP behavior with one packet tail loss a) STD b) LIN}\label{timing1P}
+\caption{Timing diagram of MPTCP behavior with one packet tail loss a) RTO b) TLP}\label{timing1P}
 \end{figure}
@@
 \end{center}
-\caption{Timing diagram of MPTCP behavior with one packet and next probe loss a) STD b) LIN}\label{timing1PP}
+\caption{Timing diagram of MPTCP behavior with one packet and next probe loss a) RTO b) TLP}\label{timing1PP}
 \end{figure}
@@
 \end{center}
-\caption{Timing diagram of MPTCP behavior with two packet tail loss a) STD b) LIN}\label{timing2P}
+\caption{Timing diagram of MPTCP behavior with two packet tail loss a) RTO b) TLP}\label{timing2P}
 \end{figure}
@@
 Experiments performed on CORE emulator with the setup mentioned in section\ref{exsetup}.
-MPTCP is sensitive to path asymmetry in general due to the default scheduler being shortest RTT based scheduler. Wireshark and tcptrace
-analysis provides information on the TCP retransmissions at the subflow level.
+MPTCP is sensitive to path asymmetry in general due to the default scheduler being shortest RTT based scheduler.
 
-Figure~\ref{1p} provides burst completion time comparison of different ER sysctl settings considered with one packet loss. The results show the improvements
-in burst completion time with ER and TLP enabled compared to both disabled only when there is delay difference.
+The burst completion times on single packet tail loss calculated for each delay configuration and retransmission 
+setting is depicted in figure~\ref{1p}. There is no difference in performance of RTO and TLP. The retransmission
+timeout and the probe timeout are same as there is only one outstanding packet. This is a special case in TLP
+specification that adds 200ms to PTO accommodating the case of delayed ACK. The minimum of PTO and RTO is considered
+for PTO.
 
@@
 
-In packet trace inspection, it is easy to observe the use of TLP with two packet tail loss. The last packet will be retransmitted first in this case.
-With one packet tail loss one has to compare the RTT and last packet retransmission time to check the use of TLP.
-To further validate the results with one packet tail loss, we consider two packet tail loss and the results are shown in~\ref{2p}.
-In this case, ER=3 setting performs better than the other two settings in all scenarios except when the packets lost on high delay interface.
+In the case of single packet the last outstanding packet is sent as probe and in case of n-degree tail loss,
+the latest packet is sent as probe to trigger fast retransmit of the packets in between. In order to observe
+the performance of n-degree tail loss, we consider dropping the last two packets. 
+The results of two packet tail loss are shown in~\ref{2p}. In this case, TLP performs better than the other 
+two settings in all scenarios except when the packets lost on high delay interface. The delay difference is
+large enough that the retransmission on alternate path are faster than the probe on the primary path. It 
+is PTO time slower than the RTO setting.
 
@@
 
-To further study the performance of TLP in improving latency, we tried to drop the probe packet along with the last packet. In this case ER=0
-setting results much lower burst completion times. Use of TLP did not improve the burst completion time as we saw in TCP for ER=3. The reason
+To further study the performance of TLP in improving latency, we tried to drop the probe packet along with the last packet. In this case STD
+setting results much lower burst completion times. Use of TLP did not improve the burst completion time as we saw in TCP for TLP. The reason
 lies in the way the retransmission and loss probe transmission carried out in MPTCP. Retransmission occurs after an RTO time on both paths.
@@
 
-\section{Improvments to TLP for MPTCP}
+\section{Improvements to TLP for MPTCP}
 The goal of Tail Loss Probe(TLP) is to reduce tail latency of short flows.
@@
 But in the Event of Probe timeout, the loss probe packet is being sent on the same path without injecting in to scheduler.
-Modifications should reinject the probe packet in to the mptcp scheduler in the event of PTO.
-
-\subsection{Example scenario where MPTCP TLP could incur more latency}
-
-Use a traffic pattern where there is new burst to send before the RTO after a tail loss triggering TLP on a flow.
+We modified the probe sending mechanism for MPTCP by reinjecting the probe packet in to the mptcp scheduler in the event of PTO.
 
@@
 
-Initial testing with customized program to send bursts of packets with fixed burst seperation. Adjust the burst seperation to get the desired scenario.
 Results comparing the MPTCP-TLP for the considered cases are provided in figures~\ref{1pn},~\ref{2pn} and ~\ref{1ppn}. The improvement in latency
-performance is seen Single packet tail loss with probe loss scenario. Single packet loss and Two packet loss scenarios provide increase in latency.
-Analysis required in these two cases.
+performance is seen Single packet tail loss with probe loss scenario. Single packet loss and Two packet loss scenarios provide same performance as 
+before. Performance improved in the case with large delay difference and tail drop in the high delay path.
 
@@
 
+\section{Conclusion}\label{conc}
+
 \bibliographystyle{IEEEtran}

### 2017-02-27 plots/fetchDelay.sh

@@ -1,14 +1,14 @@
 #!/bin/bash
 
-FILES1=/home/vkirany/mptcptests/tcp/
-for e in 'er0' 'er2' 'er3' 'er4'
+FILES1=/home/vkirany/mptcptests/mptcp/
+for e in 'er3' 
 do
 for p in 1P 2P 1P+P
 do
 for x in 'tail20-20' 'tail20-30' 'tail30-20' 'tail20-120' 'tail120-20'
 do
 echo "ER $e and Scene $x and P $p"
-f1=$FILES1$e'/'$x'/'$p'/'Run1'/'delay
+f1=$FILES1$e'/'$x'/'$p'/'Run2'/'delay
 
  	cat $f1
 	

### 2017-02-27 plots/plot_tlp_c.py

@@ -5,37 +5,24 @@ import matplotlib.pyplot as plt
 
 n_groups = 5 
 
-p1e0=(422,484,453,445,582)
-p1e2=(423,484,453,444,582)
-p1e3=(432,482,463,443,592)
-p1e4=(433,483,464,443,592)
-
-p2e0=(378,438,409,493,581)
-p2e2=(378,437,407,494,582)
-p2e3=(301,352,355,362,676)
-p2e4=(481,542,532,552,887)
-
-p1pe0=(423,483,454,534,583)
-p1pe2=(423,483,453,536,582)
-p1pe3=(646,742,668,698,957)
-p1pe4=(647,741,667,700,956)
+p1e3=(386,432,464,394,927)
+p2e3=(277,280,355,289,1148)
+p1pe3=(627,680,667,639,1442)
 
 
-tp1e0=(490,489,540,491,1002)
-tp2e0=(586,585,645,585,1279)
-tp1pe0=(1004,1004,1073,1004,1746)
+p1e0=(385,432,453,394,878)
+p2e0=(385,397,407,435,877)
+p1pe0=(384,443,453,495,878)
 
-tp1e2=(491,490,541,490,1003)
-tp2e2=(585,586,646,586,1279)
-tp1pe2=(1004,1004,1073,1003,1743)
 
-tp1e3=(491,490,541,491,1003)
-tp2e3=(462,460,533,462,1187)
-tp1pe3=(748,748,807,747,1375)
+tp1e0=(474,474,540,474,1613)
+tp2e0=(469,469,645,469,1631,)
+tp1pe0=(973,973,1070,973,2594)
 
-tp1e4=(490,490,540,491,1004)
-tp2e4=(657,657,728,658,1371)
-tp1pe4=(748,747,808,748,1376)
+
+tp1e3=(474,474,540,474,1612)
+tp2e3=(326,326,533,326,1645)
+tp1pe3=(722,722,808,722,2103)
 
 
 fig, ax = plt.subplots()
@@ -46,17 +33,17 @@ margin = 0.05
 bar_width = (1-2*margin)/3
 
 
-rects1 = plt.bar(index+margin, tp2e0 , bar_width,
+rects1 = plt.bar(index+margin, p1pe0 , bar_width,
                  color='b',
-                 label='er=0')
+                 label='RTO')
 
-rects2 = plt.bar(index+margin + bar_width, tp2e2 , bar_width,
[…]

### 2017-02-27 plots/plot_tlp_new.py

@@ -5,18 +5,18 @@ import matplotlib.pyplot as plt
 
 n_groups = 5 
 
-p1e3=(433,482,463,443,592)
-p2e3=(301,352,355,362,676)
-p1pe3=(646,742,668,698,957)
+p1e3=(386,432,463,394,928)
+p2e3=(277,280,356,288,1148)
+p1pe3=(627,680,667,639,1442)
 
 
-np1e3=(423,390,453,389,581)
-np2e3=(303,407,407,403,674)
-np1pe3=(423,483,452,443,581)
+np1e3=(385,432,451,394,827)
+np2e3=(231,243,272,288,752)
+np1pe3=(385,443,451,499,827)
 
-tp1e3=(491,490,541,491,1003)
-tp2e3=(462,460,533,462,1187)
-tp1pe3=(748,748,807,747,1375)
+tp1e3=(474,474,540,474,1612)
+tp2e3=(326,326,533,326,1645)
+tp1pe3=(722,722,808,722,2103)
 
 
 
@@ -28,15 +28,15 @@ margin = 0.05
 bar_width = (1-2*margin)/3
 
 
-rects1 = plt.bar(index+margin, p1pe3 , bar_width,
+rects1 = plt.bar(index+margin, p2e3 , bar_width,
                  color='b',
                  label='mptcp')
 
-rects2 = plt.bar(index+margin + bar_width, np1pe3 , bar_width,
+rects2 = plt.bar(index+margin + bar_width, np2e3 , bar_width,
                  color='g',
                  label='mptcp-New')
 
-rects3 = plt.bar(index+margin + 2*bar_width, tp1pe3, bar_width,
+rects3 = plt.bar(index+margin + 2*bar_width, tp2e3, bar_width,
                  color='r',
                  label='tcp')
 

### 2017-03-05 main.tex
Update

@@
 of loss recovery greatly influences the completion time of the flows.
-In this paper we focus on the state of the art los recovery mechanisms for TCP and Multipath TCP.
+In this paper we focus on the state of the art loss recovery mechanisms for TCP and Multipath TCP.
 We use controlled tail loss scenarios
-to calculate the burst completion time. We propose an enhancement to the tail loss recovery in Multipath TCP. Our experiment results show performance improvement in controlled scenarios. 
+to evaluate the performance of loss recovery mechanisms by calculating the burst completion time for each scenario.
+Based on the observations, we propose an enhancement to the tail loss recovery in Multipath TCP.
+Our experiment results show latency performance improvement in controlled scenarios.
 \end{abstract}
@@
 
-Multipath TCP is an experimental standard proposed as an extension to TCP~\cite{rfc6824}. It allows devices with multiple interfaces to transmit simultaneously on both interfaces.  
-In Multipath TCP(MPTCP), a connection can have multiple TCP sub flows using different interfaces on different routes. Each TCP subflow imparts the functional behavior of a standard TCP flow. 
+Multipath TCP is an experimental standard proposed as an extension to TCP~\cite{rfc6824}. It allows devices with multiple interfaces to transmit data simultaneously on both interfaces.  
+Using Multipath TCP(MPTCP), a connection can have multiple TCP sub flows using different interfaces on different routes. Each TCP subflow imparts the functional behavior of a standard TCP flow. 
 MPTCP uses a connection level congestion window as well as sub flow level congestion windows for each sub flow.
 Packet losses in each sub flow are detected and recovered in a similar fashion as that of TCP.
-However, It is not restricted in the standard on how the loss recovery happens in the implementation and which sub-flow retransmits the lost packets.
+However, the standard does not restrict on, how the loss recovery should happen in the implementation and which sub-flow should be retransmit the lost packets.
 If the recovery is handled at the meta level, the lost packet may be rescheduled and retransmitted at the available sub flow with lowest RTT.
 If the recovery is handled at the flow level, the packet may be retransmitted in the same sub-flow.
-An important part of this recovery process is to follow the TCP semantics.
+An important part of MPTCP packet loss recovery process is to follow the TCP semantics and must retransmit lost packet on the initial path of transfer.
 
-TCP has two mechanisms for detecting and recovering from packet losses namely Fast Recovery (FR) and Retransmission Timeouts(RTOs).
-Fast recovery is triggered on receipt of predecided number of duplicate acknowledgements, considering it as a loss indication.
+TCP has mainly two mechanisms for detecting and recovering from packet losses namely Fast Recovery (FR) and Retransmission Timeouts(RTOs).
+Fast recovery is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as a loss indication.
 RTOs wait for the packet loss to trigger a timeout for retransmitting that packet.
 For long flows where there are sufficiently large number of packets to be transferred, FR is quicker than RTOs in detecting and recovering packet losses.
-On the other hand shorter flows which are a majority in the web traffic have to depend on RTOs for packet loss detection and recovery thus incurring more latency.
+On the other hand shorter flows which are a majority in the web traffic, have to depend on RTOs for packet loss detection and recovery, thus incurring more latency.
 A single packet loss in a short flow may take many RTTs to detect and recover.
 This scenario is also applicable to the packets at the end of the flow (or tail) in a long flow.
+Though, FR and RTOs are reliable even in tail loss recovery, they are not efficient in terms of recovery completion time.
 
-TCP Loss Probe (TLP)~\cite{Flach:2013} a mechanism that allows flows to detect and recover from tail losses much faster than an RTO, thereby speeding up short transfers.
-With TLP a packet loss in the middle
-of a packet train as well as at the tail end will now trigger the same fast recovery mechanisms.
-It assumes other algorithms such as early retransmit~\cite{rfc5827} and FACK threshold based recovery are
-present.
-
-Retransmission schemes of TCP are of significant importance in achieving lower flow completion times. TLP is the recent scheme proposed and implemented in Linux kernel. Further 
-evaluationsin~\cite{Rajiullah:2015}, showed that TLP provides significant reductions in latency for short flows with one to n-degree tail losses.
-In this paper we provide the related research on retransmission in section~\ref{relwork} leading to the experimental setup.
+TCP Loss Probe (TLP)~\cite{Flach:2013} is a mechanism to detect and recover from tail losses by employing a loss probe packet for a timeout shorter than the RTO.
+Probe timeout enables the short flows to recover packet loss at the tail by triggering the same fast recovery mechanisms.
+It assumes other algorithms such as early retransmit~\cite{rfc5827} and FACK threshold based recovery are enabled.
+TLP is available in Linux kernel and enabled by default.
+Further evaluations in~\cite{Rajiullah:2015}, showed that TLP provides significant reductions in latency for short flows with one to n-degree tail losses.
+In this paper we provide the related research on MPTCP retransmission in section~\ref{relwork} and experimental setup in section~\ref{exsetup}.
+We discuss the observations from the experiments in section~\ref{disc} that lead to our proposed improvement on current Linux implementation of TLP for MPTCP in section~\ref{impr}.
+In section~\ref{eval}, we provide evaluation of proposed TLP for MPTCP implementation in Linux and conclude the paper in section~\ref{conc}.
 
@@
 
+Retransmission schemes of TCP are of significant importance in achieving lower flow completion time.
+Several research studies led to improved retransmission strategies for TCP.
+The current Linux TCP implementation provides scope for improvement in specific cases of packet loss scenarios~\cite{prr}.
+In a TCP flow, sequence numbers are used to identify the lost packets and retransmit the respective packets.
 Multipath TCP uses two levels of sequence numbers to support efficient data transfer, namely data sequence and TCP sequence.
@@
 However, the interaction with two levels makes the problem of retransmission more challenging than in TCP.
-The dual sequence numbering enables MPTCP to respond to loss of packets or loss of subflows by retransmitting the lost packets on an alternate path.
+The dual sequence numbering enables MPTCP to respond to loss of packets or loss of subflow by retransmitting the lost packets on an alternate path.
 It allows a packet to be retransmitted on a different subflow than that of the originally sent flow.
 The Linux implementation of MPTCP uses a set of retransmission heuristics to decide on the retransmissions.
-The data outstanding on timed out subflows should be rescheduled for transmission on different subflows using first timeout as the indicator.
+The data outstanding on a timed out subflow should be rescheduled for transmission on different subflow using first timeout as the indicator.
 Fast retransmit on a subflow will not trigger retransmission on another subflow.
-It is often the case in MPTCP connection that there is path asymmetry among the subflow characteristics.
-Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmtting on the fast paths.
+The target devices of MPTCP contain a degree of asymmetry in their interface characteristics leading to flows with different delay characteristics.
+Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fast paths.
+Quick retransmission comes with a cost of redundant packets while the MPTCP standard suggests more conservative approaches.
+MPTCP is designed to exploit the multiple interfaces and often one or more of these interfaces could be wireless with completely different characteristics than that of wired interfaces.
+Authors in~\cite{Shin} argue that the calculation RTO for MPTCP flows should include the interface characteristics to improve loss recovery time.
 
-\section{Scope}\label{scope}
+\subsection{Scope}\label{scope}
 
 This paper analyzes the performance of state of the art loss recovery mechanism of MPTCP and presents case for improvement in the tail loss scenario. 
-For understanding the loss recovery and retransmission policies in MPTCP linux implementation version 0.91, we consider specific cases of tail losses.  
-No tuning of TCP settings performed such as adjusting the delayed ACK behavior or time and flip on Nagle algorithm. We consider the standard settings
-of Linux kernel. In the following experiments, we do not consider the congestion control interaction with the retransmission. We consider mainly failure
-induced packet loss and its recovery.
+For understanding the loss recovery and retransmission policies in MPTCP Linux implementation version 0.91, we consider specific cases of tail losses.  
+No tuning of TCP settings performed such as adjusting the delayed ACK behavior. We consider the standard settings of Linux kernel. In the following experiments, 
+we do not consider the congestion control interaction with the retransmission. We consider mainly path failure or characteristics induced packet loss and its recovery.
+The Linux implementation of TCP is considered as the baseline TCP retransmission for implementation flexibility and availability of TLP support, though it differs from IETF standard.
 
@@
 
-Experiments use CORE emulation platform with a simple topology as depicted in Figure~\ref{fig1}.
-Client is connected to two wireless interfaces 3G/4G and WLAN. Server is connected to a wired router.
+Experiments use CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}.
+with client node connected to two wireless interfaces 3G/4G and WLAN and server node connected to a wired router.
 Characteristics of the connection setup and assumptions about the parameters are provided in table~\ref{tab1}.
+The parameters are chosen to be close to real values of bandwidth and delay of respective technologies.
 There is no attempt in this study to focus on the effect of link characteristics in retransmission performance.
@@
 MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between client and server on the MPTCP protocol 
-support and interface availability. So one has to wait in time and packets for the second subflow establishment to understand the MPTCP retransmissions. We send two bursts of 80 
+support and interface availability. So one has to wait in time and packets for the second subflow establishment to understand the MPTCP retransmissions in multiple flows. We send two bursts of 80 
 packets each and drop tail packets on one of the interface. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that
-subflow. Three test cases are evaluated with one packet tail loss, two packet tail loss and  one packet along with probe loss.
+subflow. Three test cases are evaluated with one packet tail loss, two packet tail loss and one packet along with probe(in TLP) or retransmission(in RTO) loss.
 
@@
 ER and TLP. Standard TCP used by other operating systems might not use ER and TLP. We use RTO to refer sysctl setting of 0 and TLP to refer sysctl setting of 3 throughout the paper.
-In standard TCP (RTO), disabling ER and TLP gives the performance of RTO. In Linux default TCP, enabling delayed ER gives the combined performance of RTO, delayed ER and TLP.
-In standard TCP, probe loss scenario is equivalent to lossing the retransmitted packet.
+In Linux, disabling ER and TLP gives the performance of RTO. In Linux default TCP, enabling delayed ER gives the combined performance of RTO, delayed ER and TLP.
+In RTO, probe loss scenario is equivalent to losing the retransmitted packet.
 
@@
 interfaces with one way delay on each interface ranging from 20ms to 120ms. We consider 5 scenarios with delay pairings 20ms-20ms, 20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand 
-the effect of delay difference in the performance. These scenarios are choosen to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in link delays.
+the effect of delay difference in the performance. These scenarios are chosen to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in link delays.
 
@@
 Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
-starts with the results using TCP and compares it with that of the MPTCP. Figures ~\ref{t1p},~\ref{t2p},~\ref{t1pp} represent the comparison of TCP 
-burst completion times. Path asymmetry is irrelevant as the TCP uses single path.
+starts with the results using TCP and compare it with that of the MPTCP. Figures ~\ref{t1p},~\ref{t2p},~\ref{t1pp} represent the comparison of TCP 
+burst completion times. Path asymmetry is irrelevant in this case as the TCP uses single path.
 
-In the case of single packet tail loss, there is no difference in burst completion times of RTO and TLP settings. TLP implementation in linux accounts
+In the case of single packet tail loss, there is no difference in burst completion times of RTO and TLP settings. TLP implementation in Linux accounts
 for the delayed ACK and waits 200ms before sending probe. Hence retransmission timeout is the mode of loss recovery in both RTO and TLP. This can be avoided by tuning
-the ACKs as discussed in~\cite{Rajiullah:2015}.
-In the case of two packet tail loss, TLP performs better than RTO as it employs a combination of TLP and ER to recover from losses. 
-If the probe or retransmission is lost, then TLP still performs better than STD as it waits for one RTO and one probe timeout instead of two RTOs.
+the ACKs as discussed in~\cite{Rajiullah:2015}. In the case of two packet tail loss, TLP performs better than RTO as it employs a combination of TLP and ER to recover from losses. 
+If the probe or retransmission is lost, then TLP still performs better than RTO as it waits for one RTO and one probe timeout instead of two RTOs.
 
@@
 \includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
-\caption{Single packet tail loss together with probe loss using TCP}\label{t1pp}
+\caption{Single packet tail loss together with retransmission loss using TCP}\label{t1pp}
 \end{center}
@@
 
-Experiments performed on CORE emulator with the setup mentioned in section\ref{exsetup}.
+Experiments performed on CORE emulator with the setup mentioned in section~\ref{exsetup}.
 MPTCP is sensitive to path asymmetry in general due to the default scheduler being shortest RTT based scheduler.
@@
 two settings in all scenarios except when the packets lost on high delay interface. The delay difference is
-large enough that the retransmission on alternate path are faster than the probe on the primary path. It 
+large enough that the retransmission on alternate path is faster than the probe on the primary path. It 
 is PTO time slower than the RTO setting.
@@
 
-To further study the performance of TLP in improving latency, we tried to drop the probe packet along with the last packet. In this case STD
+To further study the performance of TLP in improving latency, we tried to drop the probe packet along with the last packet. In this case RTO
 setting results much lower burst completion times. Use of TLP did not improve the burst completion time as we saw in TCP for TLP. The reason
@@
 Even if retransmission fails on one path, the client receives the packet on the other path as shown in Fig~\ref{timing1P}. For TLP, in
-current implementation, the loss probe sent on the actual path of first transmission but not on both paths. This incorrect impartation of
+current implementation, the loss probe sent on the actual path of first transmission but not on both paths. This incomplete impartation of
 TLP from TCP to MPTCP led to the increase burst completion time by an RTO for MPTCP.
@@
 
-\section{Improvements to TLP for MPTCP}
+\section{Improvements to TLP for MPTCP}\label{impr}
 The goal of Tail Loss Probe(TLP) is to reduce tail latency of short flows.
-It achieves this by converting retransmission timeouts (RTOs) occuring due to tail losses (losses at end of transactions) into fast recovery.
+It achieves this by converting retransmission timeouts (RTOs) occurring due to tail losses (losses at end of transactions) into fast recovery.
 TLP transmits one packet in two round-trips when a connection is in Open state and isn't receiving any ACKs.
@@
 Current implementation of TLP is at the TCP flow level.
-In the event of RTO timeout the MPTCP retransmission happens with a reinjection in to scheduler along with sending lost packet on the same path.
-But in the Event of Probe timeout, the loss probe packet is being sent on the same path without injecting in to scheduler.
-We modified the probe sending mechanism for MPTCP by reinjecting the probe packet in to the mptcp scheduler in the event of PTO.
+In the event of RTO timeout, the MPTCP retransmission happens with a re-injection in to scheduler along with sending lost packet on the same path.
+But in the Event of Probe timeout, the loss probe packet is being sent on the same path without being re-injected in to scheduler.
+We propose a modification the probe sending mechanism for MPTCP by re-injecting the probe packet in to the MPTCP scheduler in the event of PTO.
+As the probe timeout is meant for acting faster than RTO timeout with essentially same consequence of retransmission, it is relevant to follow the process upon an RTO.
+It is expected that this modification is useful when the alternate path has lower delay than the packet original transmission path.
+The timing diagram depicted in figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP on each of the loss cases.
 
-\section{Evaluation of proposal}
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.46\textwidth]{images/timingER3NewTLP.pdf}
+\end{center}
+\caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) One Packet b) Two Packet c) One packet and Retransmission}\label{timingNew}
+\end{figure}
 
-Results comparing the MPTCP-TLP for the considered cases are provided in figures~\ref{1pn},~\ref{2pn} and ~\ref{1ppn}. The improvement in latency
-performance is seen Single packet tail loss with probe loss scenario. Single packet loss and Two packet loss scenarios provide same performance as 
-before. Performance improved in the case with large delay difference and tail drop in the high delay path.
+\section{Evaluation of modified MPTCP TLP}\label{eval}
+
+Results comparing the burst completion time for new MPTCP TLP with that of TCP and MPTCP TLP for the considered cases are provided in figures~\ref{1pn},~\ref{2pn} and ~\ref{1ppn}. 
+In figure~\ref{1pn}, mptcp-New performs better than mptcp in all scenarios, performs equal to tcp in symmetric scenarios and performs better than tcp in asymmetric scenarios.
+This improvement is due to the wait for RTO is eliminated for a packet to be retransmitted on the alternate path.
+In two packet tail loss scenario as in figure~\ref{2pn}, mptcp-New performs better than mptcp in all scenarios and better than tcp in all scenarios except when the alternate path has higher delay than the original path.
+In the single packet together with probe loss scenario as in figure~\ref{1ppn}, mptcp-New performs better than both mptcp and tcp in all scenarios with PTO triggering the probe retransmission on alternate path much earlier than the RTO.
 
@@
 \includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/1PNew.pdf}
-\caption{Single packet tail loss comparision with MPTCP-TLP}\label{1pn}
+\caption{Single packet tail loss comparison with MPTCP-TLP}\label{1pn}
 \end{center}
@@
 \section{Conclusion}\label{conc}
+This paper provides the short comparative analysis of retransmission behavior with TCP and MPTCP protocols using possible standard loss patterns for tail losses.
+The three cases of tail losses together with deterministic loss pattern allow us to dissect the retransmission behavior at packet level to understand possible improvements.
+We propose a less conservative approach to trigger retransmission on an alternate path by altering the loss indication as PTO timeout than the RTO timeout as in the Linux implementation.
+Our experiments on CORE emulator using  modified Linux implementation, prove that the proposed approach is in fact improves the burst completion time in most cases and equals the existing
+implementation in other cases.
+Proposed approach improve the performance when the alternate path has lower delay than the original loss path and the advantage increases with degree of asymmetry between the paths.
+This approach is very effective in case of probe loss which otherwise incur RTO timeout to trigger retransmission on either path.
 

### 2017-03-07 main.tex
Update after comments

@@
-\documentclass[10pt,draftcls,twocolumn]{IEEEconf}
+\documentclass[10pt,conference,compsoc]{IEEEtran}
 \usepackage[utf8]{inputenc}
 \usepackage{graphicx}
+\usepackage[nocompress]{cite}
 
 % Title Page
-\title{Loss recovery with tail loss probe in Multipath TCP}
-\author{KAU}
+\title{Probe or Wait: Handling tail losses using Multipath TCP}
+
+\author{\IEEEauthorblockN{Kiran~Yedugundla, Per~Hurtig, Anna~Brunstrom}
+\IEEEauthorblockA{Dept. of Computer Science, Karlstad University, Karlstad, Sweden}}
 
@@
 \begin{abstract}
-Packet losses are a well known cause of performance hindrance in the Internet. Transport protocols recover from the packet losses in order to provide reliable end to end communication. The efficiency
-of loss recovery greatly influences the completion time of the flows.
+Packet losses are known to affect the performance of latency sensitive applications in the Internet such as media streaming and gaming.
+Transport protocols recover from packet loss in order to provide reliable end to end communication and improving the quality of user experience.
+The efficiency of loss recovery greatly influences the completion time of flows.
 In this paper we focus on the state of the art loss recovery mechanisms for TCP and Multipath TCP.
-We use controlled tail loss scenarios
-to evaluate the performance of loss recovery mechanisms by calculating the burst completion time for each scenario.
-Based on the observations, we propose an enhancement to the tail loss recovery in Multipath TCP.
-Our experiment results show latency performance improvement in controlled scenarios.
+We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms by calculating the burst completion time for each scenario.
+Based on the observations, we propose an enhancement to the tail loss recovery in Multipath TCP to improve the loss recovery time.
+Our experiment results show end to end latency performance improvement in controlled scenarios by reducing the burst completion time.
 \end{abstract}
@@
 
-Multipath TCP is an experimental standard proposed as an extension to TCP~\cite{rfc6824}. It allows devices with multiple interfaces to transmit data simultaneously on both interfaces.  
-Using Multipath TCP(MPTCP), a connection can have multiple TCP sub flows using different interfaces on different routes. Each TCP subflow imparts the functional behavior of a standard TCP flow. 
-MPTCP uses a connection level congestion window as well as sub flow level congestion windows for each sub flow.
-Packet losses in each sub flow are detected and recovered in a similar fashion as that of TCP.
-However, the standard does not restrict on, how the loss recovery should happen in the implementation and which sub-flow should be retransmit the lost packets.
-If the recovery is handled at the meta level, the lost packet may be rescheduled and retransmitted at the available sub flow with lowest RTT.
-If the recovery is handled at the flow level, the packet may be retransmitted in the same sub-flow.
+Multipath TCP(MPTCP) is an experimental standard proposed as an extension to TCP~\cite{rfc6824}.
+It allows devices with multiple interfaces to transmit data simultaneously on both interfaces, there by improving the throughput of end to end connection.
+Using MPTCP, a connection can have multiple TCP subflows using different interfaces on different routes.
+Each TCP subflow imparts the functional behavior of a standard TCP flow.
+
+TCP has mainly two mechanisms for detecting and recovering from packet losses namely Fast Retransmit (FR) and Retransmission Timeout~(RTO)~\cite{Flach:2013}.
+Fast retransmit is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as a loss indication.
+In case of RTO, sender waits for the packet loss to trigger a timeout before retransmitting that packet.
+For long flows where there are sufficiently large number of packets to be transferred, FR is quicker than RTO in detecting and recovering packet losses.
+On the other hand, shorter flows have to depend on RTO for packet loss detection and recovery, thus incurring more latency.
+A single packet loss in a short flow may take many RTTs to detect and recover.
+This scenario is also applicable to the packets at the end of the flow (or tail) in a long flow.
+Though, FR and RTO are reliable even in tail loss recovery, they are not efficient in terms of recovery completion time.
+
+ MPTCP connection starts with one TCP flow to send data and with available network interfaces and protocol support, additional flows are added.
+MPTCP uses a connection level congestion window as well as subflow level congestion windows for each subflow.
+Data is split across the flows and each data packet is scheduled on one of the subflows based on minimum RTT.
+Packet losses in each subflow are detected and recovered in a similar fashion as that of TCP.
+However, the MPTCP standard does not restrict on, how the loss recovery should happen in the implementation and which subflow should be retransmit the lost packets.
+Loss recovery can be conducted at different levels: the meta-level and the subflow level.
+If the recovery is handled at the meta level, the lost packet may be rescheduled and retransmitted at the available subflow with lowest RTT.
+If the recovery is handled at the flow level, the packet may be retransmitted in the same subflow.
 An important part of MPTCP packet loss recovery process is to follow the TCP semantics and must retransmit lost packet on the initial path of transfer.
 
-TCP has mainly two mechanisms for detecting and recovering from packet losses namely Fast Recovery (FR) and Retransmission Timeouts(RTOs).
-Fast recovery is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as a loss indication.
-RTOs wait for the packet loss to trigger a timeout for retransmitting that packet.
-For long flows where there are sufficiently large number of packets to be transferred, FR is quicker than RTOs in detecting and recovering packet losses.
-On the other hand shorter flows which are a majority in the web traffic, have to depend on RTOs for packet loss detection and recovery, thus incurring more latency.
-A single packet loss in a short flow may take many RTTs to detect and recover.
-This scenario is also applicable to the packets at the end of the flow (or tail) in a long flow.
-Though, FR and RTOs are reliable even in tail loss recovery, they are not efficient in terms of recovery completion time.
-
 TCP Loss Probe (TLP)~\cite{Flach:2013} is a mechanism to detect and recover from tail losses by employing a loss probe packet for a timeout shorter than the RTO.
-Probe timeout enables the short flows to recover packet loss at the tail by triggering the same fast recovery mechanisms.
-It assumes other algorithms such as early retransmit~\cite{rfc5827} and FACK threshold based recovery are enabled.
+Probe timeout enables the short flows to recover packet loss at the tail by triggering the fast retransmit.
+It assumes other algorithms such as Early Retransmit~(ER)\cite{rfc5827} and Forward Acknowledgement~(FACK)~\cite{FACK} threshold based recovery are enabled.
 TLP is available in Linux kernel and enabled by default.
 Further evaluations in~\cite{Rajiullah:2015}, showed that TLP provides significant reductions in latency for short flows with one to n-degree tail losses.
-In this paper we provide the related research on MPTCP retransmission in section~\ref{relwork} and experimental setup in section~\ref{exsetup}.
-We discuss the observations from the experiments in section~\ref{disc} that lead to our proposed improvement on current Linux implementation of TLP for MPTCP in section~\ref{impr}.
-In section~\ref{eval}, we provide evaluation of proposed TLP for MPTCP implementation in Linux and conclude the paper in section~\ref{conc}.
+
+In this paper, we analyze the retransmission behavior of MPTCP in selected tail loss cases to identify the scope for improvement.
+We propose a change in the retransmission strategy of MPTCP to improve the latency performance.
+We provide the related research on MPTCP retransmission in section~\ref{relwork}, and experimental setup in section~\ref{exsetup}.
+We discuss the observations from the experiments in section~\ref{disc}, that lead to our proposed improvement on current Linux implementation of TLP for MPTCP in section~\ref{impr}.
+In section~\ref{eval}, we provide an evaluation of the proposed TLP for MPTCP implementation in Linux and conclude the paper in section~\ref{conc}.
 
@@
 Several research studies led to improved retransmission strategies for TCP.
-The current Linux TCP implementation provides scope for improvement in specific cases of packet loss scenarios~\cite{prr}.
+The current Linux TCP implementation provides space for improvement in specific cases of packet loss scenarios~\cite{prr}.
 In a TCP flow, sequence numbers are used to identify the lost packets and retransmit the respective packets.
@@
 However, the interaction with two levels makes the problem of retransmission more challenging than in TCP.
-The dual sequence numbering enables MPTCP to respond to loss of packets or loss of subflow by retransmitting the lost packets on an alternate path.
+The dual sequence numbering enables MPTCP to respond to loss of packets by retransmitting the lost packets on an alternate path.
 It allows a packet to be retransmitted on a different subflow than that of the originally sent flow.
+
 The Linux implementation of MPTCP uses a set of retransmission heuristics to decide on the retransmissions.
-The data outstanding on a timed out subflow should be rescheduled for transmission on different subflow using first timeout as the indicator.
+The data outstanding on a timed out subflow should be rescheduled for transmission on a different subflow using timeout as the indicator.
 Fast retransmit on a subflow will not trigger retransmission on another subflow.
-The target devices of MPTCP contain a degree of asymmetry in their interface characteristics leading to flows with different delay characteristics.
+The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 4G vs WLAN, leading to flows with different delay characteristics.
 Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fast paths.
-Quick retransmission comes with a cost of redundant packets while the MPTCP standard suggests more conservative approaches.
+Quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always to follow TCP semantics.
+MPTCP standard suggests more conservative approaches.
 MPTCP is designed to exploit the multiple interfaces and often one or more of these interfaces could be wireless with completely different characteristics than that of wired interfaces.
@@
 
-\subsection{Scope}\label{scope}
+%\subsection{Scope}\label{scope}
 
-This paper analyzes the performance of state of the art loss recovery mechanism of MPTCP and presents case for improvement in the tail loss scenario. 
-For understanding the loss recovery and retransmission policies in MPTCP Linux implementation version 0.91, we consider specific cases of tail losses.  
-No tuning of TCP settings performed such as adjusting the delayed ACK behavior. We consider the standard settings of Linux kernel. In the following experiments, 
-we do not consider the congestion control interaction with the retransmission. We consider mainly path failure or characteristics induced packet loss and its recovery.
-The Linux implementation of TCP is considered as the baseline TCP retransmission for implementation flexibility and availability of TLP support, though it differs from IETF standard.
+This paper analyzes the performance of state of the art loss recovery mechanism of MPTCP and presents a case for improvement in the tail loss scenario. 
+For understanding the loss recovery and retransmission policies in MPTCP Linux implementation version 0.91, we consider several specific cases of tail losses.  
+No tuning of TCP settings such as, adjusting the delayed ACK behavior are performed.
+We consider the standard settings of the Linux kernel.
+We consider mainly various cases of tail loss and its recovery.
+The Linux implementation of TCP is considered as the baseline TCP retransmission for implementation flexibility and availability of TLP support.
 
@@
 
-Experiments use CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}.
-with client node connected to two wireless interfaces 3G/4G and WLAN and server node connected to a wired router.
-Characteristics of the connection setup and assumptions about the parameters are provided in table~\ref{tab1}.
+Experiments use the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}.
+with client node connected to two wireless interfaces 3G or 4G and WLAN and server node connected to a wired router.
+Characteristics of the connection setup and assumptions about the parameters are provided in Table~\ref{tab1}.
 The parameters are chosen to be close to real values of bandwidth and delay of respective technologies.
-There is no attempt in this study to focus on the effect of link characteristics in retransmission performance.
+There is no attempt in this study to focus on the effect of link detailed characteristics in retransmission performance.
 
@@
 \end{tabular}
-\caption{Connection parameters}\label{tab1}
+\caption{Emulation parameters}\label{tab1}
 \end{center}
@@
 In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
-Losses generated by using associating netem with corresponding interfaces and dropping the packets. This process is simplified by using the KAUNetem tool~\cite{Garcia2016}. 
+Losses are generated by associating netem with corresponding interfaces and dropping select packets. This process is simplified by using the KAUNetem tool~\cite{Garcia2016}. 
 MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between client and server on the MPTCP protocol 
 support and interface availability. So one has to wait in time and packets for the second subflow establishment to understand the MPTCP retransmissions in multiple flows. We send two bursts of 80 
-packets each and drop tail packets on one of the interface. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that
-subflow. Three test cases are evaluated with one packet tail loss, two packet tail loss and one packet along with probe(in TLP) or retransmission(in RTO) loss.
+packets each and drop tail packets on one of the interfaces. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that
+subflow. Three test cases are evaluated with one packet tail loss, two packet tail loss and one packet along with probe~(in TLP) or retransmission~(in RTO) loss.
 
-In Linux, TCP retransmission features are controlled by a sysctl setting tcp.early.retrans. It has 5 possible values with ranging from 0 to 4 with 3 being Linux default, enables both
-ER and TLP. Standard TCP used by other operating systems might not use ER and TLP. We use RTO to refer sysctl setting of 0 and TLP to refer sysctl setting of 3 throughout the paper.
+In Linux, TCP retransmission features are controlled by a sysctl setting tcp.early.retrans.
+It has 5 possible values ranging from 0 to 4 with 3 being the Linux default.
+The default setting value enables both
+ER and TLP. TCP in other operating systems might not use ER and TLP. We use RTO to refer to a sysctl setting of 0 and TLP to refer to a sysctl setting of 3 throughout the paper.
 In Linux, disabling ER and TLP gives the performance of RTO. In Linux default TCP, enabling delayed ER gives the combined performance of RTO, delayed ER and TLP.
@@
 
-In each setting, we calculate burst completion time for two bursts of 80 packets each. The setup has server and client with Linux supporting MPTCP running on them. Client has two wireless 
-interfaces with one way delay on each interface ranging from 20ms to 120ms. We consider 5 scenarios with delay pairings 20ms-20ms, 20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand 
-the effect of delay difference in the performance. These scenarios are chosen to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in link delays.
+The setup has server and client with Linux supporting MPTCP running on them.
+Client has two wireless interfaces with one way delay on each interface ranging from 20ms to 120ms.
+We consider 5 scenarios with delay pairings 20ms-20ms, 20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance as shown in Table~\ref{tab1}.
+These scenarios are chosen to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in link delays.
+For each experiment trail, we send two bursts of data from server to client, each burst containing 80 packets.
+First burst allows the MPTCP connection to be setup taking in to account that second flow is added after some time and packets.
+We calculate the burst completion time for second burst that has specified tail losses.
 
@@
 
-In the case of single packet tail loss, there is no difference in burst completion times of RTO and TLP settings. TLP implementation in Linux accounts
-for the delayed ACK and waits 200ms before sending probe. Hence retransmission timeout is the mode of loss recovery in both RTO and TLP. This can be avoided by tuning
+In the case of single packet tail loss, there is no difference in burst completion times of RTO and TLP settings. The retransmission
+timeout and the probe timeout are same as there is only one outstanding packet. This is a special case in TLP
+specification that adds 200ms to PTO accommodating the case of delayed ACK. TLP implementation in Linux accounts
+for the delayed ACK and waits 200ms before sending probe. The minimum of PTO and RTO is considered for PTO.
+
+Hence retransmission timeout is the mode of loss recovery in both RTO and TLP. This can be avoided by tuning
 the ACKs as discussed in~\cite{Rajiullah:2015}. In the case of two packet tail loss, TLP performs better than RTO as it employs a combination of TLP and ER to recover from losses. 
-If the probe or retransmission is lost, then TLP still performs better than RTO as it waits for one RTO and one probe timeout instead of two RTOs.
+If the probe or retransmission is lost, then TLP still performs better than RTO as it waits for one probe timeout and one RTO instead of two RTOs.
 
@@
 
-The expected retransmission behavior of MPTCP for standard TCP and Linux default settings is shown 
+The expected retransmission behavior of MPTCP for RTO and Linux default settings is shown 
 in~\ref{timing1P},~\ref{timing1PP} and ~\ref{timing2P}. The actual pattern might differ in cases,
@@
 The burst completion times on single packet tail loss calculated for each delay configuration and retransmission 
-setting is depicted in figure~\ref{1p}. There is no difference in performance of RTO and TLP. The retransmission
-timeout and the probe timeout are same as there is only one outstanding packet. This is a special case in TLP
-specification that adds 200ms to PTO accommodating the case of delayed ACK. The minimum of PTO and RTO is considered
-for PTO.
+setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO and TLP.
 
@@
 the performance of n-degree tail loss, we consider dropping the last two packets. 
-The results of two packet tail loss are shown in~\ref{2p}. In this case, TLP performs better than the other 
-two settings in all scenarios except when the packets lost on high delay interface. The delay difference is
-large enough that the retransmission on alternate path is faster than the probe on the primary path. It 
+The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the other 
+two settings in all scenarios except when the packets are lost on high delay interface. In this scenario, the delay difference is
+large enough that the retransmission of last packet on alternate path is faster than the probe on the primary path. It 
 is PTO time slower than the RTO setting.
@@
 
-To further study the performance of TLP in improving latency, we tried to drop the probe packet along with the last packet. In this case RTO
-setting results much lower burst completion times. Use of TLP did not improve the burst completion time as we saw in TCP for TLP. The reason
-lies in the way the retransmission and loss probe transmission carried out in MPTCP. Retransmission occurs after an RTO time on both paths.
-Even if retransmission fails on one path, the client receives the packet on the other path as shown in Fig~\ref{timing1P}. For TLP, in
-current implementation, the loss probe sent on the actual path of first transmission but not on both paths. This incomplete impartation of
-TLP from TCP to MPTCP led to the increase burst completion time by an RTO for MPTCP.
+To further study the performance of TLP in the case when there is longer path interruption on one path, we tried to drop the probe packet along with the last packet. In this scenario, the RTO
+setting results in much lower burst completion times. Use of TLP did not improve the burst completion time as we saw in TCP for TLP. The reason
+lies in the way the retransmission and loss probe transmission is carried out in MPTCP. Retransmission occurs after an RTO time on both paths.
+Even if the retransmission fails on one path, the client receives the packet on the other path as shown in Fig~\ref{timing1P}. For TLP, in
+current implementation, the loss probe is sent on the actual path of the first transmission but not on both paths. This incomplete impartation of
+TLP from TCP to MPTCP led to the increase in burst completion time by an RTO for MPTCP.
 
@@
 When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout.
-Our results show that current TLP implementation does not improve the performance in MPTCP and incur more latency.
+Our results show that the current TLP implementation does not improve the performance in MPTCP in cases when there is path interruption and incur more latency.
 
-Current implementation of TLP is at the TCP flow level.
-In the event of RTO timeout, the MPTCP retransmission happens with a re-injection in to scheduler along with sending lost packet on the same path.
-But in the Event of Probe timeout, the loss probe packet is being sent on the same path without being re-injected in to scheduler.
-We propose a modification the probe sending mechanism for MPTCP by re-injecting the probe packet in to the MPTCP scheduler in the event of PTO.
-As the probe timeout is meant for acting faster than RTO timeout with essentially same consequence of retransmission, it is relevant to follow the process upon an RTO.
+The current implementation of TLP is at the TCP flow level.
+In the event of RTO timeout, the MPTCP retransmission happens with a re-injection in to the scheduler along with sending the lost packet on the same path.
+But in the event of Probe timeout, the loss probe packet is being sent on the same path without being re-injected in to the scheduler.
+We propose a modification to probe sending mechanism for MPTCP by re-injecting the probe packet in to the MPTCP scheduler in the event of PTO.
+As the probe timeout is meant for acting faster than the RTO timeout with essentially the same consequence as a retransmission, it is relevant to follow the same process as upon an RTO.
 It is expected that this modification is useful when the alternate path has lower delay than the packet original transmission path.
-The timing diagram depicted in figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP on each of the loss cases.
+The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied loss.
 
@@
 \end{center}
-\caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) One Packet b) Two Packet c) One packet and Retransmission}\label{timingNew}
+\caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) One packet b) Two packet c) One packet and Retransmission}\label{timingNew}
 \end{figure}
@@
 
-Results comparing the burst completion time for new MPTCP TLP with that of TCP and MPTCP TLP for the considered cases are provided in figures~\ref{1pn},~\ref{2pn} and ~\ref{1ppn}. 
-In figure~\ref{1pn}, mptcp-New performs better than mptcp in all scenarios, performs equal to tcp in symmetric scenarios and performs better than tcp in asymmetric scenarios.
-This improvement is due to the wait for RTO is eliminated for a packet to be retransmitted on the alternate path.
-In two packet tail loss scenario as in figure~\ref{2pn}, mptcp-New performs better than mptcp in all scenarios and better than tcp in all scenarios except when the alternate path has higher delay than the original path.
-In the single packet together with probe loss scenario as in figure~\ref{1ppn}, mptcp-New performs better than both mptcp and tcp in all scenarios with PTO triggering the probe retransmission on alternate path much earlier than the RTO.
+Results comparing the burst completion time for new MPTCP TLP with that of TCP and MPTCP TLP for the considered cases are provided in Figures~\ref{1pn},~\ref{2pn} and ~\ref{1ppn}. 
+In Figure~\ref{1pn}, MPTCP-New performs better than MPTCP in all scenarios, performs equal to TCP in symmetric scenarios and performs better than tcp in asymmetric scenarios.
+This improvement is due to RTO a packet to be retransmitted on the alternate path.
+In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New performs better than MPTCP in all scenarios and better than TCP in all scenarios except when the alternate path has higher delay than the original path.
+In the single packet together with probe loss scenario as in Figure~\ref{1ppn}, MPTCP-New performs better than both MPTCP and TCP in all scenarios with PTO triggering the probe retransmission on alternate path much earlier than the RTO.
 
@@
 \includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/1PNew.pdf}
-\caption{Single packet tail loss comparison with MPTCP-TLP}\label{1pn}
+\caption{Single packet tail loss comparison with MPTCP, TCP}\label{1pn}
 \end{center}
@@
 \includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/2PNew.pdf}
-\caption{Two packet tail loss comparison with MPTCP-TLP}\label{2pn}
+\caption{Two packet tail loss comparison with MPTCP, TCP}\label{2pn}
 \end{center}
@@
 \includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/1PPNew.pdf}
-\caption{Single packet tail loss together with probe loss with MPTCP-TLP}\label{1ppn}
+\caption{Single packet tail loss together with probe loss with MPTCP, TCP}\label{1ppn}
 \end{center}
@@
 \section{Conclusion}\label{conc}
-This paper provides the short comparative analysis of retransmission behavior with TCP and MPTCP protocols using possible standard loss patterns for tail losses.
+This paper provides the short comparative analysis of retransmission behavior with TCP and MPTCP protocols using various loss patterns for tail losses.
 The three cases of tail losses together with deterministic loss pattern allow us to dissect the retransmission behavior at packet level to understand possible improvements.
-We propose a less conservative approach to trigger retransmission on an alternate path by altering the loss indication as PTO timeout than the RTO timeout as in the Linux implementation.
-Our experiments on CORE emulator using  modified Linux implementation, prove that the proposed approach is in fact improves the burst completion time in most cases and equals the existing
+We propose a less conservative approach to trigger retransmission on an alternate path in the event of tail loss probe timeout, which otherwise triggered with an RTO. 
+Our emulation experiments  using a modified Linux implementation, show that the proposed approach in fact improves the burst completion time in most cases and equals the existing
 implementation in other cases.
-Proposed approach improve the performance when the alternate path has lower delay than the original loss path and the advantage increases with degree of asymmetry between the paths.
-This approach is very effective in case of probe loss which otherwise incur RTO timeout to trigger retransmission on either path.
+The proposed approach improves the performance when the alternate path has lower delay than the original loss path and the advantage increases with degree of asymmetry between the paths.
+Temporary path failures can cause probe loss along with packet loss.
+The approach is very effective in case of probe loss which otherwise incur RTO timeout to trigger retransmission on either path.
+However, the latency improvement comes at a cost of additional network traffic from the retransmitted packets.
+Further, we plan to investigate more conservative approaches for other retransmission events.
 

### 2017-03-08 main.tex
Update

@@
 
-Multipath TCP(MPTCP) is an experimental standard proposed as an extension to TCP~\cite{rfc6824}.
-It allows devices with multiple interfaces to transmit data simultaneously on both interfaces, there by improving the throughput of end to end connection.
+Multipath TCP~(MPTCP) is an experimental standard proposed as an extension to TCP~\cite{rfc6824}.
+It allows devices with multiple interfaces to transmit data simultaneously on both interfaces, thereby improving the throughput of end-to-end connections.
 Using MPTCP, a connection can have multiple TCP subflows using different interfaces on different routes.
@@
 
-TCP has mainly two mechanisms for detecting and recovering from packet losses namely Fast Retransmit (FR) and Retransmission Timeout~(RTO)~\cite{Flach:2013}.
+TCP has two mechanisms for detecting and recovering from packet losses: Fast Retransmit (FR) and Retransmission Timeout~(RTO)~\cite{Flach:2013}.
 Fast retransmit is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as a loss indication.
-In case of RTO, sender waits for the packet loss to trigger a timeout before retransmitting that packet.
-For long flows where there are sufficiently large number of packets to be transferred, FR is quicker than RTO in detecting and recovering packet losses.
+In case of RTO, a sender waits for the packet loss to trigger a timeout before retransmitting that packet.
+During timeout, the sender cannot send any packets, but with FR the out-of-order packets are presumed lost upon receiving duplicate acknowledgements.
+The sender can start retransmission without waiting for timeout.
+For flows where there are sufficiently many packets to be transferred, FR is quicker than RTO in detecting and recovering packet losses.
+Fast retransmit coupled with fast recovery can send data outstanding whene there is enough data to send in long flows.
 On the other hand, shorter flows have to depend on RTO for packet loss detection and recovery, thus incurring more latency.
 A single packet loss in a short flow may take many RTTs to detect and recover.
-This scenario is also applicable to the packets at the end of the flow (or tail) in a long flow.
-Though, FR and RTO are reliable even in tail loss recovery, they are not efficient in terms of recovery completion time.
+This scenario is also applicable to the packets at the end of the flow (or tail), in a long flow.
+Although, RTO is reliable, even in tail loss recovery, it is not efficient in terms of recovery completion time.
 
- MPTCP connection starts with one TCP flow to send data and with available network interfaces and protocol support, additional flows are added.
+ MPTCP connections start with one TCP flow to send data over, and may add more flows to the connection if configured to.
 MPTCP uses a connection level congestion window as well as subflow level congestion windows for each subflow.
@@
 Packet losses in each subflow are detected and recovered in a similar fashion as that of TCP.
-However, the MPTCP standard does not restrict on, how the loss recovery should happen in the implementation and which subflow should be retransmit the lost packets.
-Loss recovery can be conducted at different levels: the meta-level and the subflow level.
-If the recovery is handled at the meta level, the lost packet may be rescheduled and retransmitted at the available subflow with lowest RTT.
-If the recovery is handled at the flow level, the packet may be retransmitted in the same subflow.
-An important part of MPTCP packet loss recovery process is to follow the TCP semantics and must retransmit lost packet on the initial path of transfer.
+Loss recovery can be conducted at different levels: the MPTCP connection level (meta level) and at subflow level.
+If the recovery is handled at meta level, a lost packet may be rescheduled and retransmitted on the available subflow with the lowest RTT.
+If recovery is handled at the subflow level, the packet may be retransmitted on the same subflow.
+An important part of MPTCP packet loss recovery process is to follow the TCP semantics and retransmit lost packets on the initial path of transfer.
+However, the MPTCP standard does not restrict on, how the loss recovery should happen in the implementation and which subflow should retransmit the lost packets.
 
 TCP Loss Probe (TLP)~\cite{Flach:2013} is a mechanism to detect and recover from tail losses by employing a loss probe packet for a timeout shorter than the RTO.
-Probe timeout enables the short flows to recover packet loss at the tail by triggering the fast retransmit.
+Probe timeout enables short flows to recover packet loss at the tail, by triggering FR.
 It assumes other algorithms such as Early Retransmit~(ER)\cite{rfc5827} and Forward Acknowledgement~(FACK)~\cite{FACK} threshold based recovery are enabled.
@@
 We propose a change in the retransmission strategy of MPTCP to improve the latency performance.
-We provide the related research on MPTCP retransmission in section~\ref{relwork}, and experimental setup in section~\ref{exsetup}.
-We discuss the observations from the experiments in section~\ref{disc}, that lead to our proposed improvement on current Linux implementation of TLP for MPTCP in section~\ref{impr}.
-In section~\ref{eval}, we provide an evaluation of the proposed TLP for MPTCP implementation in Linux and conclude the paper in section~\ref{conc}.
+We provide related research on MPTCP retransmission in section~\ref{relwork}, and the experimental setup in section~\ref{exsetup}.
+We discuss observations from the experiments in section~\ref{disc}.
+Based on the observations, we propose an improvement to the current Linux implementation of TLP for MPTCP in section~\ref{impr}.
+In section~\ref{eval}, we provide an evaluation of the proposed TLP for MPTCP mechanism and conclude the paper in section~\ref{conc}.
 
@@
 
-Retransmission schemes of TCP are of significant importance in achieving lower flow completion time.
-Several research studies led to improved retransmission strategies for TCP.
-The current Linux TCP implementation provides space for improvement in specific cases of packet loss scenarios~\cite{prr}.
-In a TCP flow, sequence numbers are used to identify the lost packets and retransmit the respective packets.
-Multipath TCP uses two levels of sequence numbers to support efficient data transfer, namely data sequence and TCP sequence.
-Data sequence is for the end to end data transfer and TCP sequence number is for individual flow data sequence similar to TCP.
-A loss occurring in an individual TCP flow corresponds to a loss in end to end data.
+Retransmission schemes are of significant importance in achieving low flow completion time.
+Several research studies has led to improved retransmission strategies for TCP.
+The current Linux TCP implementation provides room for improvement in several packet loss scenarios.
+
+In a TCP flow, sequence numbers are used to identify lost packets and retransmit them.
+MPTCP uses two levels of sequence numbers to support efficient data transfer, namely data sequence and TCP sequence.
+Data sequence is for the end-to-end data transfer and TCP sequence number is for individual flow data sequence similar to TCP.
+There is an association between data sequence number and TCP sequence number within a subflow.
+A loss occurring in an individual TCP flow corresponds to a loss in end-to-end data.
 Multipath TCP, as an extension of TCP, uses the basics of TCP retransmission.
@@
 The dual sequence numbering enables MPTCP to respond to loss of packets by retransmitting the lost packets on an alternate path.
-It allows a packet to be retransmitted on a different subflow than that of the originally sent flow.
 
-The Linux implementation of MPTCP uses a set of retransmission heuristics to decide on the retransmissions.
+The Linux implementation of MPTCP uses a set of retransmission heuristics to handle retransmissions.
 The data outstanding on a timed out subflow should be rescheduled for transmission on a different subflow using timeout as the indicator.
@@
 The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 4G vs WLAN, leading to flows with different delay characteristics.
-Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fast paths.
+Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fastest paths.
 Quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always to follow TCP semantics.
-MPTCP standard suggests more conservative approaches.
-MPTCP is designed to exploit the multiple interfaces and often one or more of these interfaces could be wireless with completely different characteristics than that of wired interfaces.
+MPTCP standard suggests a more conservative approach.
+MPTCP is designed to exploit the multiple interfaces and often one or more of these could be wireless with completely different characteristics than that of wired interfaces.
 Authors in~\cite{Shin} argue that the calculation RTO for MPTCP flows should include the interface characteristics to improve loss recovery time.
@@
 For understanding the loss recovery and retransmission policies in MPTCP Linux implementation version 0.91, we consider several specific cases of tail losses.  
-No tuning of TCP settings such as, adjusting the delayed ACK behavior are performed.
-We consider the standard settings of the Linux kernel.
-We consider mainly various cases of tail loss and its recovery.
+We use the standard settings of the Linux kernel to understand the performance for various cases of tail loss and its recovery.
 The Linux implementation of TCP is considered as the baseline TCP retransmission for implementation flexibility and availability of TLP support.
@@
 
-Experiments use the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}.
+Experiments use the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}
 with client node connected to two wireless interfaces 3G or 4G and WLAN and server node connected to a wired router.
+CORE emulates the Linux networking stack in each node of the topology by using lightweight virtualization technique known as Linux network namespaces.
 Characteristics of the connection setup and assumptions about the parameters are provided in Table~\ref{tab1}.
@@
 In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
-Losses are generated by associating netem with corresponding interfaces and dropping select packets. This process is simplified by using the KAUNetem tool~\cite{Garcia2016}. 
-MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between client and server on the MPTCP protocol 
-support and interface availability. So one has to wait in time and packets for the second subflow establishment to understand the MPTCP retransmissions in multiple flows. We send two bursts of 80 
-packets each and drop tail packets on one of the interfaces. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that
+Losses are generated by associating netem with corresponding interfaces and dropping select packets.
+This process is simplified by using the KAUNetem tool~\cite{Garcia2016}, which allows experimenter to e.g., drop packets based on their positions in a flow.
+MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between the client and the server.
+So one has to wait both in time and packets for the second subflow establishment.
+We send two bursts of 80 packets each and drop tail packets at one of the interfaces.
+This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that
 subflow. Three test cases are evaluated with one packet tail loss, two packet tail loss and one packet along with probe~(in TLP) or retransmission~(in RTO) loss.

### 2017-03-09 main.tex
Update after Per comments

@@
 Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fastest paths.
-Quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always to follow TCP semantics.
+Such quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always to follow TCP semantics.
 MPTCP standard suggests a more conservative approach.
@@
 
-In Linux, TCP retransmission features are controlled by a sysctl setting tcp.early.retrans.
-It has 5 possible values ranging from 0 to 4 with 3 being the Linux default.
-The default setting value enables both
-ER and TLP. TCP in other operating systems might not use ER and TLP. We use RTO to refer to a sysctl setting of 0 and TLP to refer to a sysctl setting of 3 throughout the paper.
-In Linux, disabling ER and TLP gives the performance of RTO. In Linux default TCP, enabling delayed ER gives the combined performance of RTO, delayed ER and TLP.
-In RTO, probe loss scenario is equivalent to losing the retransmitted packet.
+In Linux, TCP supports multiple retransmission mechanisms that can be controlled by a system setting. The default setting in Linux enables both
+ER and TLP, which may not be the case with other operating systems. In this paper, we use TLP to denote the default Linux setting and RTO to denote the use of retransmission timeout only.
 
@@
 Client has two wireless interfaces with one way delay on each interface ranging from 20ms to 120ms.
+The intention is to evaluate the effect of symmetry, mild asymmetry and higher aysmmetry in the link delays.
 We consider 5 scenarios with delay pairings 20ms-20ms, 20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance as shown in Table~\ref{tab1}.
-These scenarios are chosen to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in link delays.
 For each experiment trail, we send two bursts of data from server to client, each burst containing 80 packets.
@@
 
-Experiments performed on CORE emulator with the setup mentioned in section~\ref{exsetup}.
-MPTCP is sensitive to path asymmetry in general due to the default scheduler being shortest RTT based scheduler.
-
-The burst completion times on single packet tail loss calculated for each delay configuration and retransmission 
-setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO and TLP.
+MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit
+the data until its congestion window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance
+is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
+configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO and TLP.
 

### 2017-03-13 main.tex
Update after Anna Comments - Round2

@@
 \usepackage[nocompress]{cite}
+\usepackage{caption}
+\usepackage{subcaption}
+
+%\newcommand{\per}[1]{}
+%\newcommand{\anna}[1]{}
+%\newcommand{\kiran}[1]{}
 
@@
 In case of RTO, a sender waits for the packet loss to trigger a timeout before retransmitting that packet.
-During timeout, the sender cannot send any packets, but with FR the out-of-order packets are presumed lost upon receiving duplicate acknowledgements.
-The sender can start retransmission without waiting for timeout.
+During timeout, the sender cannot send any packets, but with FR the out-of-order packets trigger duplicate acknowledgements.
+The sender can start retransmission of missing packets without waiting for timeout.
 For flows where there are sufficiently many packets to be transferred, FR is quicker than RTO in detecting and recovering packet losses.
 Fast retransmit coupled with fast recovery can send data outstanding whene there is enough data to send in long flows.
-On the other hand, shorter flows have to depend on RTO for packet loss detection and recovery, thus incurring more latency.
+On the other hand, shorter flows more frequently depend on RTO for packet loss detection and recovery, thus incurring more latency.
 A single packet loss in a short flow may take many RTTs to detect and recover.
 This scenario is also applicable to the packets at the end of the flow (or tail), in a long flow.
-Although, RTO is reliable, even in tail loss recovery, it is not efficient in terms of recovery completion time.
+Although, RTO is reliable, even in tail loss~(losses at the end of transmission) recovery, it is not efficient in terms of recovery completion time.
+
+TCP Loss Probe (TLP)~\cite{Flach:2013} is a mechanism to detect and recover from tail losses by employing a loss probe packet for a timeout shorter than the RTO.
+Probe timeout enables short flows to recover packet loss at the tail, by triggering FR.
+It assumes other algorithms such as Early Retransmit~(ER)\cite{rfc5827} and Forward Acknowledgement~(FACK)~\cite{FACK} threshold based recovery are enabled.
+TLP is available in Linux kernel and enabled by default.
+Further evaluations in~\cite{Rajiullah:2015}, showed that TLP provides significant reductions in latency for short flows with one to n-degree tail losses.
 
@@
 If the recovery is handled at meta level, a lost packet may be rescheduled and retransmitted on the available subflow with the lowest RTT.
-If recovery is handled at the subflow level, the packet may be retransmitted on the same subflow.
-An important part of MPTCP packet loss recovery process is to follow the TCP semantics and retransmit lost packets on the initial path of transfer.
-However, the MPTCP standard does not restrict on, how the loss recovery should happen in the implementation and which subflow should retransmit the lost packets.
-
-TCP Loss Probe (TLP)~\cite{Flach:2013} is a mechanism to detect and recover from tail losses by employing a loss probe packet for a timeout shorter than the RTO.
-Probe timeout enables short flows to recover packet loss at the tail, by triggering FR.
-It assumes other algorithms such as Early Retransmit~(ER)\cite{rfc5827} and Forward Acknowledgement~(FACK)~\cite{FACK} threshold based recovery are enabled.
-TLP is available in Linux kernel and enabled by default.
-Further evaluations in~\cite{Rajiullah:2015}, showed that TLP provides significant reductions in latency for short flows with one to n-degree tail losses.
+If recovery is handled at the subflow level, the packet must be retransmitted on the same subflow.
+An important part of loss recovery process at subflow level is to follow the TCP semantics and retransmit lost packets on the initial path of transfer.
+However, the MPTCP standard~\cite{rfc6824} does not restrict on, how the loss recovery should happen in the implementation and which subflow should retransmit the lost packets.
 
@@
 We propose a change in the retransmission strategy of MPTCP to improve the latency performance.
-We provide related research on MPTCP retransmission in section~\ref{relwork}, and the experimental setup in section~\ref{exsetup}.
-We discuss observations from the experiments in section~\ref{disc}.
+We provide related research on MPTCP retransmission in section~\ref{relwork}, and describe the experimental setup in section~\ref{exsetup}.
+We discuss observations from the experiments on state of the art retransmission behavior of TCP and MPTCP in section~\ref{disc}.
 Based on the observations, we propose an improvement to the current Linux implementation of TLP for MPTCP in section~\ref{impr}.
@@
 MPTCP uses two levels of sequence numbers to support efficient data transfer, namely data sequence and TCP sequence.
-Data sequence is for the end-to-end data transfer and TCP sequence number is for individual flow data sequence similar to TCP.
-There is an association between data sequence number and TCP sequence number within a subflow.
+Data sequence numbers are for the end-to-end data transfer and TCP sequence numbers are for individual flow data sequence.
+There is an association between data sequence numbers and TCP sequence numbers within a subflow.
 A loss occurring in an individual TCP flow corresponds to a loss in end-to-end data.
@@
 Such quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always to follow TCP semantics.
-MPTCP standard suggests a more conservative approach.
+MPTCP standard~\cite{rfc6824} suggests a more conservative approach.
 MPTCP is designed to exploit the multiple interfaces and often one or more of these could be wireless with completely different characteristics than that of wired interfaces.
-Authors in~\cite{Shin} argue that the calculation RTO for MPTCP flows should include the interface characteristics to improve loss recovery time.
+Authors in~\cite{Shin} argue that the calculation of RTO for MPTCP flows should include the interface characteristics to improve loss recovery time.
 
@@
 Experiments use the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}
-with client node connected to two wireless interfaces 3G or 4G and WLAN and server node connected to a wired router.
-CORE emulates the Linux networking stack in each node of the topology by using lightweight virtualization technique known as Linux network namespaces.
+with the client node connected to two wireless interfaces 3G or 4G and WLAN and the server node connected to a wired router.
+CORE uses the Linux networking stack in each node of the topology by using lightweight virtualization technique known as Linux network namespaces.
 Characteristics of the connection setup and assumptions about the parameters are provided in Table~\ref{tab1}.
@@
 So one has to wait both in time and packets for the second subflow establishment.
-We send two bursts of 80 packets each and drop tail packets at one of the interfaces.
-This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that
-subflow. Three test cases are evaluated with one packet tail loss, two packet tail loss and one packet along with probe~(in TLP) or retransmission~(in RTO) loss.
+We send two bursts of 80 packets each and drop tail packets in the second burst at one of the interfaces.
+This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that subflow.
+We calculate the burst completion time for second burst that has specified tail losses.
+Three test cases are evaluated with one packet tail loss, two packet tail loss and one packet along with probe~(in TLP) or retransmission~(in RTO) loss.
 
@@
 We consider 5 scenarios with delay pairings 20ms-20ms, 20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance as shown in Table~\ref{tab1}.
-For each experiment trail, we send two bursts of data from server to client, each burst containing 80 packets.
-First burst allows the MPTCP connection to be setup taking in to account that second flow is added after some time and packets.
-We calculate the burst completion time for second burst that has specified tail losses.
 
@@
 Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
-starts with the results using TCP and compare it with that of the MPTCP. Figures ~\ref{t1p},~\ref{t2p},~\ref{t1pp} represent the comparison of TCP 
-burst completion times. Path asymmetry is irrelevant in this case as the TCP uses single path.
+starts with the results using TCP and compare it with that of MPTCP. Figures ~\ref{t1p},~\ref{t2p},~\ref{t1pp} represent the comparison of TCP 
+burst completion times. Path asymmetry is irrelevant in this case as TCP uses a single path.
 
 In the case of single packet tail loss, there is no difference in burst completion times of RTO and TLP settings. The retransmission
-timeout and the probe timeout are same as there is only one outstanding packet. This is a special case in TLP
-specification that adds 200ms to PTO accommodating the case of delayed ACK. TLP implementation in Linux accounts
-for the delayed ACK and waits 200ms before sending probe. The minimum of PTO and RTO is considered for PTO.
+timeout and the probe timeout are same as there is only one outstanding packet. This is a special case in the TLP
+specification that adds 200~ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts
+for the delayed ACK and waits 200~ms before sending a probe. The minimum of PTO and RTO is considered for PTO.
 
@@
 the ACKs as discussed in~\cite{Rajiullah:2015}. In the case of two packet tail loss, TLP performs better than RTO as it employs a combination of TLP and ER to recover from losses. 
-If the probe or retransmission is lost, then TLP still performs better than RTO as it waits for one probe timeout and one RTO instead of two RTOs.
+If the probe or retransmission is lost, then TLP still performs better than RTO as it waits for one probe timeout and one RTO instead of two RTOs, thus avoiding the exponential backoff of the RTO.
 
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.46\textwidth,natwidth=578.16,natheight=433.62]{plots/T1P.pdf}
-\caption{Single packet tail loss using TCP}\label{t1p}
-\end{center}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.46\textwidth,natwidth=578.16,natheight=433.62]{plots/T2P.pdf}
-\caption{Two packet tail loss using TCP}\label{t2p}
-\end{center}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
-\caption{Single packet tail loss together with retransmission loss using TCP}\label{t1pp}
-\end{center}
-\end{figure}
+\begin{figure*}[!tbp]
+ \begin{subfigure}[b]{0.32\textwidth}
+	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T1P.pdf}
+	\caption{Single packet}\label{t1p}
+ \end{subfigure}
+ \hfill
+ \begin{subfigure}[b]{0.32\textwidth}
+	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T2P.pdf}
+	\caption{Two packet }\label{t2p}
+ \end{subfigure}
+ \hfill
+ \begin{subfigure}[b]{0.32\textwidth}
+	\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
+	\caption{Single packet and retransmission }\label{t1pp}
+ \end{subfigure}
+ \caption{Tail loss scenarios using TCP}
+\end{figure*}
 
@@
 The expected retransmission behavior of MPTCP for RTO and Linux default settings is shown 
-in~\ref{timing1P},~\ref{timing1PP} and ~\ref{timing2P}. The actual pattern might differ in cases,
-where the RTT is larger than 200ms.
+in Figures.~\ref{timing1P},~\ref{timing1PP} and ~\ref{timing2P}. The actual pattern might differ in cases,
+where the RTT is larger than 200~ms.
 
@@
 \begin{center}
+\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timing2P.pdf}
+\end{center}
+\caption{Timing diagram of MPTCP behavior with two packet tail loss a) RTO b) TLP}\label{timing2P}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
 \includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timing1PP.pdf}
@@
 
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timing2P.pdf}
-\end{center}
-\caption{Timing diagram of MPTCP behavior with two packet tail loss a) RTO b) TLP}\label{timing2P}
-\end{figure}
-
 MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit
@@
 is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
-configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO and TLP.
+configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO and TLP for 
+cases when the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
+scenarios~(30ms-20ms,120ms-20ms), the RTO retransmission happens in the second path with lowest RTT following the behavior depicted
+in Figure~\ref{timing1P}.
 
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.46\textwidth,natwidth=578.16,natheight=433.62]{plots/1P.pdf}
-\caption{Single packet tail loss using MPTCP}\label{1p}
-\end{center}
-\end{figure}
+\begin{figure*}[!tbp]
+ \begin{subfigure}[b]{0.32\textwidth}
+\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/1P.pdf}
+\caption{Single packet}\label{1p}
+ \end{subfigure}
+ \hfill
+ \begin{subfigure}[b]{0.32\textwidth}
+\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/2P.pdf}
+\caption{Two packet}\label{2p}
+ \end{subfigure}
+ \hfill
+ \begin{subfigure}[b]{0.32\textwidth}
+\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/1PP.pdf}
+\caption{Single packet and probe loss}\label{1pp}
+ \end{subfigure}
+ \caption{Tail loss scenarios using MPTCP}
+\end{figure*}
 
-In the case of single packet the last outstanding packet is sent as probe and in case of n-degree tail loss,
+In the case of single packet tail loss, the last outstanding packet is sent as probe and in case of n-degree tail loss,
 the latest packet is sent as probe to trigger fast retransmit of the packets in between. In order to observe
@@
 The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the other 
-two settings in all scenarios except when the packets are lost on high delay interface. In this scenario, the delay difference is
-large enough that the retransmission of last packet on alternate path is faster than the probe on the primary path. It 
-is PTO time slower than the RTO setting.
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.46\textwidth,natwidth=578.16,natheight=433.62]{plots/2P.pdf}
-\caption{Two packet tail loss using MPTCP}\label{2p}
-\end{center}
-\end{figure}
+two settings in all scenarios except when the packets are lost on high delay interface. In this scenario, the MPTCP retransmission follows the 
+behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of last packet on alternate path is 
+faster than the probe on the primary path.
 
 To further study the performance of TLP in the case when there is longer path interruption on one path, we tried to drop the probe packet along with the last packet. In this scenario, the RTO
-setting results in much lower burst completion times. Use of TLP did not improve the burst completion time as we saw in TCP for TLP. The reason
-lies in the way the retransmission and loss probe transmission is carried out in MPTCP. Retransmission occurs after an RTO time on both paths.
-Even if the retransmission fails on one path, the client receives the packet on the other path as shown in Fig~\ref{timing1P}. For TLP, in
+setting results in much lower burst completion times.
+Comparing with the performance of TLP in TCP from Figure~\ref{t1pp}, where we saw that use of TLP improved the burst completion time, TLP in MPTCP did not improve the performance.
+The reason lies in the way the retransmission and loss probe transmission is carried out in MPTCP.
+Retransmission occurs after an RTO time on both paths.
+Even if the retransmission fails on one path, the client receives the packet on the other path as shown in Figure~\ref{timing1PP}. For TLP, in
 current implementation, the loss probe is sent on the actual path of the first transmission but not on both paths. This incomplete impartation of
 TLP from TCP to MPTCP led to the increase in burst completion time by an RTO for MPTCP.
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/1PP.pdf}
-\caption{Single packet tail loss together with probe loss using MPTCP}\label{1pp}
-\end{center}
-\end{figure}
 
@@
 \section{Improvements to TLP for MPTCP}\label{impr}
-The goal of Tail Loss Probe(TLP) is to reduce tail latency of short flows.
-It achieves this by converting retransmission timeouts (RTOs) occurring due to tail losses (losses at end of transactions) into fast recovery.
+The goal of Tail Loss Probe~(TLP) is to reduce tail latency of short flows.
+It achieves this by converting retransmission timeouts (RTOs) occurring due to tail losses into fast recovery.
 TLP transmits one packet in two round-trips when a connection is in Open state and isn't receiving any ACKs.
@@
 When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout.
-Our results show that the current TLP implementation does not improve the performance in MPTCP in cases when there is path interruption and incur more latency.
+Our results show that the current TLP implementation for MPTCP helps improving the latency performance in cases when the paths are symmetric or the first path~(for the packet) has lowest RTT.
+However, it does not improve the performance in cases when the first path is not the lowest RTT path or when there is longer path interruption.
 
@@
 As the probe timeout is meant for acting faster than the RTO timeout with essentially the same consequence as a retransmission, it is relevant to follow the same process as upon an RTO.
-It is expected that this modification is useful when the alternate path has lower delay than the packet original transmission path.
+It is expected that this modification is useful when the alternate path has lower delay than the packet original transmission path or when there is longer path interruption.
 The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied loss.
@@
 Results comparing the burst completion time for new MPTCP TLP with that of TCP and MPTCP TLP for the considered cases are provided in Figures~\ref{1pn},~\ref{2pn} and ~\ref{1ppn}.
-In Figure~\ref{1pn}, MPTCP-New performs better than MPTCP in all scenarios, performs equal to TCP in symmetric scenarios and performs better than tcp in asymmetric scenarios.
-This improvement is due to RTO a packet to be retransmitted on the alternate path.
-In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New performs better than MPTCP in all scenarios and better than TCP in all scenarios except when the alternate path has higher delay than the original path.
+In Figure~\ref{1pn}, MPTCP-New performs better than MPTCP in all scenarios, performs equal to TCP in symmetric scenarios and performs better than TCP in asymmetric scenarios.
+This improvement is for the reason that MPTCP-New allows the packet to be retransmitted on the alternate path much before the RTO.
+In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New performs better than TCP in all scenarios and better than MPTCP in all scenarios except when the alternate path has a significantly higher delay than the original path, where performance is the same.
 In the single packet together with probe loss scenario as in Figure~\ref{1ppn}, MPTCP-New performs better than both MPTCP and TCP in all scenarios with PTO triggering the probe retransmission on alternate path much earlier than the RTO.
@@
 \section{Conclusion}\label{conc}
-This paper provides the short comparative analysis of retransmission behavior with TCP and MPTCP protocols using various loss patterns for tail losses.
-The three cases of tail losses together with deterministic loss pattern allow us to dissect the retransmission behavior at packet level to understand possible improvements.
+This paper provides a short comparative analysis of retransmission behavior with TCP and MPTCP protocols using various loss patterns for tail loss.
+The three cases of tail loss together with deterministic loss patterns allow us to dissect the retransmission behavior at packet level to understand possible improvements.
 We propose a less conservative approach to trigger retransmission on an alternate path in the event of tail loss probe timeout, which otherwise triggered with an RTO. 
@@
 However, the latency improvement comes at a cost of additional network traffic from the retransmitted packets.
-Further, we plan to investigate more conservative approaches for other retransmission events.
+Further, we plan to investigate less conservative approaches for other retransmission events.
 

### 2017-03-13 main.tex
Update

@@
 In this paper we focus on the state of the art loss recovery mechanisms for TCP and Multipath TCP.
-We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms by calculating the burst completion time for each scenario.
+We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms.
 Based on the observations, we propose an enhancement to the tail loss recovery in Multipath TCP to improve the loss recovery time.
-Our experiment results show end to end latency performance improvement in controlled scenarios by reducing the burst completion time.
+Our experiment results show consistent end to end latency performance improvement in considered scenarios.
 \end{abstract}
@@
 For flows where there are sufficiently many packets to be transferred, FR is quicker than RTO in detecting and recovering packet losses.
-Fast retransmit coupled with fast recovery can send data outstanding whene there is enough data to send in long flows.
+Fast retransmit coupled with fast recovery can send data outstanding when there is enough data to send in long flows.
 On the other hand, shorter flows more frequently depend on RTO for packet loss detection and recovery, thus incurring more latency.
@@
 
+\kern-1em
 \section{Background and Related Work}\label{relwork}
@@
 Retransmission schemes are of significant importance in achieving low flow completion time.
-Several research studies has led to improved retransmission strategies for TCP.
-The current Linux TCP implementation provides room for improvement in several packet loss scenarios.
+Several IETF RFCs such as SACK based recovery~\cite{rfc6675}, Limited Transmit~\cite{rfc3042} and Early Retransmission~\cite{rfc5827} have led to improved retransmission strategies for TCP.
+TLP is part of the various deployable loss recovery mechanisms proposed in~\cite{Flach:2013}, suggesting that current Linux TCP implementation provides room for improvement in several packet loss scenarios.
 
@@
 The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 4G vs WLAN, leading to flows with different delay characteristics.
+Significant delay differences in available paths can cause receive buffer limitations, as the receiver has to wait for the data sent on slow path, while buffering the data on the fast path.
+In such cases, Opportunistic retransmission~\cite{Costin} improves the latency by using fast path to retransmit the data sent on an another path.
 Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fastest paths.
@@
 
+\kern-1em
 \section{Experimental Setup}\label{exsetup}
@@
 Characteristics of the connection setup and assumptions about the parameters are provided in Table~\ref{tab1}.
-The parameters are chosen to be close to real values of bandwidth and delay of respective technologies.
+%The parameters are chosen to be close to real values of bandwidth and delay of respective technologies.
 There is no attempt in this study to focus on the effect of link detailed characteristics in retransmission performance.
@@
 
-      \textbf{RTT} & \multicolumn{10}{c|}{20ms-120ms(3G/4G/WLAN)}  \\
+      \textbf{RTT} & \multicolumn{10}{c|}{20\,ms-120\,ms(3G/4G/WLAN)}  \\
       \hline 	
@@
 
+\kern-1em
 \subsection{Testing retransmission with deterministic loss patterns}
@@
 We calculate the burst completion time for second burst that has specified tail losses.
-Three test cases are evaluated with one packet tail loss, two packet tail loss and one packet along with probe~(in TLP) or retransmission~(in RTO) loss.
+The objective of this experiment is to evaluate retransmission mechanisms with tail loss (single packet loss and multiple tail loss) and temporary path loss.
+Three test cases are considered for experiments: single packet tail loss and two packet tail loss for tail loss evaluation and single packet along with probe~(in TLP) or retransmission~(in RTO) loss for path loss evaluation.
 
@@
 The setup has server and client with Linux supporting MPTCP running on them.
-Client has two wireless interfaces with one way delay on each interface ranging from 20ms to 120ms.
-The intention is to evaluate the effect of symmetry, mild asymmetry and higher aysmmetry in the link delays.
-We consider 5 scenarios with delay pairings 20ms-20ms, 20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance as shown in Table~\ref{tab1}.
+Client has two wireless interfaces with one way delay on each interface ranging from 20\,ms to 120\,ms.
+The intention is to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in the link delays.
+We consider 5 scenarios with delay pairings 20\,ms-20\,ms, 20\,ms-30\,ms, 20\,ms-120\,ms, 30\,ms-20\,ms, 120\,ms-20\,ms to understand the effect of delay difference in the performance as shown in Table~\ref{tab1}.
 
+\kern-1em
 \section{Observations and Discussion}\label{disc}
@@
 tail loss.
-
+\kern-5em
 \subsection{TCP}
@@
 timeout and the probe timeout are same as there is only one outstanding packet. This is a special case in the TLP
-specification that adds 200~ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts
-for the delayed ACK and waits 200~ms before sending a probe. The minimum of PTO and RTO is considered for PTO.
+specification that adds 200\,ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts
+for the delayed ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for PTO.
 
@@
 	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T2P.pdf}
-	\caption{Two packet }\label{t2p}
+	\caption{Two packets }\label{t2p}
  \end{subfigure}
@@
 	\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
-	\caption{Single packet and retransmission }\label{t1pp}
+	\caption{Single packet and probe loss }\label{t1pp}
  \end{subfigure}
@@
 \end{figure*}
-
+\kern-1em
 \subsection{MPTCP}
@@
 in Figures.~\ref{timing1P},~\ref{timing1PP} and ~\ref{timing2P}. The actual pattern might differ in cases,
-where the RTT is larger than 200~ms.
+where the RTT is larger than 200\,ms.
 
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timing1P.pdf}
-\end{center}
-\caption{Timing diagram of MPTCP behavior with one packet tail loss a) RTO b) TLP}\label{timing1P}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timing2P.pdf}
-\end{center}
-\caption{Timing diagram of MPTCP behavior with two packet tail loss a) RTO b) TLP}\label{timing2P}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timing1PP.pdf}
-\end{center}
-\caption{Timing diagram of MPTCP behavior with one packet and next probe loss a) RTO b) TLP}\label{timing1PP}
-\end{figure}
+\begin{figure*}[!tbp]
+ \begin{subfigure}[b]{0.32\textwidth}
+	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing1P.pdf}
+	\caption{Single packet}\label{timing1P}
+ \end{subfigure}
+ \hfill
+ \begin{subfigure}[b]{0.32\textwidth} 
+	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing2P.pdf}
+	\caption{Two packets }\label{timing2P}
+ \end{subfigure} 
+ \hfill
+ \begin{subfigure}[b]{0.32\textwidth}
+  	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing1PP.pdf}
+	\caption{Single packet and probe loss}\label{timing1PP}
+ \end{subfigure}
+ \caption{Timing diagrams of MPTCP behavior on tail losses with a) RTO b) TLP }	
+\end{figure*}
 
@@
 configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO and TLP for 
-cases when the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
-scenarios~(30ms-20ms,120ms-20ms), the RTO retransmission happens in the second path with lowest RTT following the behavior depicted
+cases when both paths are symmetric or the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
+scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens in the second path with lowest RTT following the behavior depicted
 in Figure~\ref{timing1P}.
@@
 \includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/2P.pdf}
-\caption{Two packet}\label{2p}
+\caption{Two packets}\label{2p}
  \end{subfigure}
@@
 
+\kern-1em
 \section{Improvements to TLP for MPTCP}\label{impr}
@@
 \begin{center}
-\includegraphics[angle=0, width=0.46\textwidth]{images/timingER3NewTLP.pdf}
+\includegraphics[angle=0, width=0.46\textwidth]{images/timingER3NewTLP1.pdf}
 \end{center}
-\caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) One packet b) Two packet c) One packet and Retransmission}\label{timingNew}
+\caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) Single packet b) Two packets c) Single packet and probe loss}\label{timingNew}
 \end{figure}
 
+\kern-1em
 \section{Evaluation of modified MPTCP TLP}\label{eval}
@@
 
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/1PNew.pdf}
-\caption{Single packet tail loss comparison with MPTCP, TCP}\label{1pn}
-\end{center}
-\end{figure}
+\begin{figure*}[!tbp]
+\begin{subfigure}[b]{0.32\textwidth}
+\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/1PNew.pdf}
+\caption{Single packet}\label{1pn}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.32\textwidth}
+\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/2PNew.pdf}
+\caption{Two packets}\label{2pn}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.32\textwidth}
+\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/1PPNew.pdf}
+\caption{Single packet and probe loss}\label{1ppn}
+\end{subfigure}
+\caption{Performance comparison of MPTCP-New with MPTCP and TCP}
+\end{figure*}
 
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/2PNew.pdf}
-\caption{Two packet tail loss comparison with MPTCP, TCP}\label{2pn}
-\end{center}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.46\textwidth, natwidth=578.16,natheight=433.62]{plots/1PPNew.pdf}
-\caption{Single packet tail loss together with probe loss with MPTCP, TCP}\label{1ppn}
-\end{center}
-\end{figure}
-
+\kern-1em
 \section{Conclusion}\label{conc}
@@
 
+\kern-1em
 \bibliographystyle{IEEEtran}

### 2017-03-13 plots/plot_tlp.py

@@ -1,5 +1,11 @@
 import numpy as np
 import matplotlib.pyplot as plt
+import matplotlib 
+
+matplotlib.rc('axes',labelsize=18)
+matplotlib.rc('xtick',labelsize=14)
+matplotlib.rc('ytick', labelsize=14)
+
 
 #data = np.loadtxt(sys.argv[1])
 
@@ -46,15 +52,15 @@ margin = 0.05
 bar_width = (1-2*margin)/3
 
 
-rects1 = plt.bar(index+margin, p2e0 , bar_width,
+rects1 = plt.bar(index+margin, p1e0 , bar_width,
                  color='b',
                  label='er=0')
 
-rects2 = plt.bar(index+margin + bar_width, p2e2 , bar_width,
+rects2 = plt.bar(index+margin + bar_width, p1e2 , bar_width,
                  color='g',
                  label='er=2')
 
-rects3 = plt.bar(index+margin + 2*bar_width, p2e3, bar_width,
+rects3 = plt.bar(index+margin + 2*bar_width, p1e3, bar_width,
                  color='r',
                  label='er=3')
 

### 2017-03-13 plots/plot_tlp_c.py

@@ -1,5 +1,12 @@
 import numpy as np
 import matplotlib.pyplot as plt
+import matplotlib
+#data = np.loadtxt(sys.argv[1])
+
+matplotlib.rc('axes',labelsize=18)
+matplotlib.rc('xtick',labelsize=14)
+matplotlib.rc('ytick', labelsize=14)
+
 
 #data = np.loadtxt(sys.argv[1])
 
@@ -33,11 +40,11 @@ margin = 0.05
 bar_width = (1-2*margin)/3
 
 
-rects1 = plt.bar(index+margin, p1pe0 , bar_width,
+rects1 = plt.bar(index+margin, p2e0 , bar_width,
                  color='b',
                  label='RTO')
 
-rects2 = plt.bar(index+margin + bar_width, p1pe3 , bar_width,
+rects2 = plt.bar(index+margin + bar_width, p2e3 , bar_width,
                  color='g',
                  label='TLP')
 

### 2017-03-13 plots/plot_tlp_new.py

@@ -1,7 +1,10 @@
 import numpy as np
 import matplotlib.pyplot as plt
+import matplotlib
 
-#data = np.loadtxt(sys.argv[1])
+matplotlib.rc('axes',labelsize=18)
+matplotlib.rc('xtick',labelsize=14)
+matplotlib.rc('ytick', labelsize=14)
 
 n_groups = 5 
 
@@ -28,15 +31,15 @@ margin = 0.05
 bar_width = (1-2*margin)/3
 
 
-rects1 = plt.bar(index+margin, p2e3 , bar_width,
+rects1 = plt.bar(index+margin, p1pe3 , bar_width,
                  color='b',
                  label='mptcp')
 
-rects2 = plt.bar(index+margin + bar_width, np2e3 , bar_width,
+rects2 = plt.bar(index+margin + bar_width, np1pe3 , bar_width,
                  color='g',
                  label='mptcp-New')
 
-rects3 = plt.bar(index+margin + 2*bar_width, tp2e3, bar_width,
+rects3 = plt.bar(index+margin + 2*bar_width, tp1pe3, bar_width,
                  color='r',
                  label='tcp')
 

### 2017-03-15 main.tex
Final update

@@
 Probe timeout enables short flows to recover packet loss at the tail, by triggering FR.
-It assumes other algorithms such as Early Retransmit~(ER)\cite{rfc5827} and Forward Acknowledgement~(FACK)~\cite{FACK} threshold based recovery are enabled.
+It assumes that other algorithms such as Early Retransmit~(ER)\cite{rfc5827} and Forward Acknowledgement~(FACK)~\cite{FACK} threshold based recovery are enabled.
 TLP is available in Linux kernel and enabled by default.
@@
 
-\kern-1em
+%\kern-1em
 \section{Background and Related Work}\label{relwork}
@@
 The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 4G vs WLAN, leading to flows with different delay characteristics.
-Significant delay differences in available paths can cause receive buffer limitations, as the receiver has to wait for the data sent on slow path, while buffering the data on the fast path.
-In such cases, Opportunistic retransmission~\cite{Costin} improves the latency by using fast path to retransmit the data sent on an another path.
+Significant delay differences in available paths can cause receive buffer limitations, as the receiver has to wait for the data sent on a slow path, while buffering the data on the fast path.
+In such cases, opportunistic retransmission~\cite{Costin} improves the latency by using fast path to retransmit the data sent on an another path.
 Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fastest paths.
-Such quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always to follow TCP semantics.
-MPTCP standard~\cite{rfc6824} suggests a more conservative approach.
+Such quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always as well to follow TCP semantics.
+The MPTCP standard~\cite{rfc6824} suggests a more conservative approach.
 MPTCP is designed to exploit the multiple interfaces and often one or more of these could be wireless with completely different characteristics than that of wired interfaces.
@@
 For understanding the loss recovery and retransmission policies in MPTCP Linux implementation version 0.91, we consider several specific cases of tail losses.  
-We use the standard settings of the Linux kernel to understand the performance for various cases of tail loss and its recovery.
+We use the standard settings of the Linux kernel.
 The Linux implementation of TCP is considered as the baseline TCP retransmission for implementation flexibility and availability of TLP support.
 
-\kern-1em
+%\kern-1em
 \section{Experimental Setup}\label{exsetup}
 
-Experiments use the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}
-with the client node connected to two wireless interfaces 3G or 4G and WLAN and the server node connected to a wired router.
-CORE uses the Linux networking stack in each node of the topology by using lightweight virtualization technique known as Linux network namespaces.
+Experiments use the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}.
+The client node connect to two wireless interfaces 3G or 4G and WLAN and the server node connect to a wired router.
+CORE uses the Linux networking stack in each node of the topology by using a lightweight virtualization technique known as Linux network namespaces.
 Characteristics of the connection setup and assumptions about the parameters are provided in Table~\ref{tab1}.
@@
 
-      \textbf{RTT} & \multicolumn{10}{c|}{20\,ms-120\,ms(3G/4G/WLAN)}  \\
+      \textbf{Oneway Delay} & \multicolumn{10}{c|}{20\,ms-120\,ms}  \\
       \hline 	
-      \textbf{Bandwidth} & \multicolumn{10}{c|}{54Mbps(3G/4G/WLAN)}  \\
+      \textbf{Bandwidth} & \multicolumn{10}{c|}{54Mbps}  \\
       \hline
@@
 
-\kern-1em
+%\kern-1em
 \subsection{Testing retransmission with deterministic loss patterns}
@@
 This process is simplified by using the KAUNetem tool~\cite{Garcia2016}, which allows experimenter to e.g., drop packets based on their positions in a flow.
-MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between the client and the server.
+A MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between the client and the server.
 So one has to wait both in time and packets for the second subflow establishment.
@@
 This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that subflow.
-We calculate the burst completion time for second burst that has specified tail losses.
+We calculate the burst completion time for the second burst that has specified tail losses.
 The objective of this experiment is to evaluate retransmission mechanisms with tail loss (single packet loss and multiple tail loss) and temporary path loss.
@@
 The setup has server and client with Linux supporting MPTCP running on them.
-Client has two wireless interfaces with one way delay on each interface ranging from 20\,ms to 120\,ms.
+The Client has two interfaces with one way delay on each interface ranging from 20\,ms to 120\,ms.
 The intention is to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in the link delays.
@@
 \section{Observations and Discussion}\label{disc}
-This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, retransmission or probe loss and two packet 
-tail loss.
+This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, two packet tail loss and retransmission or probe loss.
 \kern-5em
@@
 Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
-starts with the results using TCP and compare it with that of MPTCP. Figures ~\ref{t1p},~\ref{t2p},~\ref{t1pp} represent the comparison of TCP 
+starts with the results using TCP and compare it with that of MPTCP. Figure~\ref{tcpf} represent the comparison of TCP 
 burst completion times. Path asymmetry is irrelevant in this case as TCP uses a single path.
@@
 In the case of single packet tail loss, there is no difference in burst completion times of RTO and TLP settings. The retransmission
-timeout and the probe timeout are same as there is only one outstanding packet. This is a special case in the TLP
+timeout and the probe timeout are the same as there is only one outstanding packet. This is a special case in the TLP
 specification that adds 200\,ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts
@@
  \end{subfigure}
- \caption{Tail loss scenarios using TCP}
+ \caption{Tail loss scenarios using TCP}\label{tcpf}
 \end{figure*}
-\kern-1em
+%\kern-1em
 \subsection{MPTCP}
@@
 The expected retransmission behavior of MPTCP for RTO and Linux default settings is shown 
-in Figures.~\ref{timing1P},~\ref{timing1PP} and ~\ref{timing2P}. The actual pattern might differ in cases,
-where the RTT is larger than 200\,ms.
-
+in Figure.~\ref{mptiming}. The actual pattern might differ in cases, where the RTT is larger than 200\,ms. 
+\renewcommand{\thesubfigure}{\roman{subfigure}}
 \begin{figure*}[!tbp]
@@
  \end{subfigure}
- \caption{Timing diagrams of MPTCP behavior on tail losses with a) RTO b) TLP }	
+ \caption{Timing diagrams of MPTCP behavior on tail losses with a) RTO b) TLP }\label{mptiming}	
 \end{figure*}
@@
 
+\renewcommand{\thesubfigure}{\alph{subfigure}}
 \begin{figure*}[!tbp]
@@
 The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the other 
-two settings in all scenarios except when the packets are lost on high delay interface. In this scenario, the MPTCP retransmission follows the 
-behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of last packet on alternate path is 
+two settings in all scenarios except when the packets are lost on a high delay interface. In this scenario, the MPTCP retransmission follows the 
+behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of the last packet on the alternate path is 
 faster than the probe on the primary path.
@@
 
-\kern-1em
+%\kern-1em
 \section{Improvements to TLP for MPTCP}\label{impr}
@@
 Our results show that the current TLP implementation for MPTCP helps improving the latency performance in cases when the paths are symmetric or the first path~(for the packet) has lowest RTT.
-However, it does not improve the performance in cases when the first path is not the lowest RTT path or when there is longer path interruption.
+However, it reduces the performance in cases when the first path is not the lowest RTT path or when there is longer path interruption.
 
@@
 But in the event of Probe timeout, the loss probe packet is being sent on the same path without being re-injected in to the scheduler.
-We propose a modification to probe sending mechanism for MPTCP by re-injecting the probe packet in to the MPTCP scheduler in the event of PTO.
+We propose a modification to the probe sending mechanism for MPTCP by re-injecting the probe packet in to the MPTCP scheduler in the event of PTO.
 As the probe timeout is meant for acting faster than the RTO timeout with essentially the same consequence as a retransmission, it is relevant to follow the same process as upon an RTO.
 It is expected that this modification is useful when the alternate path has lower delay than the packet original transmission path or when there is longer path interruption.
-The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied loss.
+The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied loss scenarios.
 
@@
 
-\kern-1em
+%\kern-1em
 \section{Evaluation of modified MPTCP TLP}\label{eval}
 
-Results comparing the burst completion time for new MPTCP TLP with that of TCP and MPTCP TLP for the considered cases are provided in Figures~\ref{1pn},~\ref{2pn} and ~\ref{1ppn}.
-In Figure~\ref{1pn}, MPTCP-New performs better than MPTCP in all scenarios, performs equal to TCP in symmetric scenarios and performs better than TCP in asymmetric scenarios.
+Results comparing the burst completion time for new MPTCP TLP with that of TCP and MPTCP TLP for the considered cases are provided in Figure~\ref{mpnew}.
+In Figure~\ref{1pn}, MPTCP-New performs better than TCP in all scenarios, performs equal to MPTCP in symmetric scenarios and performs better than MPTCP in asymmetric scenarios with a lower delay on the alternate path.
 This improvement is for the reason that MPTCP-New allows the packet to be retransmitted on the alternate path much before the RTO.
@@
 \end{subfigure}
-\caption{Performance comparison of MPTCP-New with MPTCP and TCP}
+\caption{Performance comparison of MPTCP-New with MPTCP and TCP}\label{mpnew}
 \end{figure*}
 
-\kern-1em
+%\kern-1em
 \section{Conclusion}\label{conc}
@@
 
-\kern-1em
+%\kern-1em
 \bibliographystyle{IEEEtran}

### 2017-04-24 responses/reviewresponse.tex
Update

\documentclass[a4paper,twoside,11pt]{reviewresponse}

% 1. Load and set up proper language packages
\usepackage[utf8x]{inputenc}
%\usepackage[latin9]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

% 2. Complete the paper data
\newcommand{\myAuthors}{{Kiran~Yedugundla, Per~Hurtig, Anna~Brunstrom}}
\newcommand{\myAuthorsShort}{Kiran~et. al}
%\newcommand{\myEmail}{kiran.yedugundla@kau.se}
\newcommand{\myTitle}{Probe or Wait: Handling tail losses using Multipath TCP}
\newcommand{\myShortTitle}{Response to reviewers}
\newcommand{\myJournal}{IFIP Networking - FIT Workshop 2017}
\newcommand{\myDept}{{$^{\displaystyle 1}$Department of Computer Science, Karlstad University, Karlstad, Sweden. }}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[linktoc=all]{hyperref}
%\usepackage[linktoc=all,bookmarks,bookmarksopen=true,bookmarksnumbered=true]{hyperref}

\hypersetup{
pdfauthor = {\myAuthorsShort},
pdftitle = {\myTitle},
pdfsubject = {\myJournal},
colorlinks = true,
linkcolor=black!70!green,          % color of internal links
citecolor=black!70!green,        % color of links to bibliography
filecolor=magenta,      % color of file links
urlcolor=black!70!green           % color of external links
}

\newcommand{\ab}[1]{\textcolor{red}{{\sf AB: #1}}}

\begin{document}

\thispagestyle{plain}

\begin{center}
 {\LARGE\myTitle} \vspace{0.5cm} \\
 {\large\myShortTitle} \vspace{0.5cm}\\
 {\large\myJournal} \vspace{0.5cm} \\
 \today \vspace{0.5cm} \\
 \myAuthors \\
 %\url{\myEmail} \vspace{1cm} \\
 \small\myDept
\end{center}

%\tableofcontents

\begin{abstract}
We thank the anonymous reviewers for their efforts and valuable feedback. 
In this document, we propose changes to the draft in response to comments/suggestions received from reviewers.

\end{abstract}

\clearpage
\section{Reviewer 1}
\setcounter{reviewer} {1}
\rsum{
Paper Summary:
Authors evaluate the current use of tail loss probe (TLP) in the context of Linux MPTCP implementation, and propose a modification that improves the flow completion times of MPTCP connections when faced with tail losses.

Strengths: What are the major reasons to accept the paper? [Be  brief.]

The paper identifies a specific issue, and proposes a clear \& easy to comprehend enhancement to existing Linux MPTCP implementation of TLP. The results appear to be positive.

*** Weaknesses: What are the most important reasons NOT to accept the paper? [Be brief.]

The experimental analysis does not appear exhaustive (or if it was, this doesn't really come through in the presentation). But having said that, the analysis appears *sufficient* for this workshop's purposes.

 *** Detailed Comments: Please provide detailed comments that will be helpful to the TPC for assessing the paper. Also provide feedback to the authors.

The paper's writing is easy to read, their contribution is clear and succinct. From page 5: "We propose [...] re-injecting the probe packet in to the MPTCP scheduler in the event of PTO. [...] It is expected that this modification is useful when the alternate path has lower delay than the packet original transmission path or when there is longer path interruption."

This would seem a suitable paper for a workshop, although perhaps the impact of their proposed enhancement is not explored as much as we might wish. In particular, it would have been stimulating to get a bigger-picture impact of current MPTCP use of TLP on applications and their end-users. This would help put the paper's proposed enhancement into context (i.e. what degree of difference will it make in how the Internet works tomorrow?)
}

\textbf{Response}

We agree with the reviewer that paper is not comprehensive in studying the effect of the proposed mechanism for various traffic scenarios in the Internet. We considered that for our future research work and the corresponding text is updated.

\clearpage

\section{Reviewer 2}
\setcounter{reviewer} {2}
\rsum{
 *** Paper summary: Paper summary

This paper proposes a modified error recovery heuristic for Multipath TCP. The authors show that a small change of the tail loss probe handling can improve the completion time of data transfers compared to the existing algorithms used by the Linux Multipath TCP implementation. The benefits are demonstrated by a number of measurements using a simple workload and simple, static network scenarios.

*** Strengths: What are the major reasons to accept the paper? [Be brief.]

The paper is a short but solid study on a timely topic.

*** Weaknesses: What are the most important reasons NOT to accept the paper? [Be brief.]

For actually evaluating the proposed technique more comprehensive studies would be useful.
}
\rcomment{

*** Detailed Comments: Please provide detailed comments that will be helpful to the TPC for assessing the paper. Also provide feedback to the authors.

- The Section 1 and 2 are comparatively detailed and long compared to Section 5, which is the key contribution of the paper. A final version could expand more in Section 5, which is short and not easy to follow.

- The of the acronym "RTO" both for a timer and for a retransmission algorithm is a bit confusing

- In Figure 2, is there a reason for the small increase for TLP in Subfigure (b) 120ms?

- Figure 3 and 5 are not easy to parse and could be improved:

  * In general, as these figures are essential for understanding the paper, the visualization could be expanded to better highlight differences and possibly suboptimal behavior.

  * For instance, the sequence chart in Figure 3 and (in particular) 5 seem to show the case with symmetric delays (same delivery delay on both paths). It could be interesting to have a look at the sequence charts for asymmetric delays.

  * The terms "Cli-l1" and "Cli-l2" are not intuitive.

  * If the retransmission timeout and the tail-loss probe timeout have different values, this is also not visualized in the sequence chart.

- Section 5 could discuss more comprehensively what could be changed in "MPTCP-New". For instance, couldn't it make sense to use different heuristics depending on whether a loss is on the path with lowest RTT, or not?

- If a retransmission scheme increases network load, this requires careful analysis. For instance, can it happen that the load on a congested path can increase? This should be discussed in more detail. In general, the analysis could be more comprehensive.

- The final paper should be proof-read as there are some minor editorial issues.

}
\clearpage
\textbf{Response}

We plan to address the following points from the comments:\\

1. Expand Section 5 to elaborate the contribution of the paper to enhance readability.\\
2. RTO is used mainly to suggest that it is the standard (not Linux) TCP retransmission policy. RTO is used to elevate the importance of RTO timer in the retransmission policy.\\
3. In the case of Figure 2 b), PTO is very close to RTO and the second packet is recovered with ER when TLP is used unlike resending all unacked packets on RTO. \\
4. Due to the page limit constraints, it is not feasible to provide the figures for asymmetric delays. However the current figures could be updated to enhance the visualization. \\
   Cli-I1 is Client Interface 1. It could be changed to Inf1-Client. We have simplified the timing figures to make it clean. \\
5. All transmissions are limited by congestion window and that covers the increased load scenario. We consider the suggestion for our future research.\\
6. We try to enhance the readability by careful proof reading.\\

\clearpage
\section{Reviewer 3}
\setcounter{reviewer} {3}

\rsum{

 *** Paper summary: Paper summary

The paper investigates LTP for MCTCP. The authors investigate 3 scenarios for TCP, as baseline, and MPTCP both with and without TLP using emulations. Their results show a weakness of the existing TLP implementation for MPTCP in a scenario where the TLP is also lost on the link with higher delay. Further the authors propose a change to address this detected problem and show the improvements with their modification in the same scenarios.

*** Strengths: What are the major reasons to accept the paper? [Be brief.]

The experiment is well performed and the paper is well written. Further the authors propose an change in the implementation that demonstrates to improve the results.

*** Weaknesses: What are the most important reasons NOT to accept the paper? [Be brief.]

The paper investigates an existing implementation of TLP for MPTCP. While the results are interesting, the investigated scenarios are limited, catching a problem in a potentially rare case. The proposed improvement rather seems like fixing an incomplete implementation of TLP for MPTCP. While this proposed changed is a valuable contribution, it's not clear to me if this contribution is sufficient for a 6-page paper.

}

\rcomment{
*** Detailed Comments: Please provide detailed comments that will be helpful to the TPC for assessing the paper. Also provide feedback to the authors.

Thanks for the well-written paper. One minor comment: Maybe use different colors for figure 6 because I first assume that blue and green would match to some of the previous graphs. Also it is not really clearly described anywhere that all results in figure 6 are with the TLP version of TCP and MPTCP respectively.

Tiny nit:
s/Further evaluations in [5], showed/Further, evaluations in [5] showed/

}
\textbf{Response}
We plan to address the following points from the comments:\\
1. Figure 6 Colors will be updated as per suggestion. \\
2. We update the text in Section 6 to elucidate that the results in Figure 6 are with the TLP enabled TCP and MPTCP respectively.\\
3. Text updated.\\

\clearpage
\section{Reviewer 4} 
\setcounter{reviewer} {4}
\rsum{

*** Paper summary: Paper summary

The paper studies tail losses in TCP and Multipath TCP, and presents a modification to the existing algorithm to enable faster burst completion times.

*** Strengths: What are the major reasons to accept the paper? [Be  brief.]

Interesting and novel work that is timely as MPTCP is an emerging transport protocol. The evaluation is quite nice.
}

\rcomment{
*** Weaknesses: What are the most important reasons NOT to accept the  paper? [Be brief.]

Overall the weakness are more in the line of clarifications and positioning:
- How big an issue tail loss is in reality? Is it really a problem, what data is there to show it is a problem worth solving? In particular since you mentioned some downsides of TLP (are there others?).
- Introduction doesn't mention results, btw.
- You have two times asymmetric use cases, e.g., 20-30 and 30-20, but I didn't understand why? Why aren't those the same if the bit rates on the links are the same (as per Table 1)? Or are the BDP different on the links?
- Figure 2 is not really discussed, in particular why (b) shows that TLP is worse than RTO?
- How does TLP work really? How does it know the transfer is ending and not in the middle? You don't really explain how TLP is triggered vs. normal retransmissions. Does TLP affect "normal" data transfers at all?

*** Detailed Comments: Please provide detailed comments that will be helpful to the TPC for assessing the paper. Also provide feedback to the authors.

Overall it is a nice paper that would benefit from several clarifications. The weaknesses above would be important to tackle.

}

\textbf{Response}

We plan to address the following points from the comments:\\
1. Introduction section update to include importance of tail loss, the problem and gist of results.\\
2. Path asymmetry is important as the paper considers deterministic packet loss in primary path. We will try and clarify the difference between the scenarios in the text.\\
3. Figure 2 is discussed in Section 4.1 and we update the text to clarify the question on figure 2 (b).\\
4. We try to update the Section 2 to provide more clarity on the working details of TLP. \\

% Uncomment in case references are needed
\bibliographystyle{apalike}
\bibliography{responsereferences}

\end{document}

### 2017-04-24 main.tex
Update

@@
 TLP is available in Linux kernel and enabled by default.
-Further evaluations in~\cite{Rajiullah:2015}, showed that TLP provides significant reductions in latency for short flows with one to n-degree tail losses.
+Further, evaluations in~\cite{Rajiullah:2015} showed that TLP provides significant reductions in latency for short flows with one to n-degree tail losses.
 

### 2017-04-24 responses/reviewresponse.tex
Update

@@
 
-We agree with the reviewer that paper is not comprehensive in studying the effect of the proposed mechanism for various traffic scenarios in the Internet.
-We considered that for our future research work and the corresponding text is updated.
+We agree with the reviewer that paper is not comprehensive (due to page limitations) in studying the effect of the proposed mechanism for various traffic scenarios in the Internet.
+We considered a comprehensive study for our future research work and we try to clarify that in the Conclusions section.
 
@@
 
-1. Expand Section 5 to elaborate the contribution of the paper to enhance readability.\\
-2. RTO is used mainly to suggest that it is the standard (not Linux) TCP retransmission policy. RTO is used to elevate the importance of RTO timer in the retransmission policy.\\
-3. In the case of Figure 2 b), PTO is very close to RTO and the second packet is recovered with ER when TLP is used unlike resending all unacked packets on RTO. \\
+1. Expanded Section 5 to elaborate the contribution of the paper to enhance readability.\\
+2. We updated the text in Section 4 by clearly mentioning RTO timer and RTO retransmission policy to enhance readability.\\
+3.
+In the case of Figure 2 b), PTO is very close to RTO and the second packet is recovered with ER which has a small delay when TLP is used, unlike resending all unacked packets on RTO expiration.
+Corresponding explanation is included in Section 4.1 \\
 4. Due to the page limit constraints, it is not feasible to provide the figures for asymmetric delays. However the current figures could be updated to enhance the visualization. \\
-   Cli-I1 is Client Interface 1. It could be changed to Inf1-Client. We have simplified the timing figures to make it clean. \\
-5. All transmissions are limited by congestion window and that covers the increased load scenario. We consider the suggestion for our future research.\\
-6. We try to enhance the readability by careful proof reading.\\
+   Cli-I1 is Client Interface 1. It could be changed to Client(I1). We have simplified the timing figures to make it clean. \\
+5. MPTCP-New has currently no heuristics. We clarify that in Section 5.
+6.
+In relation to increased load, all transmissions are still limited by the cwnd and the overhead does not create congestion.
+We consider the suggestion on comprehensive analysis for our future research.\\
+7. We try to enhance the readability by careful proof reading.\\
 
@@
 We plan to address the following points from the comments:\\
-1. Introduction section update to include importance of tail loss, the problem and gist of results.\\
+1. Introduction section is updated to include importance of addressing the problem of tail losses in the network traffic and gist of results.\\
 2. Path asymmetry is important as the paper considers deterministic packet loss in primary path. We will try and clarify the difference between the scenarios in the text.\\

### 2017-04-24 main.tex
Update

@@
 Although, RTO is reliable, even in tail loss~(losses at the end of transmission) recovery, it is not efficient in terms of recovery completion time.
+Tail losses are one of the major causes of end-to-end latency detoriation in modern networks~\cite{Flach:2013}.
 
@@
 In section~\ref{eval}, we provide an evaluation of the proposed TLP for MPTCP mechanism and conclude the paper in section~\ref{conc}.
+Our experimental results indicate improvement in the performance of TLP by reducing recovery time in the event of tail losses.
 
@@
 
-In the case of single packet tail loss, there is no difference in burst completion times of RTO and TLP settings. The retransmission
+In the case of single packet tail loss as shown in Figure~\ref{t1p}, there is no difference in burst completion times of RTO and TLP settings. The retransmission
 timeout and the probe timeout are the same as there is only one outstanding packet. This is a special case in the TLP
 specification that adds 200\,ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts
-for the delayed ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for PTO.
+for the delayed ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for PTO. In the case of two packet tail loss as shown in Figure~\ref{t2p},
+TLP performs better than the RTO retransmission for lower delay values.
+As the delay on the loss path increases RTO retransmission performs slightly better than TLP.
+In 120\,ms case, PTO is very close to RTO expiration and the second packet is recovered using ER which incurs additional delay in case of TLP.
+In RTO retransmission, all unacknowledged packets are retransmitted on a RTO expiration.
 
 Hence retransmission timeout is the mode of loss recovery in both RTO and TLP. This can be avoided by tuning
-the ACKs as discussed in~\cite{Rajiullah:2015}. In the case of two packet tail loss, TLP performs better than RTO as it employs a combination of TLP and ER to recover from losses. 
-If the probe or retransmission is lost, then TLP still performs better than RTO as it waits for one probe timeout and one RTO instead of two RTOs, thus avoiding the exponential backoff of the RTO.
+the ACKs as discussed in~\cite{Rajiullah:2015}.
+In the case of two packet tail loss, TLP performs better than RTO retransmission as it employs a combination of TLP and ER to recover from losses.
+If the probe or retransmission is lost, then TLP still performs better than RTO retransmission as it waits for one probe timeout and one RTO experiation instead of two RTO expirations, thus avoiding the exponential backoff of the RTO.
 
@@
 
-The expected retransmission behavior of MPTCP for RTO and Linux default settings is shown 
+The expected retransmission behavior of MPTCP for RTO retransmission and Linux default settings is shown 
 in Figure.~\ref{mptiming}. The actual pattern might differ in cases, where the RTT is larger than 200\,ms. 
@@
 is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
-configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO and TLP for 
+configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO retransmission and TLP for 
 cases when both paths are symmetric or the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
@@
 However, the latency improvement comes at a cost of additional network traffic from the retransmitted packets.
+This study is limited to emulated synthetic traffic and static drop pattern that serve the purpose of localizing the loss events.
+A comprehensive study covering various traffic types and loss patterns is considered for future investigation.
 Further, we plan to investigate less conservative approaches for other retransmission events.

### 2017-04-24 plots/plot_tlp_new.py

@@ -31,15 +31,15 @@ margin = 0.05
 bar_width = (1-2*margin)/3
 
 
-rects1 = plt.bar(index+margin, p1pe3 , bar_width,
-                 color='b',
+rects1 = plt.bar(index+margin, p2e3 , bar_width,
+                 color='m',
                  label='mptcp')
 
-rects2 = plt.bar(index+margin + bar_width, np1pe3 , bar_width,
-                 color='g',
+rects2 = plt.bar(index+margin + bar_width, np2e3 , bar_width,
+                 color='indigo',
                  label='mptcp-New')
 
-rects3 = plt.bar(index+margin + 2*bar_width, tp1pe3, bar_width,
+rects3 = plt.bar(index+margin + 2*bar_width, tp2e3, bar_width,
                  color='r',
                  label='tcp')
 

### 2017-04-26 main.tex
Update

@@
 The efficiency of loss recovery greatly influences the completion time of flows.
-In this paper we focus on the state of the art loss recovery mechanisms for TCP and Multipath TCP.
+In this paper we focus on state of the art loss recovery mechanisms for TCP and Multipath TCP.
 We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms.
@@
 
-This paper analyzes the performance of state of the art loss recovery mechanism of MPTCP and presents a case for improvement in the tail loss scenario. 
-For understanding the loss recovery and retransmission policies in MPTCP Linux implementation version 0.91, we consider several specific cases of tail losses.  
+This paper analyzes the performance of state of the art loss recovery mechanism of MPTCP and presents a case for improvement in the tail loss scenario.For understanding the loss recovery and retransmission policies in MPTCP Linux implementation version 0.91, we consider several specific cases of tail losses.
 We use the standard settings of the Linux kernel.
@@
 
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.45\textwidth]{images/fortest.pdf}
+\caption{Topology used for Emulation}\label{fig1}
+\end{center}
+\end{figure}
 %\kern-1em
@@
 
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.45\textwidth]{images/fortest.pdf}
-\caption{Topology used for Emulation}\label{fig1}
-\end{center}
-\end{figure}
 \begin{center}
@@
 \kern-5em
+\begin{figure*}[!tbp]
+ \begin{subfigure}[b]{0.32\textwidth}
+	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T1P.pdf}
+	\caption{Single packet}\label{t1p}
+ \end{subfigure}
+ \hfill
+ \begin{subfigure}[b]{0.32\textwidth}
+	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T2P.pdf}
+	\caption{Two packets }\label{t2p}
+ \end{subfigure}
+ \hfill
+ \begin{subfigure}[b]{0.32\textwidth}
+	\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
+	\caption{Single packet and probe loss }\label{t1pp}
+ \end{subfigure}
+ \caption{Tail loss scenarios using TCP}\label{tcpf}
+\end{figure*}
 \subsection{TCP}
@@
 
-\begin{figure*}[!tbp]
- \begin{subfigure}[b]{0.32\textwidth}
-	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T1P.pdf}
-	\caption{Single packet}\label{t1p}
- \end{subfigure}
- \hfill
- \begin{subfigure}[b]{0.32\textwidth}
-	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T2P.pdf}
-	\caption{Two packets }\label{t2p}
- \end{subfigure}
- \hfill
- \begin{subfigure}[b]{0.32\textwidth}
-	\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
-	\caption{Single packet and probe loss }\label{t1pp}
- \end{subfigure}
- \caption{Tail loss scenarios using TCP}\label{tcpf}
-\end{figure*}
 %\kern-1em
@@
 
-MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit
-the data until its congestion window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance
-is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
-configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO retransmission and TLP for 
-cases when both paths are symmetric or the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
-scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens in the second path with lowest RTT following the behavior depicted
-in Figure~\ref{timing1P}.
-
 \renewcommand{\thesubfigure}{\alph{subfigure}}
@@
 
+MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit
+the data until its congestion window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance
+is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
+configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO retransmission and TLP for 
+cases when both paths are symmetric or the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
+scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens in the second path with lowest RTT following the behavior depicted
+in Figure~\ref{timing1P}.
+
 In the case of single packet tail loss, the last outstanding packet is sent as probe and in case of n-degree tail loss,
@@
 The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the other 
-two settings in all scenarios except when the packets are lost on a high delay interface. In this scenario, the MPTCP retransmission follows the 
-behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of the last packet on the alternate path is 
-faster than the probe on the primary path.
+two settings in all scenarios except when the packets are lost on a high delay interface.
+In this scenario, the MPTCP retransmission follows the behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of the last packet on the alternate path is faster than the probe on the primary path.
 

### 2017-04-27 plots/plot_tlp_c.py
Update

@@ -40,11 +40,11 @@ margin = 0.05
 bar_width = (1-2*margin)/3
 
 
-rects1 = plt.bar(index+margin, p2e0 , bar_width,
+rects1 = plt.bar(index+margin, p1pe0 , bar_width,
                  color='b',
                  label='RTO')
 
-rects2 = plt.bar(index+margin + bar_width, p2e3 , bar_width,
+rects2 = plt.bar(index+margin + bar_width, p1pe3 , bar_width,
                  color='g',
                  label='TLP')
 

### 2017-04-27 plots/plot_tlp_new.py

@@ -31,16 +31,16 @@ margin = 0.05
 bar_width = (1-2*margin)/3
 
 
-rects1 = plt.bar(index+margin, p2e3 , bar_width,
-                 color='m',
+rects1 = plt.bar(index+margin, p1pe3 , bar_width,
+                 color='navy',
                  label='mptcp')
 
-rects2 = plt.bar(index+margin + bar_width, np2e3 , bar_width,
+rects2 = plt.bar(index+margin + bar_width, np1pe3 , bar_width,
                  color='indigo',
                  label='mptcp-New')
 
-rects3 = plt.bar(index+margin + 2*bar_width, tp2e3, bar_width,
-                 color='r',
+rects3 = plt.bar(index+margin + 2*bar_width, tp1pe3, bar_width,
+                 color='seagreen',
                  label='tcp')
 
 

### 2017-04-27 main.tex
Template update
added copyright info
tweaked ack-text
fixed figure placement

@@
-\documentclass[10pt,conference,compsoc]{IEEEtran}
+\documentclass[10pt,conference]{IEEEtran}
 \usepackage[utf8]{inputenc}
@@
 \author{\IEEEauthorblockN{Kiran~Yedugundla, Per~Hurtig, Anna~Brunstrom}
-\IEEEauthorblockA{Dept. of Computer Science, Karlstad University, Karlstad, Sweden}}
+\IEEEauthorblockA{Dept. of Computer Science, Karlstad University, Karlstad, Sweden\\
+    \{name.surname\}@kau.se
+}}
+
+\IEEEoverridecommandlockouts
+\IEEEpubid{%
+   \makebox[\columnwidth]{ISBN 978-3-901882-94-4~\copyright~2017 IFIP \hfill}%
+   \hspace{\columnsep}%
+   \makebox[\columnwidth]{ }%
+}
 
@@
 
-\begin{figure}[!ht]
+%\kern-1em
+\section{Experimental Setup}\label{exsetup}
+\begin{figure}
 \begin{center}
@@
 \end{figure}
-%\kern-1em
-\section{Experimental Setup}\label{exsetup}
 
@@
 \begin{center}
-
     \begin{table}
@@
 
+\begin{figure*}
+ \begin{subfigure}{0.32\textwidth}
+	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T1P.pdf}
+	\caption{Single packet}\label{t1p}
+ \end{subfigure}
+ \hfill
+ \begin{subfigure}{0.32\textwidth}
+	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T2P.pdf}
+	\caption{Two packets }\label{t2p}
+ \end{subfigure}
+ \hfill
+ \begin{subfigure}{0.32\textwidth}
+	\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
+	\caption{Single packet and probe loss }\label{t1pp}
+ \end{subfigure}
+ \caption{Tail loss scenarios using TCP}\label{tcpf}
+\end{figure*}
+
 %\kern-1em
@@
 
-\kern-1em
 \section{Observations and Discussion}\label{disc}
 This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, two packet tail loss and retransmission or probe loss.
-\kern-5em
-\begin{figure*}[!tbp]
- \begin{subfigure}[b]{0.32\textwidth}
-	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T1P.pdf}
-	\caption{Single packet}\label{t1p}
- \end{subfigure}
- \hfill
- \begin{subfigure}[b]{0.32\textwidth}
-	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T2P.pdf}
-	\caption{Two packets }\label{t2p}
- \end{subfigure}
- \hfill
- \begin{subfigure}[b]{0.32\textwidth}
-	\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
-	\caption{Single packet and probe loss }\label{t1pp}
- \end{subfigure}
- \caption{Tail loss scenarios using TCP}\label{tcpf}
-\end{figure*}
 \subsection{TCP}
@@
 
+MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit
+the data until its congestion window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance
+is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
+configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO retransmission and TLP for 
+cases when both paths are symmetric or the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
+scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens in the second path with lowest RTT following the behavior depicted
+in Figure~\ref{timing1P}.
+
 \renewcommand{\thesubfigure}{\alph{subfigure}}
@@
 \end{figure*}
-
-MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit
-the data until its congestion window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance
-is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
-configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO retransmission and TLP for 
-cases when both paths are symmetric or the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
-scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens in the second path with lowest RTT following the behavior depicted
-in Figure~\ref{timing1P}.
 
@@
 
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.46\textwidth]{images/timingER3NewTLP1.pdf}
-\end{center}
-\caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) Single packet b) Two packets c) Single packet and probe loss}\label{timingNew}
-\end{figure}
-
-%\kern-1em
-\section{Evaluation of modified MPTCP TLP}\label{eval}
-
-Results comparing the burst completion time for new MPTCP TLP with that of TCP and MPTCP TLP for the considered cases are provided in Figure~\ref{mpnew}.
-In Figure~\ref{1pn}, MPTCP-New performs better than TCP in all scenarios, performs equal to MPTCP in symmetric scenarios and performs better than MPTCP in asymmetric scenarios with a lower delay on the alternate path.
-This improvement is for the reason that MPTCP-New allows the packet to be retransmitted on the alternate path much before the RTO.
-In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New performs better than TCP in all scenarios and better than MPTCP in all scenarios except when the alternate path has a significantly higher delay than the original path, where performance is the same.
-In the single packet together with probe loss scenario as in Figure~\ref{1ppn}, MPTCP-New performs better than both MPTCP and TCP in all scenarios with PTO triggering the probe retransmission on alternate path much earlier than the RTO.
-
-\begin{figure*}[!tbp]
+\begin{figure*}
 \begin{subfigure}[b]{0.32\textwidth}
@@
 
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[angle=0, width=0.46\textwidth]{images/timingER3.pdf}
+\end{center}
+\caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) Single packet b) Two packets c) Single packet and probe loss}\label{timingNew}
+\end{figure}
+
+%\kern-1em
+\section{Evaluation of modified MPTCP TLP}\label{eval}
+
+Results comparing the burst completion time for new MPTCP TLP with that of TCP and MPTCP TLP for the considered cases are provided in Figure~\ref{mpnew}.
+In Figure~\ref{1pn}, MPTCP-New performs better than TCP in all scenarios, performs equal to MPTCP in symmetric scenarios and performs better than MPTCP in asymmetric scenarios with a lower delay on the alternate path.
+This improvement is for the reason that MPTCP-New allows the packet to be retransmitted on the alternate path much before the RTO.
+In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New performs better than TCP in all scenarios and better than MPTCP in all scenarios except when the alternate path has a significantly higher delay than the original path, where performance is the same.
+In the single packet together with probe loss scenario as in Figure~\ref{1ppn}, MPTCP-New performs better than both MPTCP and TCP in all scenarios with PTO triggering the probe retransmission on alternate path much earlier than the RTO.
+
 %\kern-1em
@@
 
+\section*{Acknowledgment}
+This work was in part funded by The Knowledge Foundation (KKS) through the SIDUS READY project.
+The views expressed are solely those of the author(s).
+
 %\kern-1em

### 2017-04-27 main.tex
Update

@@
 Packet losses are known to affect the performance of latency sensitive applications in the Internet such as media streaming and gaming. 
-Transport protocols recover from packet loss in order to provide reliable end to end communication and improving the quality of user experience.
-The efficiency of loss recovery greatly influences the completion time of flows.
-In this paper we focus on state of the art loss recovery mechanisms for TCP and Multipath TCP.
-We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms.
-Based on the observations, we propose an enhancement to the tail loss recovery in Multipath TCP to improve the loss recovery time.
-Our experiment results show consistent end to end latency performance improvement in considered scenarios.
+Transport protocols recover from packet loss in order to provide reliable end-to-end communication and improve user experience. The efficiency of 
+loss recovery mechanisms influences the completion time of flows. In this paper we focus on state-of-the-art loss recovery mechanisms for TCP 
+and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms, and based on the observations, 
+we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results using Linux 
+Multipath TCP implementation show consistent end to end latency performance improvement in considered scenarios. 
 \end{abstract}
@@
 
-Multipath TCP~(MPTCP) is an experimental standard proposed as an extension to TCP~\cite{rfc6824}.
-It allows devices with multiple interfaces to transmit data simultaneously on both interfaces, thereby improving the throughput of end-to-end connections.
-Using MPTCP, a connection can have multiple TCP subflows using different interfaces on different routes.
-Each TCP subflow imparts the functional behavior of a standard TCP flow.
+Multipath TCP~(MPTCP)~\cite{rfc6824} is an experimental standard proposed as an extension to TCP. MPTCP allows devices with multiple interfaces to 
+transmit data simultaneously on both interfaces, thereby improving the throughput of end-to-end connections. Using MPTCP, a connection can have 
+multiple TCP subflows using different interfaces on different routes. Each TCP subflow imparts the functional behavior of a standard TCP flow.
 
 TCP has two mechanisms for detecting and recovering from packet losses: Fast Retransmit (FR) and Retransmission Timeout~(RTO)~\cite{Flach:2013}. 
-Fast retransmit is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as a loss indication.
-In case of RTO, a sender waits for the packet loss to trigger a timeout before retransmitting that packet.
-During timeout, the sender cannot send any packets, but with FR the out-of-order packets trigger duplicate acknowledgements.
-The sender can start retransmission of missing packets without waiting for timeout.
-For flows where there are sufficiently many packets to be transferred, FR is quicker than RTO in detecting and recovering packet losses.
-Fast retransmit coupled with fast recovery can send data outstanding when there is enough data to send in long flows.
-On the other hand, shorter flows more frequently depend on RTO for packet loss detection and recovery, thus incurring more latency.
-A single packet loss in a short flow may take many RTTs to detect and recover.
-This scenario is also applicable to the packets at the end of the flow (or tail), in a long flow.
-Although, RTO is reliable, even in tail loss~(losses at the end of transmission) recovery, it is not efficient in terms of recovery completion time.
-Tail losses are one of the major causes of end-to-end latency detoriation in modern networks~\cite{Flach:2013}.
+Fast retransmit is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as an indication of loss. In the case 
+of RTO, a sender waits for the packet loss to trigger a timeout before retransmitting packet. During such a timeout, the sender cannot send any 
+packets. However, with FR the out-of-order packets trigger duplicate acknowledgements. The sender can start retransmission of missing packets 
+without waiting for timeout. For flows where there are sufficiently many packets to be transferred, FR is quicker than RTO in detecting and 
+recovering packet loss. Fast retransmit coupled with fast recovery can send data outstanding when there is enough data to send in long flows. 
+On the other hand, shorter flows more frequently depend on RTO for packet loss detection and recovery, thus incurring more latency. A single 
+packet loss in a short flow may take many RTTs to detect and recover. This scenario is also applicable to the packets at the end of the flow 
+(or tail), of a long flow. Although RTO is reliable even in tail loss recovery, it is not efficient in terms of recovery completion time. Tail 
+losses are one of the major causes of end-to-end latency detoriation in modern networks~\cite{Flach:2013}.
 
-TCP Loss Probe (TLP)~\cite{Flach:2013} is a mechanism to detect and recover from tail losses by employing a loss probe packet for a timeout shorter than the RTO.
-Probe timeout enables short flows to recover packet loss at the tail, by triggering FR.
-It assumes that other algorithms such as Early Retransmit~(ER)\cite{rfc5827} and Forward Acknowledgement~(FACK)~\cite{FACK} threshold based recovery are enabled.
-TLP is available in Linux kernel and enabled by default.
-Further, evaluations in~\cite{Rajiullah:2015} showed that TLP provides significant reductions in latency for short flows with one to n-degree tail losses.
+TCP Loss Probe (TLP)~\cite{Flach:2013} is a mechanism to detect and recover from tail losses by employing a loss probe packet for a timeout shorter
+than the RTO. Probe timeout enables short flows to recover packet loss at the tail, by triggering FR. It assumes that other algorithms such as 
+Early Retransmit~(ER)\cite{rfc5827} and Forward Acknowledgement~(FACK)~\cite{FACK} threshold based recovery are enabled. TLP is available in the Linux 
+kernel and enabled by default. Further, evaluations in~\cite{Rajiullah:2015} show that TLP provides significant latency reductions for short 
+flows with one to n-degree tail losses.
 
- MPTCP connections start with one TCP flow to send data over, and may add more flows to the connection if configured to.
-MPTCP uses a connection level congestion window as well as subflow level congestion windows for each subflow.
-Data is split across the flows and each data packet is scheduled on one of the subflows based on minimum RTT.
-Packet losses in each subflow are detected and recovered in a similar fashion as that of TCP.
-Loss recovery can be conducted at different levels: the MPTCP connection level (meta level) and at subflow level.
-If the recovery is handled at meta level, a lost packet may be rescheduled and retransmitted on the available subflow with the lowest RTT.
-If recovery is handled at the subflow level, the packet must be retransmitted on the same subflow.
-An important part of loss recovery process at subflow level is to follow the TCP semantics and retransmit lost packets on the initial path of transfer.
-However, the MPTCP standard~\cite{rfc6824} does not restrict on, how the loss recovery should happen in the implementation and which subflow should retransmit the lost packets.
+ MPTCP connections start with one TCP flow to send data over, and may add more flows to the connection if configured to. MPTCP uses a connection 
+ level congestion window as well as subflow level congestion windows for each subflow. Data is split across the flows and each data packet is 
+ scheduled on one of the subflows based on the minimum RTT. Packet losses in each subflow are detected and recovered in a similar fashion as that of 
+ TCP. Loss recovery can be conducted at different levels: the MPTCP connection level (meta level) and at subflow level. If the recovery is handled 
+ at meta level, a lost packet may be rescheduled and retransmitted on the available subflow with the lowest RTT. If recovery is handled at the 
+ subflow level, the packet must be retransmitted on the same subflow. An important part of loss recovery process at the subflow level is to follow 
+ the TCP semantics and retransmit lost packets on the initial path of transfer. However, the MPTCP standard~\cite{rfc6824} does not specify exactly, 
+ how the loss recovery should happen in the implementation and which subflow should retransmit the lost packets.
 
-In this paper, we analyze the retransmission behavior of MPTCP in selected tail loss cases to identify the scope for improvement.
-We propose a change in the retransmission strategy of MPTCP to improve the latency performance.
-We provide related research on MPTCP retransmission in section~\ref{relwork}, and describe the experimental setup in section~\ref{exsetup}.
-We discuss observations from the experiments on state of the art retransmission behavior of TCP and MPTCP in section~\ref{disc}.
-Based on the observations, we propose an improvement to the current Linux implementation of TLP for MPTCP in section~\ref{impr}.
-In section~\ref{eval}, we provide an evaluation of the proposed TLP for MPTCP mechanism and conclude the paper in section~\ref{conc}.
-Our experimental results indicate improvement in the performance of TLP by reducing recovery time in the event of tail losses.
+In this paper, we analyze the retransmission behavior of MPTCP in selected tail loss cases to identify the scope for improvement. We propose a 
+change in the retransmission strategy of MPTCP to improve the latency performance. We provide related research on MPTCP retransmission mechanisms in 
+Section~\ref{relwork}, and describe the experimental setup in Section~\ref{exsetup}. We discuss observations from the experiments on state 
+of the art retransmission behavior of TCP and MPTCP in Section~\ref{disc}. Based on the observations, we propose an improvement to the current 
+Linux implementation of TLP for MPTCP in Section~\ref{impr}. In Section~\ref{eval}, we provide an evaluation of the proposed TLP for MPTCP 
+mechanism and conclude the paper in Section~\ref{conc}. Our experimental results indicate improvement in the performance of TLP by reducing 
+recovery time in the event of tail losses.
 
@@
 
-Retransmission schemes are of significant importance in achieving low flow completion time.
-Several IETF RFCs such as SACK based recovery~\cite{rfc6675}, Limited Transmit~\cite{rfc3042} and Early Retransmission~\cite{rfc5827} have led to improved retransmission strategies for TCP.
-TLP is part of the various deployable loss recovery mechanisms proposed in~\cite{Flach:2013}, suggesting that current Linux TCP implementation provides room for improvement in several packet loss scenarios.
+Retransmission schemes are of significant importance in achieving a low flow completion time. Several IETF RFCs such as SACK based 
+recovery~\cite{rfc6675}, Limited Transmit~\cite{rfc3042} and Early Retransmission~\cite{rfc5827} have led to improved retransmission 
+strategies for TCP. TLP is part of the various deployable loss recovery mechanisms proposed in~\cite{Flach:2013}, suggesting that current 
+Linux TCP implementation provides room for improvement in several packet loss scenarios.
 
-In a TCP flow, sequence numbers are used to identify lost packets and retransmit them.
-MPTCP uses two levels of sequence numbers to support efficient data transfer, namely data sequence and TCP sequence.
-Data sequence numbers are for the end-to-end data transfer and TCP sequence numbers are for individual flow data sequence.
-There is an association between data sequence numbers and TCP sequence numbers within a subflow.
-A loss occurring in an individual TCP flow corresponds to a loss in end-to-end data.
-Multipath TCP, as an extension of TCP, uses the basics of TCP retransmission.
-However, the interaction with two levels makes the problem of retransmission more challenging than in TCP.
-The dual sequence numbering enables MPTCP to respond to loss of packets by retransmitting the lost packets on an alternate path.
+In a TCP flow, sequence numbers are used to identify lost packets and retransmit them. MPTCP uses two levels of sequence numbers to support 
+efficient data transfer, namely data sequence and TCP sequence. Data sequence numbers are for the end-to-end data transfer and TCP sequence 
+numbers are for individual flow data sequence. There is an association between data sequence numbers and TCP sequence numbers within a subflow.  
+A loss occurring in an individual TCP flow corresponds to a loss in end-to-end data. Multipath TCP, as an extension of TCP, uses the 
+TCP retransmission mechanism. However, the interaction between the two levels makes the problem of retransmitting more challenging than in TCP. The dual sequence 
+numbering enables MPTCP to respond to loss of packets by retransmitting the lost packets on an alternate path.
 
-The Linux implementation of MPTCP uses a set of retransmission heuristics to handle retransmissions.
-The data outstanding on a timed out subflow should be rescheduled for transmission on a different subflow using timeout as the indicator.
-Fast retransmit on a subflow will not trigger retransmission on another subflow.
-The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 4G vs WLAN, leading to flows with different delay characteristics.
-Significant delay differences in available paths can cause receive buffer limitations, as the receiver has to wait for the data sent on a slow path, while buffering the data on the fast path.
-In such cases, opportunistic retransmission~\cite{Costin} improves the latency by using fast path to retransmit the data sent on an another path.
-Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fastest paths.
-Such quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always as well to follow TCP semantics.
-The MPTCP standard~\cite{rfc6824} suggests a more conservative approach.
-MPTCP is designed to exploit the multiple interfaces and often one or more of these could be wireless with completely different characteristics than that of wired interfaces.
-Authors in~\cite{Shin} argue that the calculation of RTO for MPTCP flows should include the interface characteristics to improve loss recovery time.
+The Linux implementation of MPTCP uses a set of retransmission heuristics to handle retransmissions. The data outstanding on a timed out subflow 
+should be rescheduled for transmission on a different subflow using timeout as the indicator. Fast retransmit on a subflow will not trigger 
+retransmission on another subflow. The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 
+4G vs WLAN, leading to flows with different delay characteristics. Significant delay differences in available paths can cause receive buffer 
+limitations, as the receiver has to wait for the data sent on a slow path, while buffering the data on the fast path. In such cases, opportunistic 
+retransmission~\cite{Costin} improves the latency by using fast path to retransmit the data sent on an another path. Authors in~\cite{fuso}, try 
+to exploit the path diversity by quickly retransmitting on the fastest paths. Such quick retransmission comes with a cost of redundant packets as 
+each individual TCP flow should retransmit always as well to follow TCP semantics. The MPTCP standard~\cite{rfc6824} suggests a more conservative 
+approach. MPTCP is designed to exploit the multiple interfaces and often one or more of these could be wireless with completely different 
+characteristics than that of wired interfaces. Authors in~\cite{Shin} argue that the calculation of RTO for MPTCP flows should include the 
+interface characteristics to improve loss recovery time.
 
@@
 
-This paper analyzes the performance of state of the art loss recovery mechanism of MPTCP and presents a case for improvement in the tail loss scenario.For understanding the loss recovery and retransmission policies in MPTCP Linux implementation version 0.91, we consider several specific cases of tail losses.
-We use the standard settings of the Linux kernel.
-The Linux implementation of TCP is considered as the baseline TCP retransmission for implementation flexibility and availability of TLP support.
+This paper analyzes the performance of state-of-the-art loss recovery mechanism for MPTCP and presents a case for improvement in the tail loss 
+scenario. For understanding the loss recovery and retransmission policies in the MPTCP Linux implementation (version 0.91), we consider several specific 
+cases of tail losses. We use the default settings of the Linux kernel. The Linux implementation of TCP is considered as the baseline TCP 
+retransmission for implementation flexibility and availability of TLP support.
 
@@
 
-Experiments use the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}.
+Experiments are run in the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}.
 The client node connect to two wireless interfaces 3G or 4G and WLAN and the server node connect to a wired router.
@@
 Characteristics of the connection setup and assumptions about the parameters are provided in Table~\ref{tab1}.
-%The parameters are chosen to be close to real values of bandwidth and delay of respective technologies.
-There is no attempt in this study to focus on the effect of link detailed characteristics in retransmission performance.
+There is no attempt in this study to focus on the effect of link specific characteristics in retransmission performance.
 
@@
 %\kern-1em
-\subsection{Testing retransmission with deterministic loss patterns}
-In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
-Losses are generated by associating netem with corresponding interfaces and dropping select packets.
-This process is simplified by using the KAUNetem tool~\cite{Garcia2016}, which allows experimenter to e.g., drop packets based on their positions in a flow.
-A MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between the client and the server.
-So one has to wait both in time and packets for the second subflow establishment.
-We send two bursts of 80 packets each and drop tail packets in the second burst at one of the interfaces.
-This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that subflow.
-We calculate the burst completion time for the second burst that has specified tail losses.
-The objective of this experiment is to evaluate retransmission mechanisms with tail loss (single packet loss and multiple tail loss) and temporary path loss.
-Three test cases are considered for experiments: single packet tail loss and two packet tail loss for tail loss evaluation and single packet along with probe~(in TLP) or retransmission~(in RTO) loss for path loss evaluation.
+\subsection{Testing retransmission mechanism with deterministic loss patterns}
+In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a 
+deterministic drop pattern. Losses are generated by associating netem with corresponding interfaces and dropping select packets. This process 
+is simplified by using the KAUNetem tool~\cite{Garcia2016}, which allows an experimenter to e.g., drop packets based on their positions in a flow. 
+A MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between the client 
+and the server. So one has to wait both in time and packets for the second subflow establishment. We send two bursts of 80 packets each and drop 
+tail packets in the second burst at one of the interfaces. This is to ensure that the necessary and sufficient conditions for probe triggering 
+are met and tail loss probe is generated on that subflow. We calculate the burst completion time for the second burst that has specified tail 
+losses. The objective of this experiment is to evaluate retransmission mechanisms with tail loss (single packet loss and multiple tail loss) and 
+temporary path loss. Three test cases are considered for the experiments: single packet tail loss and two packet tail loss for tail loss evaluation 
+and single packet along with probe~(in TLP) or retransmission~(in RTO) loss for path loss evaluation.
 
 In Linux, TCP supports multiple retransmission mechanisms that can be controlled by a system setting. The default setting in Linux enables both
-ER and TLP, which may not be the case with other operating systems. In this paper, we use TLP to denote the default Linux setting and RTO to denote the use of retransmission timeout only.
+ER and TLP, which may not be the case with other operating systems. In this paper, we use TLP to denote the default Linux setting and RTO to denote
+the use of retransmission timeout only.
 
-The setup has server and client with Linux supporting MPTCP running on them.
-The Client has two interfaces with one way delay on each interface ranging from 20\,ms to 120\,ms.
-The intention is to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in the link delays.
-We consider 5 scenarios with delay pairings 20\,ms-20\,ms, 20\,ms-30\,ms, 20\,ms-120\,ms, 30\,ms-20\,ms, 120\,ms-20\,ms to understand the effect of delay difference in the performance as shown in Table~\ref{tab1}.
+The setup has server and client with Linux supporting MPTCP running on them. The Client has two interfaces with one way delay on each interface 
+ranging from 20\,ms to 120\,ms. The intention is to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in the link delays. We 
+consider 5 scenarios with delay pairings 20\,ms-20\,ms, 20\,ms-30\,ms, 20\,ms-120\,ms, 30\,ms-20\,ms, 120\,ms-20\,ms to understand the effect of
+delay difference in the performance as shown in Table~\ref{tab1}.
 
 \section{Observations and Discussion}\label{disc}
-This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, two packet tail loss and retransmission or probe loss.
+This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, two packet tail loss and 
+retransmission or probe loss.
 \subsection{TCP}
 Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
-starts with the results using TCP and compare it with that of MPTCP. Figure~\ref{tcpf} represent the comparison of TCP 
-burst completion times. Path asymmetry is irrelevant in this case as TCP uses a single path.
+starts with the results using TCP and compares it with that of MPTCP. Figure~\ref{tcpf} represents the comparison of TCP burst completion times. Path
+asymmetry is irrelevant in this case as TCP uses a single path.
 
-In the case of single packet tail loss as shown in Figure~\ref{t1p}, there is no difference in burst completion times of RTO and TLP settings. The retransmission
-timeout and the probe timeout are the same as there is only one outstanding packet. This is a special case in the TLP
-specification that adds 200\,ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts
-for the delayed ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for PTO. In the case of two packet tail loss as shown in Figure~\ref{t2p},
-TLP performs better than the RTO retransmission for lower delay values.
-As the delay on the loss path increases RTO retransmission performs slightly better than TLP.
-In 120\,ms case, PTO is very close to RTO expiration and the second packet is recovered using ER which incurs additional delay in case of TLP.
-In RTO retransmission, all unacknowledged packets are retransmitted on a RTO expiration.
+In the case of single packet tail loss, as shown in Figure~\ref{t1p}, there is no difference in burst completion between RTO and TLP. The 
+retransmission timeout and the probe timeout are the same as there is only one outstanding packet. This is a special case in the TLP
+specification that adds 200\,ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts for the delayed 
+ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for PTO. In the case of two packet tail loss, as shown 
+in Figure~\ref{t2p}, TLP performs better than the RTO retransmission for lower delay values. As the delay on the loss path increases RTO retransmission performs 
+slightly better than TLP. In 120\,ms case, PTO is very close to RTO expiration and the second packet is recovered using ER which incurs additional 
+delay in case of TLP. In RTO retransmission, all unacknowledged packets are retransmitted on a RTO expiration.
 
-Hence retransmission timeout is the mode of loss recovery in both RTO and TLP. This can be avoided by tuning
-the ACKs as discussed in~\cite{Rajiullah:2015}.
-In the case of two packet tail loss, TLP performs better than RTO retransmission as it employs a combination of TLP and ER to recover from losses.
-If the probe or retransmission is lost, then TLP still performs better than RTO retransmission as it waits for one probe timeout and one RTO experiation instead of two RTO expirations, thus avoiding the exponential backoff of the RTO.
+Hence retransmission timeout is the mode of loss recovery in both RTO and TLP. This can be avoided by tuning the ACKs as discussed 
+in~\cite{Rajiullah:2015}. In the case of two packet tail loss, TLP performs better than RTO retransmission as it employs a combination of TLP 
+and ER to recover from losses. If the probe or retransmission is lost, then TLP still performs better than RTO retransmission as it waits for one 
+probe timeout and one RTO experiation instead of two RTO expirations, thus avoiding the exponential backoff of the RTO.
 
@@
 The expected retransmission behavior of MPTCP for RTO retransmission and Linux default settings is shown 
-in Figure.~\ref{mptiming}. The actual pattern might differ in cases, where the RTT is larger than 200\,ms. 
+in Figure~\ref{mptiming}. The actual pattern might be different in cases where the RTT is larger than 200\,ms.
+
 \renewcommand{\thesubfigure}{\roman{subfigure}}
@@
 
-MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit
-the data until its congestion window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance
-is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
-configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO retransmission and TLP for 
-cases when both paths are symmetric or the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
-scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens in the second path with lowest RTT following the behavior depicted
-in Figure~\ref{timing1P}.
+MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit the data until its congestion 
+window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance is sensitive to asymmetry in paths. The burst completion times 
+on single packet tail loss calculated for each delay configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no 
+difference in performance between RTO retransmission and TLP for cases when both paths are symmetric or the first path has lowest RTT. RTO performs 
+better than TLP when the first path is not the lowest RTT path. In such scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens 
+in the second path with lowest RTT following the behavior depicted in Figure~\ref{timing1P}.
 
@@
 
-In the case of single packet tail loss, the last outstanding packet is sent as probe and in case of n-degree tail loss,
-the latest packet is sent as probe to trigger fast retransmit of the packets in between. In order to observe
-the performance of n-degree tail loss, we consider dropping the last two packets. 
-The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the other 
-two settings in all scenarios except when the packets are lost on a high delay interface.
-In this scenario, the MPTCP retransmission follows the behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of the last packet on the alternate path is faster than the probe on the primary path.
+In the case of single packet tail loss, the last outstanding packet is sent as probe and in case of n-degree tail loss, the latest packet is sent 
+as probe to trigger fast retransmit of the packets in between. In order to observe the performance of n-degree tail loss, we consider dropping the 
+last two packets. The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the other 
+two settings in all scenarios except when packets are lost on a high delay interface. In this scenario, the MPTCP retransmission follows the 
+behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of the last packet on the alternate 
+path is faster than the probe on the primary path.
 
-To further study the performance of TLP in the case when there is longer path interruption on one path, we tried to drop the probe packet along with the last packet. In this scenario, the RTO
-setting results in much lower burst completion times.
-Comparing with the performance of TLP in TCP from Figure~\ref{t1pp}, where we saw that use of TLP improved the burst completion time, TLP in MPTCP did not improve the performance.
-The reason lies in the way the retransmission and loss probe transmission is carried out in MPTCP.
-Retransmission occurs after an RTO time on both paths.
-Even if the retransmission fails on one path, the client receives the packet on the other path as shown in Figure~\ref{timing1PP}. For TLP, in
-current implementation, the loss probe is sent on the actual path of the first transmission but not on both paths. This incomplete impartation of
-TLP from TCP to MPTCP led to the increase in burst completion time by an RTO for MPTCP.
-
-%\begin{table}[!ht]
-%\centering
-%\caption{Tail loss scenarios with tcp.early.retrans = 3 default}
-%\label{ret3}
-%\begin{tabular}{|l|l|l|l|l|}
-%\hline
-% testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
-%Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
-%Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
-%Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                &  \\ \hline
-%\end{tabular}
-%\end{table}
-
-%\begin{table}[!ht]
-%\centering
-%\caption{Tail loss scenarios with tcp.early.retrans = 0 ER disabled}
-%\label{ret0}
-%\begin{tabular}{|l|l|l|l|l|}
-%\hline
-% testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
-%Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
-%Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
-%Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
-%Tail drop 4 & 20ms - 120ms &  			&		&  \\ \hline
-%Tail drop 5 & 120ms - 20ms &  			& 		& \\ \hline 
-%\end{tabular}
-%\end{table}
-
-%\begin{table}[!ht]
-%\centering
-%\caption{Tail loss scenarios with tcp.early.retrans = 4 TLP disabled}
-%\label{ret4}
-%\begin{tabular}{|l|l|l|l|l|}
-%\hline
-% testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
-%Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
-%Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
-%Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
-%\end{tabular}
-%\end{table}
+To further study the performance of TLP in the case when there is longer path interruption on one path, we tried to drop the probe packet along 
+with the last packet. In this scenario, the RTO setting results in much lower burst completion times. Comparing with the performance of TLP in 
+TCP from Figure~\ref{t1pp}, where we saw that use of TLP improved the burst completion time, TLP in MPTCP did not improve the performance. The 
+reason lies in the way the retransmission and loss probe transmission is carried out in MPTCP. Retransmission occurs after an RTO time on both 
+paths. Even if the retransmission fails on one path, the client receives the packet on the other path as shown in Figure~\ref{timing1PP}. For TLP, 
+in the current implementation, the loss probe is sent on the actual path of the first transmission but not on both paths. This incomplete impartation 
+of TLP from TCP to MPTCP led to the increase in burst completion time by an RTO for MPTCP.
 
@@
 \section{Improvements to TLP for MPTCP}\label{impr}
-The goal of Tail Loss Probe~(TLP) is to reduce tail latency of short flows.
-It achieves this by converting retransmission timeouts (RTOs) occurring due to tail losses into fast recovery.
-TLP transmits one packet in two round-trips when a connection is in Open state and isn't receiving any ACKs.
-The transmitted packet, aka loss probe, can be either new or a retransmission.
-When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout.
-Our results show that the current TLP implementation for MPTCP helps improving the latency performance in cases when the paths are symmetric or the first path~(for the packet) has lowest RTT.
-However, it reduces the performance in cases when the first path is not the lowest RTT path or when there is longer path interruption.
+The goal of Tail Loss Probe~(TLP) is to reduce tail latency of short flows. It achieves this by converting retransmission timeouts (RTOs) occurring
+due to tail losses into fast recovery. TLP transmits one packet in two round-trips when a connection is in Open state and is not receiving any ACKs.
+The transmitted packet, or loss probe, can be either new or a retransmission. When there is tail loss, the ACK from a loss probe triggers 
+FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout. Our results show that the current TLP implementation for 
+MPTCP helps improving the latency performance in cases when the paths are symmetric or the first path~(for the packet) has lowest RTT. However, 
+it reduces the performance in cases when the first path is not the lowest RTT path or when there is longer path interruption.
 
-The current implementation of TLP is at the TCP flow level.
-In the event of RTO timeout, the MPTCP retransmission happens with a re-injection in to the scheduler along with sending the lost packet on the same path.
-But in the event of Probe timeout, the loss probe packet is being sent on the same path without being re-injected in to the scheduler.
-We propose a modification to the probe sending mechanism for MPTCP by re-injecting the probe packet in to the MPTCP scheduler in the event of PTO.
-As the probe timeout is meant for acting faster than the RTO timeout with essentially the same consequence as a retransmission, it is relevant to follow the same process as upon an RTO.
-It is expected that this modification is useful when the alternate path has lower delay than the packet original transmission path or when there is longer path interruption.
-The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied loss scenarios.
+The current implementation of TLP is at the TCP flow level. In the event of RTO timeout, the MPTCP retransmission happens with a re-injection in 
+to the scheduler along with sending the lost packet on the same path. But in the event of Probe timeout, the loss probe packet is being sent on 
+the same path without being re-injected in to the scheduler. We propose a modification to the probe sending mechanism for MPTCP by re-injecting 
+the probe packet in to the MPTCP scheduler in the event of PTO. As the probe timeout is meant for acting faster than the RTO timeout with 
+essentially the same consequence as a retransmission, it is relevant to follow the same process as upon an RTO. It is expected that this 
+modification is useful when the alternate path has lower delay than the packet original transmission path or when there is longer path 
+interruption. The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied 
+loss scenarios.
 
@@
 
-Results comparing the burst completion time for new MPTCP TLP with that of TCP and MPTCP TLP for the considered cases are provided in Figure~\ref{mpnew}.
-In Figure~\ref{1pn}, MPTCP-New performs better than TCP in all scenarios, performs equal to MPTCP in symmetric scenarios and performs better than MPTCP in asymmetric scenarios with a lower delay on the alternate path.
-This improvement is for the reason that MPTCP-New allows the packet to be retransmitted on the alternate path much before the RTO.
-In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New performs better than TCP in all scenarios and better than MPTCP in all scenarios except when the alternate path has a significantly higher delay than the original path, where performance is the same.
-In the single packet together with probe loss scenario as in Figure~\ref{1ppn}, MPTCP-New performs better than both MPTCP and TCP in all scenarios with PTO triggering the probe retransmission on alternate path much earlier than the RTO.
+Results comparing the burst completion time for new MPTCP TLP with that of TCP and MPTCP TLP for the considered cases are provided in 
+Figure~\ref{mpnew}. In Figure~\ref{1pn}, MPTCP-New performs better than TCP in all scenarios, performs equal to MPTCP in symmetric scenarios and 
+performs better than MPTCP in asymmetric scenarios with a lower delay on the alternate path. This improvement is for the reason that MPTCP-New 
+allows the packet to be retransmitted on the alternate path much before the RTO. In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New
+performs better than TCP in all scenarios and better than MPTCP in all scenarios except when the alternate path has a significantly higher delay 
+than the original path, where performance is the same. In the single packet together with probe loss scenario as in Figure~\ref{1ppn}, MPTCP-New 
+performs better than both MPTCP and TCP in all scenarios with PTO triggering the probe retransmission on alternate path much earlier than the RTO.
 
@@
 This paper provides a short comparative analysis of retransmission behavior with TCP and MPTCP protocols using various loss patterns for tail loss.
-The three cases of tail loss together with deterministic loss patterns allow us to dissect the retransmission behavior at packet level to understand possible improvements.
-We propose a less conservative approach to trigger retransmission on an alternate path in the event of tail loss probe timeout, which otherwise triggered with an RTO. 
-Our emulation experiments  using a modified Linux implementation, show that the proposed approach in fact improves the burst completion time in most cases and equals the existing
-implementation in other cases.
-The proposed approach improves the performance when the alternate path has lower delay than the original loss path and the advantage increases with degree of asymmetry between the paths.
-Temporary path failures can cause probe loss along with packet loss.
-The approach is very effective in case of probe loss which otherwise incur RTO timeout to trigger retransmission on either path.
-However, the latency improvement comes at a cost of additional network traffic from the retransmitted packets.
-This study is limited to emulated synthetic traffic and static drop pattern that serve the purpose of localizing the loss events.
-A comprehensive study covering various traffic types and loss patterns is considered for future investigation.
-Further, we plan to investigate less conservative approaches for other retransmission events.
+The three cases of tail loss together with deterministic loss patterns allow us to dissect the retransmission behavior at packet level to 
+understand possible improvements.
+We propose a less conservative approach to trigger retransmission on an alternate path in the event of tail loss probe timeout, which otherwise 
+triggered with an RTO. Our emulation experiments  using a modified Linux implementation, show that the proposed approach in fact improves the 
+burst completion time in most cases and equals the existing implementation in other cases. The proposed approach improves the performance when 
+the alternate path has lower delay than the original loss path and the advantage increases with degree of asymmetry between the paths. 
+Temporary path failures can cause probe loss along with packet loss. The approach is very effective in case of probe loss which otherwise 
+incur RTO timeout to trigger retransmission on either path. However, the latency improvement comes at a cost of additional network traffic 
+from the retransmitted packets. This study is limited to emulated synthetic traffic and static drop pattern that serve the purpose of localizing 
+the loss events. A comprehensive study covering various traffic types and loss patterns is considered for future investigation. Further, we plan 
+to investigate less conservative approaches for other retransmission events.
 
 \section*{Acknowledgment}
-This work was in part funded by The Knowledge Foundation (KKS) through the SIDUS READY project.
-The views expressed are solely those of the author(s).
+This work was in part funded by The Knowledge Foundation (KKS) through the SIDUS READY project. The views expressed are solely those of the 
+author(s).
 

### 2017-04-27 main.tex
Update
made timing diagram larger and changed section name from "Conclusion" -> "Conclusions"

@@
 
-The current implementation of TLP is at the TCP flow level. In the event of RTO timeout, the MPTCP retransmission happens with a re-injection in 
-to the scheduler along with sending the lost packet on the same path. But in the event of Probe timeout, the loss probe packet is being sent on 
-the same path without being re-injected in to the scheduler. We propose a modification to the probe sending mechanism for MPTCP by re-injecting 
-the probe packet in to the MPTCP scheduler in the event of PTO. As the probe timeout is meant for acting faster than the RTO timeout with 
-essentially the same consequence as a retransmission, it is relevant to follow the same process as upon an RTO. It is expected that this 
-modification is useful when the alternate path has lower delay than the packet original transmission path or when there is longer path 
-interruption. The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied 
-loss scenarios.
+\begin{figure*}
+\begin{center}
+    \includegraphics[angle=0, width=0.8\textwidth]{images/timingER3NewTLP1}
+\end{center}
+\caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) Single packet b) Two packets c) Single packet and probe loss}\label{timingNew}
+\end{figure*}
 
@@
 
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[angle=0, width=0.46\textwidth]{images/timingER3.pdf}
-\end{center}
-\caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) Single packet b) Two packets c) Single packet and probe loss}\label{timingNew}
-\end{figure}
+The current implementation of TLP is at the TCP flow level. In the event of RTO timeout, the MPTCP retransmission happens with a re-injection in 
+to the scheduler along with sending the lost packet on the same path. But in the event of Probe timeout, the loss probe packet is being sent on 
+the same path without being re-injected in to the scheduler. We propose a modification to the probe sending mechanism for MPTCP by re-injecting 
+the probe packet in to the MPTCP scheduler in the event of PTO. As the probe timeout is meant for acting faster than the RTO timeout with 
+essentially the same consequence as a retransmission, it is relevant to follow the same process as upon an RTO. It is expected that this 
+modification is useful when the alternate path has lower delay than the packet original transmission path or when there is longer path 
+interruption. The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied 
+loss scenarios.
 
@@
 Figure~\ref{mpnew}. In Figure~\ref{1pn}, MPTCP-New performs better than TCP in all scenarios, performs equal to MPTCP in symmetric scenarios and 
-performs better than MPTCP in asymmetric scenarios with a lower delay on the alternate path. This improvement is for the reason that MPTCP-New 
-allows the packet to be retransmitted on the alternate path much before the RTO. In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New
+performs better than MPTCP in asymmetric scenarios with a lower delay on the alternate path. MPTCP-New achieves this improvement by retransmitting
+ lost packets on the alternate path much before the RTO. In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New
 performs better than TCP in all scenarios and better than MPTCP in all scenarios except when the alternate path has a significantly higher delay 
@@
 %\kern-1em
-\section{Conclusion}\label{conc}
+\section{Conclusions}\label{conc}
 This paper provides a short comparative analysis of retransmission behavior with TCP and MPTCP protocols using various loss patterns for tail loss.

### 2017-04-27 main.tex
Update

--- 2017-04-27 main.tex
@@
 
-\begin{figure}[!ht]
+\begin{figure*}[!ht]
 \begin{center}
-\includegraphics[angle=0, width=0.46\textwidth]{images/timingER3.pdf}
+\includegraphics[angle=0, width=\textwidth]{images/timingER3NewTLP1.pdf}
 \end{center}
 \caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) Single packet b) Two packets c) Single packet and probe loss}\label{timingNew}
-\end{figure}
+\end{figure*}
 

### 2017-04-27 main.tex
update

--- 2017-04-27 main.tex
@@
 
-Multipath TCP~(MPTCP)~\cite{rfc6824} is an experimental standard proposed as an extension to TCP. MPTCP allows devices with multiple interfaces to 
+Multipath TCP~(MPTCP)~\cite{rfc6824} is an experimental standard proposed as an extension to TCP~\cite{rfc793}. MPTCP allows devices with multiple interfaces to 
 transmit data simultaneously on both interfaces, thereby improving the throughput of end-to-end connections. Using MPTCP, a connection can have 
@@
 The goal of Tail Loss Probe~(TLP) is to reduce tail latency of short flows. It achieves this by converting retransmission timeouts (RTOs) occurring
-due to tail losses into fast recovery. TLP transmits one packet in two round-trips when a connection is in Open state and is not receiving any ACKs.
-The transmitted packet, or loss probe, can be either new or a retransmission. When there is tail loss, the ACK from a loss probe triggers 
-FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout. Our results show that the current TLP implementation for 
-MPTCP helps improving the latency performance in cases when the paths are symmetric or the first path~(for the packet) has lowest RTT. However, 
-it reduces the performance in cases when the first path is not the lowest RTT path or when there is longer path interruption.
+due to tail losses into fast recovery. In Open state of TCP connection, a sender is receiving in-sequence ACKs and has not detected any lost segments.
+TLP transmits one packet in two round-trips when a connection is in Open state and is not receiving any ACKs. The transmitted packet, or loss probe, 
+can be either new or a retransmission. When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus 
+avoiding a costly retransmission timeout. Our results show that the current TLP implementation for MPTCP helps improving the latency performance in 
+cases when the paths are symmetric or the first path~(for the packet) has lowest RTT. However, it reduces the performance in cases when the first path 
+is not the lowest RTT path or when there is longer path interruption.
 
@@
 interruption. The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied 
-loss scenarios.
+loss scenarios. For asymmetric scenarios, the behavior changes with respect to scheduler perceived delay differences are assumed.
 
@@
 performs better than MPTCP in asymmetric scenarios with a lower delay on the alternate path. MPTCP-New achieves this improvement by retransmitting
- lost packets on the alternate path much before the RTO. In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New
-performs better than TCP in all scenarios and better than MPTCP in all scenarios except when the alternate path has a significantly higher delay 
-than the original path, where performance is the same. In the single packet together with probe loss scenario as in Figure~\ref{1ppn}, MPTCP-New 
-performs better than both MPTCP and TCP in all scenarios with PTO triggering the probe retransmission on alternate path much earlier than the RTO.
+ lost packet on the alternate path much before the RTO. It is also observed that there is no performance degradation when the alternate path has 
+ higher delay as the probe packet (same as the lost packet when there is only one outstanding packet in this case) on the primary path reaches 
+ completing the loss recovery. In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New performs better than TCP in all scenarios and better 
+ than MPTCP in all scenarios except when the alternate path has a significantly higher delay than the original path, where performance is the same. 
+ In the single packet together with probe loss scenario as in Figure~\ref{1ppn}, MPTCP-New performs better than both MPTCP and TCP in all scenarios 
+ with PTO triggering the probe retransmission on alternate path much earlier than the RTO.
 

### 2017-04-27 main.tex
Update

@@
 interruption. The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied 
-loss scenarios. For asymmetric scenarios, the behavior changes with respect to scheduler perceived delay differences are assumed.
+loss scenarios. In the case of single packet tail loss, lost packet is retransmitted on the alternate path at the same time as probe sent on primary
+path. Since probe is the last transmitted packet, that is same as the retransmitted first outstanding packet in single packet tail loss. When two tail packets are lost,
+probe sent on the primary is the last transmitted packet and first outstanding packet is retransmitted on the alternate path. This difference in packet
+sequence numbers is expected on the two paths due to the nature of recovery mechanism used in each path. In the case of packet and probe loss in the 
+primary path, the alternate path retransmission is same as the probe packet.
 

### 2017-04-27 main.tex
Correction

@@
 we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results using Linux 
-Multipath TCP implementation show consistent end to end latency performance improvement in considered scenarios. 
+Multipath TCP implementation show consistent end-to-end latency performance improvement in considered scenarios. 
 \end{abstract}
@@
  lost packet on the alternate path much before the RTO. It is also observed that there is no performance degradation when the alternate path has 
- higher delay as the probe packet (same as the lost packet when there is only one outstanding packet in this case) on the primary path reaches 
- completing the loss recovery. In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New performs better than TCP in all scenarios and better 
- than MPTCP in all scenarios except when the alternate path has a significantly higher delay than the original path, where performance is the same. 
- In the single packet together with probe loss scenario as in Figure~\ref{1ppn}, MPTCP-New performs better than both MPTCP and TCP in all scenarios 
- with PTO triggering the probe retransmission on alternate path much earlier than the RTO.
+ higher delay. Probe packet on the primary path reaches faster than the retransmitted path and completes loss recovery at MPTCP level. In two packet 
+ tail loss scenario as in Figure~\ref{2pn}, MPTCP-New performs better than TCP in all scenarios and better than MPTCP in all scenarios except when 
+ the alternate path has a significantly higher delay than the original path, where performance is the same. In the single packet together with probe 
+ loss scenario as in Figure~\ref{1ppn}, MPTCP-New performs better than both MPTCP and TCP in all scenarios with PTO triggering the probe retransmission
+ on alternate path much earlier than the RTO.
 

### 2017-04-27 plots/plot_tlp_new.py
Color update 2

@@ -31,15 +31,15 @@ margin = 0.05
 bar_width = (1-2*margin)/3
 
 
-rects1 = plt.bar(index+margin, p1pe3 , bar_width,
+rects1 = plt.bar(index+margin, p2e3 , bar_width,
                  color='navy',
                  label='mptcp')
 
-rects2 = plt.bar(index+margin + bar_width, np1pe3 , bar_width,
-                 color='indigo',
+rects2 = plt.bar(index+margin + bar_width, np2e3 , bar_width,
+                 color='m',
                  label='mptcp-New')
 
-rects3 = plt.bar(index+margin + 2*bar_width, tp1pe3, bar_width,
+rects3 = plt.bar(index+margin + 2*bar_width, tp2e3, bar_width,
                  color='seagreen',
                  label='tcp')
 

### 2017-04-27 main.tex
Color update 2
Updates to abstract and sections 1 through 3

@@
 Transport protocols recover from packet loss in order to provide reliable end-to-end communication and improve user experience. The efficiency of 
-loss recovery mechanisms influences the completion time of flows. In this paper we focus on state-of-the-art loss recovery mechanisms for TCP 
+loss recovery mechanisms influences the completion time of flows, and thus also the application performance as perceived by the
+end user. In this paper we focus on state-of-the-art loss recovery mechanisms for TCP 
 and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms, and based on the observations, 
-we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results using Linux 
+we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results using the Linux 
 Multipath TCP implementation show consistent end-to-end latency performance improvement in considered scenarios. 
@@
 Fast retransmit is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as an indication of loss. In the case 
-of RTO, a sender waits for the packet loss to trigger a timeout before retransmitting packet. During such a timeout, the sender cannot send any 
+of RTO, a sender waits for the packet loss to trigger a timeout before retransmitting the packet. During such a timeout interval, 
+the sender cannot retransmit any 
 packets. However, with FR the out-of-order packets trigger duplicate acknowledgements. The sender can start retransmission of missing packets 
-without waiting for timeout. For flows where there are sufficiently many packets to be transferred, FR is quicker than RTO in detecting and 
-recovering packet loss. Fast retransmit coupled with fast recovery can send data outstanding when there is enough data to send in long flows. 
-On the other hand, shorter flows more frequently depend on RTO for packet loss detection and recovery, thus incurring more latency. A single 
+without waiting for a timeout. For flows where there are sufficiently many packets to be transferred, FR is quicker than RTO in detecting and 
+recovering packet loss. During long flows, subsequent packets in the flow typically arrive at the receiver after a packet loss and based
+on the duplicate acknowledgements that they trigger, FR coupled with fast recovery can quickly retransmit the missing data. 
+On the other hand, shorter flows more frequently have to depend on RTO for packet loss detection and recovery, thus incurring more latency. A single 
 packet loss in a short flow may take many RTTs to detect and recover. This scenario is also applicable to the packets at the end of the flow 
@@
 
-TCP Loss Probe (TLP)~\cite{Flach:2013} is a mechanism to detect and recover from tail losses by employing a loss probe packet for a timeout shorter
+TCP Loss Probe (TLP)~\cite{Flach:2013} is a mechanism to detect and recover from tail losses by employing a loss probe packet sent after a timeout shorter
 than the RTO. Probe timeout enables short flows to recover packet loss at the tail, by triggering FR. It assumes that other algorithms such as 
@@
  level congestion window as well as subflow level congestion windows for each subflow. Data is split across the flows and each data packet is 
- scheduled on one of the subflows based on the minimum RTT. Packet losses in each subflow are detected and recovered in a similar fashion as that of 
+ scheduled on one of the subflows.
+The default MPTCP scheduler in Linux selects the subflow to use based on the shortest RTT.
+Packet losses in each subflow are detected and recovered in a similar fashion as that of
  TCP. Loss recovery can be conducted at different levels: the MPTCP connection level (meta level) and at subflow level. If the recovery is handled 
  at meta level, a lost packet may be rescheduled and retransmitted on the available subflow with the lowest RTT. If recovery is handled at the 
- subflow level, the packet must be retransmitted on the same subflow. An important part of loss recovery process at the subflow level is to follow 
- the TCP semantics and retransmit lost packets on the initial path of transfer. However, the MPTCP standard~\cite{rfc6824} does not specify exactly, 
- how the loss recovery should happen in the implementation and which subflow should retransmit the lost packets.
+ subflow level, the packet is retransmitted on the same subflow. An important part of the loss recovery process at the subflow level is to follow 
+ the TCP semantics and retransmit lost packets on the initial path of transfer. However, the MPTCP standard~\cite{rfc6824} does not specify exactly 
+ how the loss recovery should happen in the implementation and which subflow(s) should retransmit the lost packets.
 
@@
 Linux implementation of TLP for MPTCP in Section~\ref{impr}. In Section~\ref{eval}, we provide an evaluation of the proposed TLP for MPTCP 
-mechanism and conclude the paper in Section~\ref{conc}. Our experimental results indicate improvement in the performance of TLP by reducing 
-recovery time in the event of tail losses.
+mechanism. Our experimental results indicate a consistent improvement in the performance of TLP by reducing 
+recovery time in the event of tail losses. Finally, we conclude the paper in Section~\ref{conc}.
 
@@
 recovery~\cite{rfc6675}, Limited Transmit~\cite{rfc3042} and Early Retransmission~\cite{rfc5827} have led to improved retransmission 
-strategies for TCP. TLP is part of the various deployable loss recovery mechanisms proposed in~\cite{Flach:2013}, suggesting that current 
+strategies for TCP. TLP is part of the various deployable loss recovery mechanisms proposed in~\cite{Flach:2013}, suggesting that the evaluated 
 Linux TCP implementation provides room for improvement in several packet loss scenarios.
@@
 numbering enables MPTCP to respond to loss of packets by retransmitting the lost packets on an alternate path.  
+The Linux implementation of MPTCP uses a set of retransmission heuristics to handle retransmissions. In addition to the normal TCP retransmission 
+on the subflow level, the data outstanding on a timed out subflow 
+is rescheduled for transmission on a different subflow using timeout as the indicator. Fast retransmit on a subflow does not trigger 
+retransmission on another subflow.
 
-The Linux implementation of MPTCP uses a set of retransmission heuristics to handle retransmissions. The data outstanding on a timed out subflow 
-should be rescheduled for transmission on a different subflow using timeout as the indicator. Fast retransmit on a subflow will not trigger 
-retransmission on another subflow. The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 
+The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 
 4G vs WLAN, leading to flows with different delay characteristics. Significant delay differences in available paths can cause receive buffer 
 limitations, as the receiver has to wait for the data sent on a slow path, while buffering the data on the fast path. In such cases, opportunistic 
-retransmission~\cite{Costin} improves the latency by using fast path to retransmit the data sent on an another path. Authors in~\cite{fuso}, try 
+retransmission~\cite{Costin} improves the latency by using the fast path to retransmit the data sent on an another path. Authors in~\cite{fuso}, try 
 to exploit the path diversity by quickly retransmitting on the fastest paths. Such quick retransmission comes with a cost of redundant packets as 
@@
 Experiments are run in the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}.
-The client node connect to two wireless interfaces 3G or 4G and WLAN and the server node connect to a wired router.
+The client node connects to two wireless interfaces 3G or 4G and WLAN and the server node connects to a wired router.
 CORE uses the Linux networking stack in each node of the topology by using a lightweight virtualization technique known as Linux network namespaces.
@@
 and the server. So one has to wait both in time and packets for the second subflow establishment. We send two bursts of 80 packets each and drop 
-tail packets in the second burst at one of the interfaces. This is to ensure that the necessary and sufficient conditions for probe triggering 
-are met and tail loss probe is generated on that subflow. We calculate the burst completion time for the second burst that has specified tail 
+packets at the tail of the second burst at one of the interfaces. This is to ensure that the necessary and sufficient conditions for probe triggering 
+are met and a tail loss probe is generated on that subflow. We calculate the burst completion time for the second burst that has specified tail 
 losses. The objective of this experiment is to evaluate retransmission mechanisms with tail loss (single packet loss and multiple tail loss) and 
@@
 
-The setup has server and client with Linux supporting MPTCP running on them. The Client has two interfaces with one way delay on each interface 
-ranging from 20\,ms to 120\,ms. The intention is to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in the link delays. We 
+The setup has a server and a client with Linux supporting MPTCP running on them. The Client has two interfaces with one way delay on each interface 
+ranging from 20\,ms to 120\,ms. The intention is to evaluate the effect of symmetry, mild asymmetry and high asymmetry in the link delays. We 
 consider 5 scenarios with delay pairings 20\,ms-20\,ms, 20\,ms-30\,ms, 20\,ms-120\,ms, 30\,ms-20\,ms, 120\,ms-20\,ms to understand the effect of

### 2017-04-27 main.tex
small updates of abstract

@@
 \begin{abstract}
-Packet losses are known to affect the performance of latency sensitive applications in the Internet such as media streaming and gaming. 
+Packet losses are known to affect the performance of latency sensitive Internet applications such as media streaming and gaming. 
 Transport protocols recover from packet loss in order to provide reliable end-to-end communication and improve user experience. The efficiency of 
@@
 and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms, and based on the observations, 
-we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results using the Linux 
-Multipath TCP implementation show consistent end-to-end latency performance improvement in considered scenarios. 
+we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results, using the Linux 
+Multipath TCP implementation, show consistent end-to-end latency performance improvement in considered scenarios. 
 \end{abstract}
@@
 interruption. The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied 
-loss scenarios. In the case of single packet tail loss, lost packet is retransmitted on the alternate path at the same time as probe sent on primary
+loss scenarios. In the case of single packet tail loss, lost packet is retransmitted on an alternate path at the same time as probe sent on primary
 path. Since probe is the last transmitted packet, that is same as the retransmitted first outstanding packet in single packet tail loss. When two tail packets are lost,

### 2017-04-27 main.tex
small text changes in abstract, intro, and background.

@@
 end user. In this paper we focus on state-of-the-art loss recovery mechanisms for TCP 
-and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms, and based on the observations, 
+and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms and, based on the observations, 
 we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results, using the Linux 
@@
 Multipath TCP~(MPTCP)~\cite{rfc6824} is an experimental standard proposed as an extension to TCP~\cite{rfc793}. MPTCP allows devices with multiple interfaces to 
-transmit data simultaneously on both interfaces, thereby improving the throughput of end-to-end connections. Using MPTCP, a connection can have 
-multiple TCP subflows using different interfaces on different routes. Each TCP subflow imparts the functional behavior of a standard TCP flow.
+transmit data simultaneously on all interfaces, thereby improving the end-to-end throughput of connections. Using MPTCP, a connection can have 
+multiple TCP subflows using different interfaces on different routes. Each subflow imparts the functional behavior of a standard TCP flow.
 
-TCP has two mechanisms for detecting and recovering from packet losses: Fast Retransmit (FR) and Retransmission Timeout~(RTO)~\cite{Flach:2013}. 
+TCP has two mechanisms for detecting and recovering from packet loss: Fast Retransmit (FR) and Retransmission Timeout~(RTO)~\cite{Flach:2013}. 
 Fast retransmit is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as an indication of loss. In the case 
@@
 
-In this paper, we analyze the retransmission behavior of MPTCP in selected tail loss cases to identify the scope for improvement. We propose a 
+In this paper, we analyze the retransmission behavior of MPTCP in selected tail loss cases to identify the scope for improvement. Based on this analysis, we propose a 
 change in the retransmission strategy of MPTCP to improve the latency performance. We provide related research on MPTCP retransmission mechanisms in 
-Section~\ref{relwork}, and describe the experimental setup in Section~\ref{exsetup}. We discuss observations from the experiments on state 
-of the art retransmission behavior of TCP and MPTCP in Section~\ref{disc}. Based on the observations, we propose an improvement to the current 
+Section~\ref{relwork}, and describe the experimental setup in Section~\ref{exsetup}. We discuss observations from the experiments on
+state-of-the-art retransmission behavior of TCP and MPTCP in Section~\ref{disc}. Based on the observations, we propose an improvement to the current 
 Linux implementation of TLP for MPTCP in Section~\ref{impr}. In Section~\ref{eval}, we provide an evaluation of the proposed TLP for MPTCP 
-mechanism. Our experimental results indicate a consistent improvement in the performance of TLP by reducing 
+mechanism. Our experimental results indicate a consistent performance improvement of TLP by reducing 
 recovery time in the event of tail losses. Finally, we conclude the paper in Section~\ref{conc}.
@@
 
-This paper analyzes the performance of state-of-the-art loss recovery mechanism for MPTCP and presents a case for improvement in the tail loss 
+This paper analyzes the performance of state-of-the-art loss recovery mechanisms for MPTCP and presents a case for improvement in the tail loss 
 scenario. For understanding the loss recovery and retransmission policies in the MPTCP Linux implementation (version 0.91), we consider several specific 

### 2017-04-27 main.tex
minor text fixes in III.

@@
 %\kern-1em
-\subsection{Testing retransmission mechanism with deterministic loss patterns}
+\subsection{Testing retransmission mechanisms with deterministic loss patterns}
 In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a 

### 2017-04-27 main.tex
Updates to sec. IV

--- 2017-04-27 main.tex
@@
 \begin{abstract}
-Packet losses are known to affect the performance of latency sensitive Internet applications such as media streaming and gaming. 
+Packet losses are known to affect the performance of latency sensitive applications in the Internet such as media streaming and gaming. 
 Transport protocols recover from packet loss in order to provide reliable end-to-end communication and improve user experience. The efficiency of 
@@
 and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms, and based on the observations, 
-we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results, using the Linux 
-Multipath TCP implementation, show consistent end-to-end latency performance improvement in considered scenarios. 
+we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results using the Linux 
+Multipath TCP implementation show consistent end-to-end latency performance improvement in considered scenarios. 
 \end{abstract}
@@
 Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
-starts with the results using TCP and compares it with that of MPTCP. Figure~\ref{tcpf} represents the comparison of TCP burst completion times. Path
+starts with the results using TCP and compares it with that of MPTCP. Figure~\ref{tcpf} represents the comparison of TCP burst completion times
+in the considered scenarios. Path
 asymmetry is irrelevant in this case as TCP uses a single path.
@@
 specification that adds 200\,ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts for the delayed 
-ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for PTO. In the case of two packet tail loss, as shown 
-in Figure~\ref{t2p}, TLP performs better than the RTO retransmission for lower delay values. As the delay on the loss path increases RTO retransmission performs 
-slightly better than TLP. In 120\,ms case, PTO is very close to RTO expiration and the second packet is recovered using ER which incurs additional 
-delay in case of TLP. In RTO retransmission, all unacknowledged packets are retransmitted on a RTO expiration.
+ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for the PTO. Hence in this special case the value of 
+the retransmission timeout is used in the loss recovery in both RTO and TLP. This can be avoided by tuning the ACKs as discussed 
+in~\cite{Rajiullah:2015}.
 
-Hence retransmission timeout is the mode of loss recovery in both RTO and TLP. This can be avoided by tuning the ACKs as discussed 
-in~\cite{Rajiullah:2015}. In the case of two packet tail loss, TLP performs better than RTO retransmission as it employs a combination of TLP 
-and ER to recover from losses. If the probe or retransmission is lost, then TLP still performs better than RTO retransmission as it waits for one 
+In the case of two packet tail loss, as shown 
+in Figure~\ref{t2p}, TLP performs better than the RTO retransmission for lower delay values. As the delay on the loss path increases, RTO retransmission performs 
+slightly better than TLP. In the 120\,ms case, PTO is very close to RTO expiration and in the case of TLP the second packet is recovered 
+using ER which incurs additional delay. In RTO retransmission, the second packet is retransmitted immediately when the acknowledgement
+for the RTO retransmission is received.
+
+If the probe or retransmission is lost, as shown 
+in Figure~\ref{tcpf}, then TLP performs better than RTO retransmission as it waits for one 
 probe timeout and one RTO experiation instead of two RTO expirations, thus avoiding the exponential backoff of the RTO.
@@
 on single packet tail loss calculated for each delay configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no 
-difference in performance between RTO retransmission and TLP for cases when both paths are symmetric or the first path has lowest RTT. RTO performs 
+difference in performance between RTO retransmission and TLP for cases when both paths are symmetric or the first path has the lowest RTT. RTO performs 
 better than TLP when the first path is not the lowest RTT path. In such scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens 
@@
 as probe to trigger fast retransmit of the packets in between. In order to observe the performance of n-degree tail loss, we consider dropping the 
-last two packets. The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the other 
-two settings in all scenarios except when packets are lost on a high delay interface. In this scenario, the MPTCP retransmission follows the 
+last two packets. The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the RTO 
+setting in all scenarios except when packets are lost on a high delay interface. In this scenario, the MPTCP retransmission follows the 
 behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of the last packet on the alternate 
@@
 
-To further study the performance of TLP in the case when there is longer path interruption on one path, we tried to drop the probe packet along 
+To further study the performance of TLP in the case when there is a longer path interruption on one path, we tried to drop the probe packet along 
 with the last packet. In this scenario, the RTO setting results in much lower burst completion times. Comparing with the performance of TLP in 
-TCP from Figure~\ref{t1pp}, where we saw that use of TLP improved the burst completion time, TLP in MPTCP did not improve the performance. The 
-reason lies in the way the retransmission and loss probe transmission is carried out in MPTCP. Retransmission occurs after an RTO time on both 
+TCP from Figure~\ref{t1pp}, where we saw that the use of TLP improved the burst completion time, TLP in MPTCP did not improve the performance. The 
+reason lies in the way the retransmission and loss probe transmission are carried out in MPTCP. After an RTO, retransmissions occur on both 
 paths. Even if the retransmission fails on one path, the client receives the packet on the other path as shown in Figure~\ref{timing1PP}. For TLP, 

### 2017-04-27 main.tex

@@
 end user. In this paper we focus on state-of-the-art loss recovery mechanisms for TCP 
-and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms, and based on the observations, 
-we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results using the Linux 
-Multipath TCP implementation show consistent end-to-end latency performance improvement in considered scenarios. 
+and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms and, based on the observations, 
+we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results, using the Linux 
+Multipath TCP implementation, show consistent end-to-end latency performance improvement in considered scenarios. 
 \end{abstract}

### 2017-04-27 main.tex
small changes

@@
 \usepackage{subcaption}
+\usepackage{balance}
 
@@
 If the probe or retransmission is lost, as shown 
-in Figure~\ref{tcpf}, then TLP performs better than RTO retransmission as it waits for one 
-probe timeout and one RTO experiation instead of two RTO expirations, thus avoiding the exponential backoff of the RTO.
+in Figure~\ref{t1pp}, then TLP performs better than RTO retransmission as it waits for one 
+probe timeout and one RTO expiration instead of two RTO expirations, thus avoiding the exponential backoff of the RTO.
 

### 2017-04-27 main.tex
balance last page
Updates to sec V

--- 2017-04-27 main.tex
@@
 avoiding a costly retransmission timeout. Our results show that the current TLP implementation for MPTCP helps improving the latency performance in 
-cases when the paths are symmetric or the first path~(for the packet) has lowest RTT. However, it reduces the performance in cases when the first path 
-is not the lowest RTT path or when there is longer path interruption.
+cases when the paths are symmetric or the path where the packet is first transmitted has the lowest RTT. However, it reduces the performance in cases when the first path 
+is not the lowest RTT path or when there is longer path interruptions.
 
@@
 The current implementation of TLP is at the TCP flow level. In the event of RTO timeout, the MPTCP retransmission happens with a re-injection in 
-to the scheduler along with sending the lost packet on the same path. But in the event of Probe timeout, the loss probe packet is being sent on 
-the same path without being re-injected in to the scheduler. We propose a modification to the probe sending mechanism for MPTCP by re-injecting 
+to the scheduler along with sending the lost packet on the same path. But in the event of Probe timeout, the loss probe packet is being sent only on 
+the same path without being re-injected in to the MPTCP scheduler. We propose a modification to the probe sending mechanism for MPTCP by re-injecting 
 the probe packet in to the MPTCP scheduler in the event of PTO. As the probe timeout is meant for acting faster than the RTO timeout with 
 essentially the same consequence as a retransmission, it is relevant to follow the same process as upon an RTO. It is expected that this 
-modification is useful when the alternate path has lower delay than the packet original transmission path or when there is longer path 
+modification is useful when the alternate path has lower delay than the path where the packet was originally transmitted or when there is a longer path 
 interruption. The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied 
-loss scenarios. In the case of single packet tail loss, lost packet is retransmitted on an alternate path at the same time as probe sent on primary
-path. Since probe is the last transmitted packet, that is same as the retransmitted first outstanding packet in single packet tail loss. When two tail packets are lost,
-probe sent on the primary is the last transmitted packet and first outstanding packet is retransmitted on the alternate path. This difference in packet
-sequence numbers is expected on the two paths due to the nature of recovery mechanism used in each path. In the case of packet and probe loss in the 
-primary path, the alternate path retransmission is same as the probe packet.
+loss scenarios. In the case of single packet tail loss, the lost packet is retransmitted on an alternate path at the same time as the probe 
+is sent on the original
+path. Since the probe is the last transmitted packet, that is the same as the retransmitted first outstanding packet for single packet tail loss. When two tail packets are lost, the
+probe sent on the original path is the last transmitted packet and the first outstanding packet is retransmitted on the alternate path. This difference in packet
+sequence numbers is expected on the two paths due to the nature of the recovery mechanism used on each path. In the case of packet and probe loss on the 
+original path, the alternate path retransmission is the same as the probe packet.
 

### 2017-04-27 main.tex

@@
 
+\balance
+
 %\kern-1em

### 2017-04-27 main.tex
spell
Nit in IV

@@
 behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of the last packet on the alternate 
-path is faster than the probe on the primary path.
+path is faster than the probe on the original path.
 

### 2017-04-27 main.tex
updated section v slightly

--- 2017-04-27 main.tex
@@
 The goal of Tail Loss Probe~(TLP) is to reduce tail latency of short flows. It achieves this by converting retransmission timeouts (RTOs) occurring
-due to tail losses into fast recovery. In Open state of TCP connection, a sender is receiving in-sequence ACKs and has not detected any lost segments.
-TLP transmits one packet in two round-trips when a connection is in Open state and is not receiving any ACKs. The transmitted packet, or loss probe, 
+due to tail losses into fast recovery. The Open state of a TCP connection is whenever a sender receives in-sequence ACKs and has not detected any lost segments.
+If TCP does not receive any ACKs within two RTTs while in the Open state, TLP transmits one packet. The transmitted packet, or loss probe, 
 can be either new or a retransmission. When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus 

### 2017-04-27 main.tex

--- 2017-04-27 main.tex
@@
 (or tail), of a long flow. Although RTO is reliable even in tail loss recovery, it is not efficient in terms of recovery completion time. Tail 
-losses are one of the major causes of end-to-end latency detoriation in modern networks~\cite{Flach:2013}.
+losses are one of the major causes of end-to-end latency deterioration in modern networks~\cite{Flach:2013}.
 

### 2017-04-27 main.tex
Updates to sec VI

--- 2017-04-27 main.tex
@@
 performs better than MPTCP in asymmetric scenarios with a lower delay on the alternate path. MPTCP-New achieves this improvement by retransmitting
- lost packet on the alternate path much before the RTO. It is also observed that there is no performance degradation when the alternate path has 
- higher delay. Probe packet on the primary path reaches faster than the retransmitted path and completes loss recovery at MPTCP level. In two packet 
+ the lost packet on the alternate path much before the RTO. It is also observed that there is no performance degradation when the alternate path has 
+ higher delay. The probe packet on the original path reaches faster than the retransmitted packet on the alternate path and completes loss recovery at the MPTCP level. In two packet 
  tail loss scenario as in Figure~\ref{2pn}, MPTCP-New performs better than TCP in all scenarios and better than MPTCP in all scenarios except when 
- the alternate path has a significantly higher delay than the original path, where performance is the same. In the single packet together with probe 
- loss scenario as in Figure~\ref{1ppn}, MPTCP-New performs better than both MPTCP and TCP in all scenarios with PTO triggering the probe retransmission
- on alternate path much earlier than the RTO.
+ the alternate path has a significantly higher delay than the original path, where performance is the same. For this scenario, MPTCP-New achieves
+ this gain as the retransmission on the alternate path recovers the first lost packet and the probe packet on the original path recovers the
+ second lost packet, and thus both lost packets can be recovered in parallel. 
+ In the single packet together with probe 
+ loss scenario as in Figure~\ref{1ppn}, MPTCP-New performs better than both MPTCP and TCP in all scenarios with PTO triggering the retransmission
+ on the alternate path much earlier than the RTO.
 

### 2017-04-27 main.tex
Updates to VII

@@
 This paper provides a short comparative analysis of retransmission behavior with TCP and MPTCP protocols using various loss patterns for tail loss.
-The three cases of tail loss together with deterministic loss patterns allow us to dissect the retransmission behavior at packet level to 
+The three studied cases of tail loss together with deterministic loss patterns allow us to dissect the retransmission behavior at packet level to 
 understand possible improvements.
-We propose a less conservative approach to trigger retransmission on an alternate path in the event of tail loss probe timeout, which otherwise 
-triggered with an RTO. Our emulation experiments  using a modified Linux implementation, show that the proposed approach in fact improves the 
+We propose a less conservative approach to trigger retransmission on an alternate path in the event of tail loss probe timeout, which is otherwise 
+triggered only with an RTO. Our emulation experiments  using a modified Linux implementation, show that the proposed approach in fact improves the 
 burst completion time in most cases and equals the existing implementation in other cases. The proposed approach improves the performance when 
@@
 Temporary path failures can cause probe loss along with packet loss. The approach is very effective in case of probe loss which otherwise 
-incur RTO timeout to trigger retransmission on either path. However, the latency improvement comes at a cost of additional network traffic 
-from the retransmitted packets. This study is limited to emulated synthetic traffic and static drop pattern that serve the purpose of localizing 
+incur RTO timeout to trigger retransmission on either path. The latency improvement comes at a cost of additional network traffic 
+from the retransmitted packets.
+However, this additional traffic is small and subject to normal congestion control.
+This study is limited to emulated synthetic traffic and static drop patterns that serve the purpose of localizing
 the loss events. A comprehensive study covering various traffic types and loss patterns is considered for future investigation. Further, we plan 
